---
title: "MOFA_MAGEL2"
output: html_document
date: "2025-02-04"
---

Loading the libraries and reading the preprocessed MOFA File

```{r}
library("MOFAdata")
library("MOFA2")
library("biomaRt")
library("ggplot2")
library("data.table")
library("dplyr")
library("gprofiler2")
library("ggplot2")
library("viridis") 
library("tidyr")   
library("tibble") 
library("RColorBrewer")
library("stringr") 
library("forcats")
library("grid")

MOFA_out<- readRDS("./MOFA_PRU3.rds") 
MOFA_out@samples_metadata$group <- gsub("[1-3]", "", MOFA_out@samples_metadata$sample)
MOFA_out@samples_metadata$group <- as.factor(MOFA_out@samples_metadata$group)
model <- MOFA_out

```

Overview & Variance Plots

```{r}
plot_data_overview(MOFA_out)

plot_variance_explained(MOFA_out, x="view", y="factor")
plot_variance_explained(MOFA_out, x="group", y="factor", plot_total = T)[[2]]   + theme(axis.text.x = element_text(angle = 45,hjust = 1),
        text = element_text(size = 16))

plot_factors(MOFA_out, 
  factors = 1:5,
  color_by = "group"
)

```

Plotting all samples based on the 15 factors

```{r}
# Load ggplot2 if it's not already loaded
library(ggplot2)

plot_factor(MOFA_out,
  factor = 1:15,
  color_by = "group",
  dot_size = 0 # Make the original points invisible
) +
  # Add a new layer of points with your desired settings
  geom_point(size = 4, alpha = 0.7, stroke = 0) +
  theme(
    axis.text.x = element_blank(),  # hides x-axis labels
    axis.ticks.x = element_blank(), # optionally hide x-axis ticks too
    text = element_text(size = 14)
  )


```


```{r}
# --- 1. Load necessary libraries ---
library(MOFA2)
library(ggplot2)
library(dplyr)
library(stringr)

# --- 2. Helper function to get colors ---
gg_color_hue <- function(n) {
  hues <- seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}


# --- 4. Prepare MOFA Metadata (Copied from your app) ---
# This adds the "Mutation" and "Background" columns needed for plotting
MOFA_out@samples_metadata <- MOFA_out@samples_metadata %>%
  mutate(
    Background = str_extract(sample, "(?<=_)(8330|MGH)"),
    Background = factor(Background),
    MutationPrefix = str_extract(sample, "^[A-Za-z]+"),
    Mutation = case_when(
      MutationPrefix == "WT" ~ "WT",
      MutationPrefix == "DelC" ~ "DelC",
      MutationPrefix == "DupC" ~ "DupC",
      MutationPrefix == "HetDel" ~ "HetDel",
      MutationPrefix == "HomoDel" ~ "HomoDel",
      MutationPrefix == "PWSTI" ~ "PWSTI",
      MutationPrefix == "PWSTII" ~ "PWSTII",
      TRUE ~ MutationPrefix
    ),
    Mutation = factor(Mutation, levels = c("DelC", "DupC", "HetDel", "HomoDel", "PWSTI", "PWSTII", "WT")),
    Mutation_BG = paste0(Mutation, "_", Background)
  ) %>%
  mutate(group = Mutation) %>%
  dplyr::select(-MutationPrefix)

# --- 5. Prepare data for plotting (from the app) ---
factors_df <- get_factors(MOFA_out, factors = 1:15, as.data.frame = TRUE)
meta_df <- samples_metadata(MOFA_out)
plot_df <- left_join(factors_df, meta_df, by = "sample")

# --- 6. Define Colors for 'Mutation' (Genotype) ---
mutation_levels <- levels(plot_df$Mutation)
mutation_colors <- gg_color_hue(length(mutation_levels))
names(mutation_colors) <- mutation_levels

# --- 7. Generate the Final Plot ---
# This is the plotting code from the app, with your changes applied
final_plot <- ggplot(plot_df, aes(x = factor, y = value,
                                 fill = Mutation,    # Map fill to Mutation
                                 color = Mutation,   # Map color to Mutation (to remove outline)
                                 shape = Background)) +
  geom_vline(xintercept = 1:14 + 0.5, linetype = "dashed", color = "grey70") +
  
  # Use stroke = 0 to remove point outlines
  geom_point(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8),
             size = 3, alpha = 0.8, stroke = 0) +
  
  # Use shapes 25 (triangle down) and 21 (circle)
  scale_shape_manual(name = "Background", values = c("8330" = 25, "MGH" = 21)) +
  
  # Add matching fill and color scales
  scale_fill_manual(name = "Genotype", values = mutation_colors) +
  scale_color_manual(name = "Genotype", values = mutation_colors, guide = "none") +
  
  # Add the rest of the theme from the app
  labs(x = "Factor", y = "Factor Value") +
  theme_bw() +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    text = element_text(size = 14, color = "#545252"),
    axis.text = element_text(color = "#545252"),
    legend.position = "top"
  )

# --- 8. Display the plot ---
print(final_plot)
```



Correlation between the factors

```{r}

# --- Flipped Correlation Heatmap with Circles (RdBu Palette) ---

# 1. Load the required libraries
library(corrplot)
library(RColorBrewer) # Added for brewer.pal()

# 2. Get the correlation matrix (same as before)
factor_list <- get_factors(MOFA_out, groups = "all")
all_factors_matrix <- do.call(rbind, factor_list)
factor_cor_matrix <- cor(all_factors_matrix)

# 3. Define the color palette exactly as you requested
#    (This is a diverging blue-to-red palette)
my_colors <- colorRampPalette(rev(brewer.pal(n = 10, name = "RdBu")))(100)

# 4. Plot using corrplot
corrplot(
  factor_cor_matrix,
  method = "circle",      # Use circles
  col = my_colors,        # Use your specified palette
  order = "original",     # Keep the 1-15 order
  is.corr = TRUE,
  tl.col = "black",       # Color of text labels
  tl.srt = 45,            # Rotate text labels
  cl.lim = c(-1, 1)       # Enforce the color limit from -1 to 1
)


# --- Plot for Factor 1 (from your original code) ---

plot_factor(MOFA_out, 
  factors = 1, 
  color_by = "group") +
theme(axis.text.x = element_text(angle = 45,hjust = 1),
      text = element_text(size = 16))
```


UMAP
```{r}
# Load necessary libraries
library(MOFA2)
library(ggplot2)
library(stringr)
library(dplyr)
library(ggExtra) # Added for marginal plots

# Helper function to generate the default ggplot2 color palette
gg_color_hue <- function(n) {
  hues <- seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}

# --- 1. Ensure Metadata and UMAP are prepared ----

# Ensure sample names are in rownames for easy merging later
stopifnot("sample" %in% colnames(MOFA_out@samples_metadata))
rownames(MOFA_out@samples_metadata) <- MOFA_out@samples_metadata$sample

# Run UMAP if it hasn't been run already
if (!"UMAP" %in% names(MOFA_out@dim_red)) {
  set.seed(42) # for reproducibility
  MOFA_out <- run_umap(MOFA_out)
}

# --- 2. Create the Data Frame for Plotting ----

# Extract only the UMAP coordinate columns
umap_coords <- as.data.frame(MOFA_out@dim_red$UMAP) %>%
  dplyr::select(UMAP1, UMAP2)

# Align metadata to the UMAP coordinates using the robust rownames
metadata_aligned <- MOFA_out@samples_metadata[rownames(umap_coords), , drop = FALSE]

# Combine coordinates and metadata, and create clean 'Background' and 'Mutation' columns
plot_df <- cbind(umap_coords, metadata_aligned) %>%
  mutate(
    Background = factor(case_when(
      str_detect(group, "_80$") ~ "8330", # Using 8330 for clarity
      str_detect(group, "_MGH$") ~ "MGH",
      TRUE ~ "Unknown"
    )),
    Mutation = factor(str_replace(group, "_(80|MGH)$", ""))
  )

# --- 3. Generate the Base UMAP Plot ----
# (Note: Renamed to 'p_umap_base' and removed the print() command)

# Define colors for the Mutation types
mutation_levels <- levels(plot_df$Mutation)
mutation_colors <- gg_color_hue(length(mutation_levels))
names(mutation_colors) <- mutation_levels

# Create the base plot object
p_umap_base <- ggplot(plot_df, aes(x = UMAP1, y = UMAP2, fill = Mutation, color = Mutation, shape = Background)) +
  geom_point(size = 3.5, alpha = 0.8) +
  
  # This scale creates the color legend for "Mutation"
  scale_fill_manual(name = "Mutation", values = mutation_colors) +
  
  # This scale matches the outlines but hides its own legend
  scale_color_manual(name = "Mutation", values = mutation_colors, guide = "none") +
  
  # This scale defines the shapes
  scale_shape_manual(name = "Background", values = c("8330" = 25, "MGH" = 21, "Unknown" = 4)) +
  
  labs(
    title = "UMAP of MOFA Factors",
    subtitle = "Colored by Mutation, Shaped by Background"
  ) +
  theme_bw(base_size = 14) +
  theme(
    legend.position = "right",
    panel.grid = element_blank() # Remove grid lines
  )

# --- 4. Add Marginal Densities and Display ----
message("--- Generating MOFA UMAP with Marginal Densities ---")

# Use ggMarginal to add density plots
# It will use the 'color = Mutation' mapping for grouping
umap_marginal_plot <- ggMarginal(
  p_umap_base,
  type = "density",
  groupColour = TRUE, # Use the 'color = Mutation' mapping
  groupFill = TRUE,   # Fill the densities based on the same group
  alpha = 0.4         # Set transparency
)

# Display the final plot
print(umap_marginal_plot)
```





Plotting the top weights for each factor

```{r}
plot_top_weights(MOFA_out,
 view = "RNA",
 factor = 1:5,
 nfeatures = 10,     # Top number of features to highlight
 scale = T           # Scale weights from -1 to 1
)

plot_weights(MOFA_out,
 view = "RNA",
 factor = 1:5,
 nfeatures = 5,     # Top number of features to highlight
 scale = T, 
 text_size = 3 # Scale weights from -1 to 1
)

plot_top_weights(MOFA_out,
 view = "protein",
 factor = 1:5,
 nfeatures = 10,     # Top number of features to highlight
 scale = T           # Scale weights from -1 to 1
)

plot_weights(MOFA_out,
 view = "protein",
 factor = 1:5,
 nfeatures = 5,     # Top number of features to highlight
 scale = T, 
 text_size = 3 # Scale weights from -1 to 1
)


plot_top_weights(MOFA_out,
 view = "Ubiq",
 factor = 1:5,
 nfeatures = 10,     # Top number of features to highlight
 scale = T           # Scale weights from -1 to 1
)

plot_weights(MOFA_out,
 view = "Ubiq",
 factor = 1:5,
 nfeatures = 5,     # Top number of features to highlight
 scale = T, 
 text_size = 3 # Scale weights from -1 to 1
)

```

Heatmap of the most prominent features in Factor 1

```{r}

TARGET_FACTOR <- 1
N_FEATURES <- 25 

# Aesthetic parameters
COLOR_BREAKS_MIN <- -2
COLOR_BREAKS_MAX <- 2
ROW_LABEL_WIDTH <- 15 

color_palette <- colorRampPalette(rev(brewer.pal(n = 7, name = "RdYlBu")))(100)
color_breaks <- seq(COLOR_BREAKS_MIN, COLOR_BREAKS_MAX, length.out = 101)



### --- RNA View --- ###

# Get top features for this view
rna_weights <- get_weights(MOFA_out, views = "RNA", factors = TARGET_FACTOR)[[1]]
rna_top_features <- names(sort(abs(rna_weights[, 1]), decreasing = TRUE)) %>% head(N_FEATURES)

# Create truncated labels for displaying on the plot
rna_truncated_labels <- str_trunc(rna_top_features, width = ROW_LABEL_WIDTH, side = "right")

# Generate the plot
plot_data_heatmap(
  MOFA_out,
  view = "RNA",
  factor = TARGET_FACTOR,
  features = rna_top_features, 
  denoise = TRUE,
  cluster_rows = TRUE, cluster_cols = FALSE,
  show_rownames = TRUE, show_colnames = TRUE,
  scale = "row",
  main = paste("Top", N_FEATURES, "RNA Features (Factor", TARGET_FACTOR, ")"),
  color = color_palette,
  breaks = color_breaks,
  labels_row = rna_truncated_labels
)


### --- Protein View --- ###

# Get top features for this view
protein_weights <- get_weights(MOFA_out, views = "protein", factors = TARGET_FACTOR)[[1]]
protein_top_features <- names(sort(abs(protein_weights[, 1]), decreasing = TRUE)) %>% head(N_FEATURES)

# Create truncated labels for displaying on the plot
protein_truncated_labels <- str_trunc(protein_top_features, width = ROW_LABEL_WIDTH, side = "right")

# Generate the plot
plot_data_heatmap(
  MOFA_out,
  view = "protein",
  factor = TARGET_FACTOR,
  features = protein_top_features, 
  denoise = TRUE,
  cluster_rows = TRUE, cluster_cols = FALSE,
  show_rownames = TRUE, show_colnames = TRUE,
  scale = "row",
  main = paste("Top", N_FEATURES, "Protein Features (Factor", TARGET_FACTOR, ")"),
  color = color_palette,
  breaks = color_breaks,
  labels_row = protein_truncated_labels 
)


### --- Ubiquitination View --- ###

# Get top features for this view
ubiq_weights <- get_weights(MOFA_out, views = "Ubiq", factors = TARGET_FACTOR)[[1]]
ubiq_top_features <- names(sort(abs(ubiq_weights[, 1]), decreasing = TRUE)) %>% head(N_FEATURES)

# Create truncated labels for displaying on the plot
ubiq_truncated_labels <- str_trunc(ubiq_top_features, width = ROW_LABEL_WIDTH, side = "right")

# Generate the plot
plot_data_heatmap(
  MOFA_out,
  view = "Ubiq",
  factor = TARGET_FACTOR,
  features = ubiq_top_features, 
  denoise = TRUE,
  cluster_rows = TRUE, cluster_cols = FALSE,
  show_rownames = TRUE, show_colnames = TRUE,
  scale = "row",
  main = paste("Top", N_FEATURES, "Ubiq Features (Factor", TARGET_FACTOR, ")"),
  color = color_palette,
  breaks = color_breaks,
  labels_row = ubiq_truncated_labels 
)


```

Running GSEA for all positive and negative weights on RNA, Protein and Ubi

```{r}
utils::data(MSigDB_v6.0_C5_human)

data("MSigDB_v6.0_C5_human")

mart <- useMart("ensembl", dataset = "hsapiens_gene_ensembl")

conversion <- getBM(
  attributes = c("ensembl_gene_id", "hgnc_symbol"),
  filters = "ensembl_gene_id",
  values = colnames(MSigDB_v6.0_C5_human),
  mart = mart
)

valid_conversion <- conversion[!is.na(conversion$hgnc_symbol) & conversion$hgnc_symbol != "", ]
common_ensembl_ids <- intersect(colnames(MSigDB_v6.0_C5_human), valid_conversion$ensembl_gene_id)
MSigDB_subset <- MSigDB_v6.0_C5_human[, common_ensembl_ids]

# Re-run conversion filtering to match the subsetted MSigDB object
valid_conversion_subset <- valid_conversion[valid_conversion$ensembl_gene_id %in% common_ensembl_ids, ]
# Ensure the order of conversion matches the order of columns in MSigDB_subset
conversion_ordered <- valid_conversion_subset[match(colnames(MSigDB_subset), valid_conversion_subset$ensembl_gene_id), ]

# Rename columns
colnames(MSigDB_subset) <- conversion_ordered$hgnc_symbol
MSigDB_v6.0_C5_human_renamed <- MSigDB_subset

MSigDB_v6.0_C5_human_RNA <- MSigDB_v6.0_C5_human_renamed
MSigDB_v6.0_C5_human_Protein <- MSigDB_v6.0_C5_human_renamed
MSigDB_v6.0_C5_human_Ubiq <- MSigDB_v6.0_C5_human_renamed

colnames(MSigDB_v6.0_C5_human_RNA) <- paste(colnames(MSigDB_v6.0_C5_human_RNA), "RNA", sep = "_")
colnames(MSigDB_v6.0_C5_human_Protein) <- paste(colnames(MSigDB_v6.0_C5_human_Protein), "protein", sep = "_")
colnames(MSigDB_v6.0_C5_human_Ubiq) <- paste(colnames(MSigDB_v6.0_C5_human_Ubiq), "ubiq", sep = "_")


# --- Run GSEA for positive weights 
res.positive_RNA <- run_enrichment(MOFA_out ,
  feature.sets = MSigDB_v6.0_C5_human_RNA,
  view = "RNA",
  factors = 1:10,
  sign = "positive",
  verbose = TRUE
)

res.positive_Protein  <- run_enrichment(MOFA_out,
  feature.sets = MSigDB_v6.0_C5_human_Protein,
  view = "protein",
  factors = 1:10,
  sign = "positive",
  verbose = TRUE
)

res.positive_Ubiq  <- run_enrichment(MOFA_out,
  feature.sets = MSigDB_v6.0_C5_human_Ubiq,
  view = "Ubiq", 
  factors = 1:10,
  sign = "positive",
  verbose = TRUE
)



# --- Run GSEA for negative weights 
res.negative_RNA <- run_enrichment(MOFA_out ,
  feature.sets = MSigDB_v6.0_C5_human_RNA,
  view = "RNA",
  factors = 1:10,
  sign = "negative",
  verbose = TRUE
)

res.negative_Protein  <- run_enrichment(MOFA_out,
  feature.sets = MSigDB_v6.0_C5_human_Protein,
  view = "protein",
  factors = 1:10,
  sign = "negative",
  verbose = TRUE
)

res.negative_Ubiq  <- run_enrichment(MOFA_out,
  feature.sets = MSigDB_v6.0_C5_human_Ubiq,
  view = "Ubiq", 
  factors = 1:10,
  sign = "negative",
  verbose = TRUE
)

save(res.positive_RNA, res.positive_Protein, res.positive_Ubiq,
     res.negative_RNA, res.negative_Protein, res.negative_Ubiq,
     file = "gsea_results.RData")


```


Plotting the genes / proteins contributing the most to the top weights in factor 1
```{r}

plot_enrichment_detailed(
  enrichment.results = res.positive_RNA,
  factor = 1,
  max.pathways = 10
) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        text = element_text(size = 14)) +
  labs(title = "Detailed GSEA Plot (Factor 1, Positive Weights)",
       subtitle = "RNA View - MSigDB C5 (GO)")

plot_enrichment_detailed(
  enrichment.results = res.positive_Protein,
  factor = 1,
  max.pathways = 10
) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        text = element_text(size = 14)) +
  labs(title = "Detailed GSEA Plot (Factor 1, Positive Weights)",
       subtitle = "Protein View - MSigDB C5 (GO)")

plot_enrichment_detailed(
  enrichment.results = res.positive_Ubiq,
  factor = 1,
  max.pathways = 10
) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        text = element_text(size = 14)) +
  labs(title = "Detailed GSEA Plot (Factor 1, Positive Weights)",
       subtitle = "Ubiq View - MSigDB C5 (GO)")

```



UMAP Analysis and Plotting of MOFA Factors
```{r}
# Helper function to generate the default ggplot2 color palette
gg_color_hue <- function(n) {
  hues <- seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}

# Ensure that 'sample' is a column in the metadata and set it as rownames
stopifnot("sample" %in% colnames(MOFA_out@samples_metadata))
rownames(MOFA_out@samples_metadata) <- MOFA_out@samples_metadata$sample

# Run UMAP if it hasn't been run already
if (!"UMAP" %in% names(MOFA_out@dim_red)) {
  set.seed(42) # for reproducibility
  MOFA_out <- run_umap(MOFA_out)
}

# Extract only the UMAP coordinate columns to prevent column name conflicts
umap_coords <- as.data.frame(MOFA_out@dim_red$UMAP) %>%
  dplyr::select(UMAP1, UMAP2)

# Align metadata to the UMAP coordinates using the robust rownames
metadata_aligned <- MOFA_out@samples_metadata[rownames(umap_coords), , drop = FALSE]

# Combine UMAP coordinates and metadata into a single data frame for plotting
plot_df <- cbind(umap_coords, metadata_aligned) %>%
  mutate(
    Background = factor(case_when(
      str_detect(group, "_80$") ~ "80",
      str_detect(group, "_MGH$") ~ "MGH",
      TRUE ~ "Unknown"
    )),
    Mutation = factor(str_replace(group, "_(80|MGH)$", ""))
  )

# --- Plot 1: Colored by Background ---
background_colors <- c("80" = gg_color_hue(2)[1], "MGH" = gg_color_hue(2)[2], "Unknown" = "grey50")

umap_plot_background <- ggplot(plot_df, aes(x = UMAP1, y = UMAP2, color = Background)) +
  # Updated point size and transparency
  geom_point(size = 8, alpha = 0.6) +
  scale_color_manual(values = background_colors) +
  labs(
    title = "UMAP of MOFA Factors",
    subtitle = "Colored by Background",
    color = "Background"
  ) +
  theme_bw(base_size = 12) +
  # Updated theme to remove grid lines
  theme(
    legend.position = "right",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

# --- Plot 2: Colored by Mutation Type ---
mutation_levels <- levels(plot_df$Mutation)
mutation_colors <- gg_color_hue(length(mutation_levels))
names(mutation_colors) <- mutation_levels

umap_plot_mutation <- ggplot(plot_df, aes(x = UMAP1, y = UMAP2, color = Mutation)) +
  # Updated point size and transparency
  geom_point(size = 8, alpha = 0.6) +
  scale_color_manual(values = mutation_colors) +
  labs(
    title = "UMAP of MOFA Factors",
    subtitle = "Colored by Mutation Type",
    color = "Mutation"
  ) +
  theme_bw(base_size = 12) +
  # Updated theme to remove grid lines
  theme(
    legend.position = "right",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

# -- 4. Display the Plots ----
print(umap_plot_background)
print(umap_plot_mutation)

```

Gene Set Enrichment Analysis of MOFA Factors using gprofiler2
```{r}
# --- Define Parameters ---
TARGET_FACTOR <- 1
SOURCE_GENESET <- "REAC"
FDR_THRESHOLD <- 0.05
TOP_N_TERMS_PLOT <- 15
Y_AXIS_LABEL_WIDTH <- 50

# Define the views and their corresponding feature suffixes to be removed
VIEWS_TO_PROCESS <- c("RNA" = "_RNA$", "protein" = "_protein$", "Ubiq" = "_ubiq$")


# Helper function to run GSEA for a given set of weights
run_gsea <- function(weights, suffix_to_remove) {
  # Clean feature names by removing view-specific suffixes
  features <- unique(gsub(suffix_to_remove, "", weights$feature))

  if (length(features) < 3) return(NULL)

  # Perform enrichment analysis using gprofiler2
  gost_result <- gost(
    query = features,
    organism = "hsapiens",
    sources = SOURCE_GENESET,
    user_threshold = FDR_THRESHOLD,
    correction_method = "fdr"
  )

  if (!is.null(gost_result)) {
    gost_result$result %>%
      group_by(term_name) %>%
      slice_min(order_by = p_value, n = 1, with_ties = FALSE) %>%
      ungroup()
  } else {
    NULL
  }
}

# Get all weights for the target factor
all_weights <- get_weights(
  MOFA_out,
  factors = TARGET_FACTOR,
  as.data.frame = TRUE
)

# Run GSEA for positive and negative weights for each view
gsea_results <- lapply(names(VIEWS_TO_PROCESS), function(view_name) {
  view_suffix <- VIEWS_TO_PROCESS[[view_name]]

  # Process positive weights
  weights_pos <- all_weights %>% filter(view == view_name & value > 0)
  results_pos <- if (nrow(weights_pos) > 0) run_gsea(weights_pos, view_suffix) else NULL

  # Process negative weights
  weights_neg <- all_weights %>% filter(view == view_name & value < 0)
  results_neg <- if (nrow(weights_neg) > 0) run_gsea(weights_neg, view_suffix) else NULL

  Sys.sleep(1) # Be courteous to the g:Profiler API
  list(positive = results_pos, negative = results_neg)
})
names(gsea_results) <- names(VIEWS_TO_PROCESS)



# Calculate a global y-axis limit for consistent scaling across all plots
all_results_df <- bind_rows(unlist(gsea_results, recursive = FALSE))
global_plot_limit <- if (nrow(all_results_df) > 0) ceiling(max(-log10(all_results_df$p_value))) else 10

# Helper function to create a standardized bar plot
create_gsea_plot <- function(results_df, view_name, weight_sign) {
  if (is.null(results_df) || nrow(results_df) == 0) return(NULL)

  plot_data <- results_df %>%
    mutate(NegLog10FDR = -log10(p_value)) %>%
    slice_max(order_by = NegLog10FDR, n = TOP_N_TERMS_PLOT) %>%
    mutate(
      term_name_truncated = str_trunc(term_name, width = Y_AXIS_LABEL_WIDTH),
      term_name_ordered = fct_reorder(term_name_truncated, NegLog10FDR)
    )

  # Swapped color palettes: Positive is now Blues, Negative is now Reds
  color_palette <- if (weight_sign == "Positive") "Blues" else "Reds"

  ggplot(plot_data, aes(x = term_name_ordered, y = NegLog10FDR, fill = NegLog10FDR)) +
    geom_bar(stat = "identity") +
    coord_flip() +
    scale_fill_distiller(palette = color_palette, direction = 1, limits = c(0, global_plot_limit)) +
    scale_y_continuous(limits = c(0, global_plot_limit), expand = c(0, 0.1)) +
    labs(
      title = paste(view_name, "View Enrichment"),
      subtitle = paste(weight_sign, "Weights | Factor", TARGET_FACTOR, "|", SOURCE_GENESET),
      x = NULL,
      y = expression(-log[10]("FDR")),
      fill = expression(-log[10]("FDR"))
    ) +
    theme_minimal(base_size = 11) +
    # Added theme elements to remove the grid
    theme(
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
    )
}

# Generate a list of all plots
plot_list <- unlist(lapply(names(gsea_results), function(view) {
  list(
    create_gsea_plot(gsea_results[[view]]$positive, view, "Positive"),
    create_gsea_plot(gsea_results[[view]]$negative, view, "Negative")
  )
}), recursive = FALSE)

# Filter out any null plots that might have been generated
plot_list <- Filter(Negate(is.null), plot_list)


# Convert plots to graphical objects (grobs) to measure their components
grob_list <- lapply(plot_list, ggplotGrob)

# Find the maximum width of the y-axis labels across all plots
max_width <- do.call(grid::unit.pmax, lapply(grob_list, function(g) g$widths[g$layout$l[grepl("axis-l", g$layout$name)]]))

# Apply the maximum width to all plots to standardize them
grob_list_aligned <- lapply(grob_list, function(g) {
  g$widths[g$layout$l[grepl("axis-l", g$layout$name)]] <- max_width
  return(g)
})

# Display each aligned plot
for (grob in grob_list_aligned) {
  grid.newpage()
  grid.draw(grob)
  Sys.sleep(1)
}
```

Gene Set Enrichment Analysis of MOFA Factors (Human Phenotype Ontology)
```{r}
# --- Define Parameters ---
TARGET_FACTOR <- 1
SOURCE_GENESET <- "HP"
FDR_THRESHOLD <- 0.05
TOP_N_TERMS_PLOT <- 15
Y_AXIS_LABEL_WIDTH <- 50


# Define the views and their corresponding feature suffixes to be removed
VIEWS_TO_PROCESS <- c("RNA" = "_RNA$", "protein" = "_protein$", "Ubiq" = "_ubiq$")


# Helper function to run GSEA for a given set of weights
run_gsea <- function(weights, suffix_to_remove) {
  # Clean feature names by removing view-specific suffixes
  features <- unique(gsub(suffix_to_remove, "", weights$feature))

  if (length(features) < 3) return(NULL)

  # Perform enrichment analysis using gprofiler2
  gost_result <- gost(
    query = features,
    organism = "hsapiens",
    sources = SOURCE_GENESET,
    user_threshold = FDR_THRESHOLD,
    correction_method = "fdr"
  )

  # Return significant results, aggregating by term name to keep the most significant
  if (!is.null(gost_result)) {
    gost_result$result %>%
      group_by(term_name) %>%
      slice_min(order_by = p_value, n = 1, with_ties = FALSE) %>%
      ungroup()
  } else {
    NULL
  }
}

all_weights <- get_weights(
  MOFA_out,
  factors = TARGET_FACTOR,
  as.data.frame = TRUE
)

# Run GSEA for positive and negative weights for each view
gsea_results <- lapply(names(VIEWS_TO_PROCESS), function(view_name) {
  view_suffix <- VIEWS_TO_PROCESS[[view_name]]

  # Process positive weights
  weights_pos <- all_weights %>% filter(view == view_name & value > 0)
  results_pos <- if (nrow(weights_pos) > 0) run_gsea(weights_pos, view_suffix) else NULL

  # Process negative weights
  weights_neg <- all_weights %>% filter(view == view_name & value < 0)
  results_neg <- if (nrow(weights_neg) > 0) run_gsea(weights_neg, view_suffix) else NULL

  Sys.sleep(1) # Be courteous to the g:Profiler API
  list(positive = results_pos, negative = results_neg)
})
names(gsea_results) <- names(VIEWS_TO_PROCESS)


# Calculate a global y-axis limit for consistent scaling across all plots
all_results_df <- bind_rows(unlist(gsea_results, recursive = FALSE))
global_plot_limit <- if (nrow(all_results_df) > 0) ceiling(max(-log10(all_results_df$p_value))) else 10

# Helper function to create a standardized bar plot
create_gsea_plot <- function(results_df, view_name, weight_sign) {
  if (is.null(results_df) || nrow(results_df) == 0) return(NULL)

  plot_data <- results_df %>%
    mutate(NegLog10FDR = -log10(p_value)) %>%
    slice_max(order_by = NegLog10FDR, n = TOP_N_TERMS_PLOT) %>%
    mutate(
      term_name_truncated = str_trunc(term_name, width = Y_AXIS_LABEL_WIDTH),
      term_name_ordered = fct_reorder(term_name_truncated, NegLog10FDR)
    )

  # Swapped color palettes: Positive is now Blues, Negative is now Reds
  color_palette <- if (weight_sign == "Positive") "Blues" else "Reds"

  ggplot(plot_data, aes(x = term_name_ordered, y = NegLog10FDR, fill = NegLog10FDR)) +
    geom_bar(stat = "identity") +
    coord_flip() +
    scale_fill_distiller(palette = color_palette, direction = 1, limits = c(0, global_plot_limit)) +
    scale_y_continuous(limits = c(0, global_plot_limit), expand = c(0, 0.1)) +
    labs(
      title = paste(view_name, "View Enrichment"),
      subtitle = paste(weight_sign, "Weights | Factor", TARGET_FACTOR, "|", SOURCE_GENESET),
      x = NULL,
      y = expression(-log[10]("FDR")),
      fill = expression(-log[10]("FDR"))
    ) +
    theme_minimal(base_size = 11) +
    # Added theme elements to remove the grid
    theme(
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
    )
}

# Generate a list of all plots
plot_list <- unlist(lapply(names(gsea_results), function(view) {
  list(
    create_gsea_plot(gsea_results[[view]]$positive, view, "Positive"),
    create_gsea_plot(gsea_results[[view]]$negative, view, "Negative")
  )
}), recursive = FALSE)

plot_list <- Filter(Negate(is.null), plot_list)

# Convert plots to graphical objects (grobs) to measure their components
grob_list <- lapply(plot_list, ggplotGrob)

# Find the maximum width of the y-axis labels across all plots
max_width <- do.call(grid::unit.pmax, lapply(grob_list, function(g) g$widths[g$layout$l[grepl("axis-l", g$layout$name)]]))

# Apply the maximum width to all plots to standardize them
grob_list_aligned <- lapply(grob_list, function(g) {
  g$widths[g$layout$l[grepl("axis-l", g$layout$name)]] <- max_width
  return(g)
})

# Display each aligned plot
for (grob in grob_list_aligned) {
  grid.newpage()
  grid.draw(grob)
  Sys.sleep(1)
}

```



Export REACTOME results from Factor 1
```{r}
# ===================================================================
# EXPORT MOFA GSEA RESULTS TO XLSX
# ===================================================================

# Ensure the 'openxlsx' library is loaded
# install.packages("openxlsx") # Run this line if you don't have it installed
library(openxlsx)

# --- 1. Create a list to hold all the data frames for export ---
list_for_export <- list()

# --- 2. Loop through the nested gsea_results list ---
# The outer loop goes through the views ('RNA', 'protein', 'Ubiq')
for (view_name in names(gsea_results)) {
  
  # The inner loop goes through the weight signs ('positive', 'negative')
  for (sign_name in names(gsea_results[[view_name]])) {
    
    # Create a descriptive name for the Excel sheet
    # e.g., "RNA_Positive_Weights"
    sheet_name <- paste(view_name, sign_name, "Weights", sep = "_")
    
    # Extract the results data frame for the current combination
    results_df <- gsea_results[[view_name]][[sign_name]]
    
    # Check if the result is NULL or empty. If so, create a placeholder.
    if (is.null(results_df) || nrow(results_df) == 0) {
      list_for_export[[sheet_name]] <- data.frame(
        Message = paste("No significant", SOURCE_GENESET, "terms found.")
      )
    } else {
      # Otherwise, add the full results table to our list
      list_for_export[[sheet_name]] <- as.data.frame(results_df)
    }
  }
}

# --- 3. Write the list of data frames to a single .xlsx file ---
# Each element in the list will become a separate worksheet.
write.xlsx(
  list_for_export,
  file = paste0("MOFA_GSEA_Factor", TARGET_FACTOR, "_Results.xlsx"),
  rowNames = FALSE,
  overwrite = TRUE
)

# --- 4. Print a confirmation message ---
message(paste0(
  "Successfully exported MOFA GSEA results for Factor ", TARGET_FACTOR, 
  " to 'MOFA_GSEA_Factor", TARGET_FACTOR, "_Results.xlsx'"
))
```



Export HPO results from Factor 1
```{r}
# ===================================================================
# EXPORT MOFA GSEA (HP) RESULTS TO XLSX
# ===================================================================

# Ensure the 'openxlsx' library is loaded
# install.packages("openxlsx") # Run this line if you don't have it installed
library(openxlsx)

# --- 1. Create a list to hold all the data frames for export ---
list_for_export_hp <- list()

# --- 2. Loop through the nested gsea_results list ---
# The outer loop goes through the views ('RNA', 'protein', 'Ubiq')
for (view_name in names(gsea_results)) {
  
  # The inner loop goes through the weight signs ('positive', 'negative')
  for (sign_name in names(gsea_results[[view_name]])) {
    
    # Create a descriptive name for the Excel sheet
    # e.g., "RNA_Positive_Weights"
    sheet_name <- paste(view_name, sign_name, "Weights", sep = "_")
    
    # Extract the results data frame for the current combination
    results_df <- gsea_results[[view_name]][[sign_name]]
    
    # Check if the result is NULL or empty. If so, create a placeholder.
    if (is.null(results_df) || nrow(results_df) == 0) {
      list_for_export_hp[[sheet_name]] <- data.frame(
        Message = paste("No significant", SOURCE_GENESET, "terms found.")
      )
    } else {
      # Otherwise, add the full results table to our list
      list_for_export_hp[[sheet_name]] <- as.data.frame(results_df)
    }
  }
}

# --- 3. Write the list of data frames to a single .xlsx file ---
# The file name dynamically includes the TARGET_FACTOR and SOURCE_GENESET
output_filename <- paste0("MOFA_GSEA_", SOURCE_GENESET, "_Factor", TARGET_FACTOR, "_Results.xlsx")

write.xlsx(
  list_for_export_hp,
  file = output_filename,
  rowNames = FALSE,
  overwrite = TRUE
)

# --- 4. Print a confirmation message ---
message(paste("Successfully exported MOFA GSEA (HP) results to:", output_filename))

```

Gene Set Enrichment Analysis of MOFA Factors using gprofiler2
Combination of Factor 1 +3
```{r}
# --- Define Parameters ---
TARGET_FACTORS <- c(1, 3)     # The two factors to combine
SOURCE_GENESET <- "REAC"
FDR_THRESHOLD <- 0.05
TOP_N_TERMS_PLOT <- 15
Y_AXIS_LABEL_WIDTH <- 50

# NEW: Since we removed n=100, we must filter out tiny noise weights 
# to avoid the "Everything is a hit" error.
WEIGHT_THRESHOLD <- 0.5       

# Define the views and their corresponding feature suffixes to be removed
VIEWS_TO_PROCESS <- c("RNA" = "_RNA$", "protein" = "_protein$", "Ubiq" = "_ubiq$")

# --- 1. Data Processing: Calculate Composite Score (F1 + F3) ---

# Get weights for both factors
raw_weights <- get_weights(
  MOFA_out,
  factors = TARGET_FACTORS,
  as.data.frame = TRUE
)

# Pivot and Sum: Calculate the mathematical combination
composite_weights <- raw_weights %>%
  select(feature, view, factor, value) %>%
  pivot_wider(names_from = factor, values_from = value, names_prefix = "Factor") %>%
  # Dynamically calculate sum of the columns named FactorX and FactorY
  mutate(composite_score = rowSums(select(., starts_with("Factor"))))

# --- 2. Helper Functions ---

# Helper function to run GSEA
run_gsea <- function(weights_df, suffix_to_remove) {
  # Clean feature names
  features <- unique(gsub(suffix_to_remove, "", weights_df$feature))
  
  if (length(features) < 3) return(NULL)
  
  # Perform enrichment analysis
  gost_result <- gost(
    query = features,
    organism = "hsapiens",
    sources = SOURCE_GENESET,
    user_threshold = FDR_THRESHOLD,
    correction_method = "fdr"
  )
  
  if (!is.null(gost_result)) {
    gost_result$result %>%
      group_by(term_name) %>%
      slice_min(order_by = p_value, n = 1, with_ties = FALSE) %>%
      ungroup()
  } else {
    NULL
  }
}

# Helper function to create standardized bar plot
create_gsea_plot <- function(results_df, view_name, weight_sign) {
  if (is.null(results_df) || nrow(results_df) == 0) return(NULL)
  
  # Format title based on the combination
  combo_title <- paste("Factor", paste(TARGET_FACTORS, collapse = " + "))
  
  plot_data <- results_df %>%
    mutate(NegLog10FDR = -log10(p_value)) %>%
    slice_max(order_by = NegLog10FDR, n = TOP_N_TERMS_PLOT) %>%
    mutate(
      term_name_truncated = str_trunc(term_name, width = Y_AXIS_LABEL_WIDTH),
      term_name_ordered = fct_reorder(term_name_truncated, NegLog10FDR)
    )
  
  color_palette <- if (weight_sign == "Positive") "Blues" else "Reds"
  
  ggplot(plot_data, aes(x = term_name_ordered, y = NegLog10FDR, fill = NegLog10FDR)) +
    geom_bar(stat = "identity") +
    coord_flip() +
    scale_fill_distiller(palette = color_palette, direction = 1, limits = c(0, global_plot_limit)) +
    scale_y_continuous(limits = c(0, global_plot_limit), expand = c(0, 0.1)) +
    labs(
      title = paste(view_name, "Enrichment"),
      subtitle = paste(weight_sign, "Drivers of", combo_title),
      x = NULL,
      y = expression(-log[10]("FDR")),
      fill = expression(-log[10]("FDR"))
    ) +
    theme_minimal(base_size = 11) +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
    )
}

# --- 3. Run GSEA on Composite Scores ---

gsea_results <- lapply(names(VIEWS_TO_PROCESS), function(view_name) {
  view_suffix <- VIEWS_TO_PROCESS[[view_name]]
  
  # Filter weights for this view
  view_data <- composite_weights %>% filter(view == view_name)
  
  # Process Positive Composite Scores (Using Threshold instead of Top N)
  pos_drivers <- view_data %>% filter(composite_score > WEIGHT_THRESHOLD)
  results_pos <- if (nrow(pos_drivers) > 0) run_gsea(pos_drivers, view_suffix) else NULL
  
  # Process Negative Composite Scores (Using Threshold instead of Top N)
  neg_drivers <- view_data %>% filter(composite_score < -WEIGHT_THRESHOLD)
  results_neg <- if (nrow(neg_drivers) > 0) run_gsea(neg_drivers, view_suffix) else NULL
  
  Sys.sleep(1) 
  list(positive = results_pos, negative = results_neg)
})
names(gsea_results) <- names(VIEWS_TO_PROCESS)


# --- 4. Plotting and Alignment (Unchanged from your snippet) ---

# Calculate global limit
all_results_df <- bind_rows(unlist(gsea_results, recursive = FALSE))
global_plot_limit <- if (nrow(all_results_df) > 0) ceiling(max(-log10(all_results_df$p_value))) else 10

# Generate plot list
plot_list <- unlist(lapply(names(gsea_results), function(view) {
  list(
    create_gsea_plot(gsea_results[[view]]$positive, view, "Positive"),
    create_gsea_plot(gsea_results[[view]]$negative, view, "Negative")
  )
}), recursive = FALSE)

# Filter nulls
plot_list <- Filter(Negate(is.null), plot_list)

if (length(plot_list) > 0) {
  # Convert to grobs
  grob_list <- lapply(plot_list, ggplotGrob)
  
  # Align widths
  max_width <- do.call(grid::unit.pmax, lapply(grob_list, function(g) g$widths[g$layout$l[grepl("axis-l", g$layout$name)]]))
  grob_list_aligned <- lapply(grob_list, function(g) {
    g$widths[g$layout$l[grepl("axis-l", g$layout$name)]] <- max_width
    return(g)
  })
  
  # Draw plots
  for (grob in grob_list_aligned) {
    grid.newpage()
    grid.draw(grob)
    Sys.sleep(0.5)
  }
} else {
  message("No significant enrichment found. Try lowering WEIGHT_THRESHOLD.")
}
```



```{r}
library(MOFA2)
library(msigdbr)
library(data.table)
library(tidyverse)

# --- Define Parameters ---
TARGET_FACTOR <- 1
TARGET_VIEW <- "RNA"   # Change this if you want to test 'protein' or 'Ubiq'

# Helper function to convert data.frame to matrix (replaces 'matrix.please')
as_binary_matrix <- function(df) {
  mat <- as.matrix(df[, -1])
  rownames(mat) <- df[[1]]
  return(mat)
}

# 1. Fetch Gene Sets (GO:BP from MSigDB)
# We use data.table for fast reshaping as done in the tutorial
message("Fetching MSigDB data...")
genesets_df <- msigdbr(species = "Homo sapiens", category = "C5", subcategory = "BP") %>% 
  as.data.table()

# 2. Create Binary Matrix (1 = gene in pathway, 0 = not in pathway)
message("Creating binary gene set matrix...")
msgidb.matrix <- genesets_df[, id := 1] %>%
  dcast(gs_name ~ gene_symbol, value.var = "id", fill = 0) %>%
  as_binary_matrix()

message("Matrix created. Dimensions: ", paste(dim(msgidb.matrix), collapse = " x "))
```

```{r}
# --- Run Enrichment for Positive Weights ---
message("Running GSEA on POSITIVE weights for Factor ", TARGET_FACTOR, "...")
gsea.positive <- run_enrichment(
  MOFA_out,
  feature.sets = msgidb.matrix,
  view = TARGET_VIEW,
  factor = TARGET_FACTOR,
  sign = "positive"
)

# --- Run Enrichment for Negative Weights ---
message("Running GSEA on NEGATIVE weights for Factor ", TARGET_FACTOR, "...")
gsea.negative <- run_enrichment(
  MOFA_out,
  feature.sets = msgidb.matrix,
  view = TARGET_VIEW,
  factor = TARGET_FACTOR,
  sign = "negative"
)
```



```{r}
# --- Plot 1: Top Enriched Pathways (Positive Weights) ---
plot_enrichment(
  gsea.positive, 
  factor = TARGET_FACTOR, 
  max.pathways = 15
) + ggtitle(paste("Factor", TARGET_FACTOR, "Positive Enrichment (GO:BP)"))

# --- Plot 2: Top Enriched Pathways (Negative Weights) ---
plot_enrichment(
  gsea.negative, 
  factor = TARGET_FACTOR, 
  max.pathways = 15
) + ggtitle(paste("Factor", TARGET_FACTOR, "Negative Enrichment (GO:BP)"))

# --- Plot 3: Detailed GSEA Plot (Gene Weights inside Pathways) ---
# This shows the specific genes driving the top pathways
plot_enrichment_detailed(
  gsea.positive,
  factor = TARGET_FACTOR,
  max.genes = 10,
  max.pathways = 5
) + ggtitle(paste("Detailed Drivers: Factor", TARGET_FACTOR, "(Positive)"))

plot_enrichment_detailed(
  gsea.negative,
  factor = TARGET_FACTOR,
  max.genes = 10,
  max.pathways = 5
) + ggtitle(paste("Detailed Drivers: Factor", TARGET_FACTOR, "(Negative)"))
```
```{r}
library(MOFA2)
library(msigdbr)
library(data.table)
library(tidyverse)
library(gridExtra)

# --- 1. Define Parameters ---
TARGET_FACTOR <- 1

# Map View Names in MOFA to the Suffix they use in feature names
# These are appended to the gene symbols so they match the MOFA model
VIEW_SUFFIXES <- c(
  "RNA"     = "_RNA",
  "protein" = "_protein",
  "Ubiq"    = "_ubiq"
)

# --- 2. Prepare Base Gene Set Matrix (Reactome) ---
message("Fetching MSigDB data (Reactome)...")

# Fetch Reactome (C2:CP:REACTOME)
genesets_df <- msigdbr(species = "Homo sapiens", collection = "C2", subcollection = "CP:REACTOME") %>% 
  as.data.table()

message("Creating base binary gene set matrix...")
# Reshape to Wide Format (Rows = Pathways, Cols = Gene Symbols)
# fun.aggregate = length handles duplicates (genes appearing twice in a pathway)
wide_dt <- genesets_df[, id := 1] %>%
  dcast(gs_name ~ gene_symbol, value.var = "id", fill = 0, fun.aggregate = length)

# Convert to Matrix
# Exclude the first column (pathway names) from the data values
base_matrix <- as.matrix(wide_dt[, -1])
rownames(base_matrix) <- wide_dt$gs_name

# Ensure strictly binary and numeric
# (If a gene was counted twice, set it back to 1)
base_matrix[base_matrix > 1] <- 1
mode(base_matrix) <- "numeric"

message("Base Matrix Ready: ", paste(dim(base_matrix), collapse = " x "))


# --- 3. Main Loop: Run GSEA for Each View ---

# Helper function to plot safely without crashing
safe_plot_enrichment <- function(gsea_result, title_text) {
  tryCatch({
    # Try plotting with default alpha (usually 0.05 or 0.1)
    p <- plot_enrichment(gsea_result, factor = TARGET_FACTOR, max.pathways = 10) +
      ggtitle(title_text) +
      theme(plot.title = element_text(size=12, face="bold"))
    return(p)
  }, error = function(e) {
    # If it fails (usually "No significant pathways"), print a message instead of crashing
    message(paste("      [INFO] No significant pathways found for:", title_text))
    return(NULL)
  })
}

all_gsea_results <- list()

for (view_name in names(VIEW_SUFFIXES)) {
  
  suffix <- VIEW_SUFFIXES[[view_name]]
  message(paste0("\n=== Processing View: ", view_name, " (Suffix: ", suffix, ") ==="))
  
  # A. Feature Alignment
  # --------------------
  # Get the actual feature names from the MOFA model for this view
  mofa_features <- features_names(MOFA_out)[[view_name]]
  
  # Create a view-specific copy of the matrix
  view_matrix <- base_matrix
  
  # Append the suffix to the gene symbols (e.g., TP53 -> TP53_RNA)
  colnames(view_matrix) <- paste0(colnames(view_matrix), suffix)
  
  # Check overlap
  common_features <- intersect(colnames(view_matrix), mofa_features)
  message(sprintf("   Matched %d features between Reactome and MOFA model.", length(common_features)))
  
  # Skip if overlap is too small (avoids empty GSEA errors)
  if (length(common_features) < 10) {
    warning(paste("   Skipping", view_name, "- too few matching features. Check your suffix!"))
    next
  }
  
  # Subset matrix to only matching features
  view_matrix <- view_matrix[, common_features, drop = FALSE]
  
  
  # B. Run Enrichment
  # -----------------
  message("   Running GSEA (Positive weights)...")
  gsea_pos <- run_enrichment(
    MOFA_out, feature.sets = view_matrix, 
    view = view_name, factor = TARGET_FACTOR, sign = "positive"
  )
  
  message("   Running GSEA (Negative weights)...")
  gsea_neg <- run_enrichment(
    MOFA_out, feature.sets = view_matrix, 
    view = view_name, factor = TARGET_FACTOR, sign = "negative"
  )
  
  # Store results
  all_gsea_results[[view_name]] <- list(positive = gsea_pos, negative = gsea_neg)
  
  
  # C. Generate & Print Plots (Safely)
  # ----------------------------------
  message("   Generating plots...")
  
  # Plot Positive
  p1 <- safe_plot_enrichment(gsea_pos, paste0(view_name, ": Positive (Factor ", TARGET_FACTOR, ")"))
  if(!is.null(p1)) print(p1)
  
  # Plot Negative
  p2 <- safe_plot_enrichment(gsea_neg, paste0(view_name, ": Negative (Factor ", TARGET_FACTOR, ")"))
  if(!is.null(p2)) print(p2)
}

message("\n=== All Views Processed ===")
```

