---
title: "Untitled"
author: "Michael Eibl"
date: "`r Sys.Date()`"
output: html_document
---


```{r}
library("DESeq2")
library("stringr")
library("ggplot2")
library("ggpubr")
library("RNAseqQC")
library("dplyr")
library("patchwork")
library("org.Hs.eg.db")
library("annotate")
library("AnnotationHub")
library('EnhancedVolcano')
library("gridExtra")
library("grid")
library("clusterProfiler")
library("eulerr")
library("ComplexHeatmap")
library("openxlsx")
library("Homo.sapiens")
library("dplyr")
library("limma")
library("DEP2")
library("patchwork")
```

Function for color scheme
```{r}
gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}
```

Loading the file from Azmal
```{r}
abundance <- read.xlsx(xlsxFile = "./Ubiquitination_IP_Azmal.xlsx")

abundance <- abundance[which(!duplicated(abundance$PG.Genes)), ]
abundance <- abundance[which(!is.na(abundance$PG.Genes)),]

rownames(abundance) <- abundance$PG.Genes

colnames(abundance)

colnames(abundance)[c(9,11)] <- c("IP_G_HOM_rep1", "IP_G_HOM_rep3")
colnames(abundance) <- gsub("IP_", "", colnames(abundance))
colnames(abundance) <- gsub("Del", "DelC", colnames(abundance))
colnames(abundance) <- gsub("DUPC", "DupC", colnames(abundance))
colnames(abundance) <- gsub("Het", "HetDel", colnames(abundance))
colnames(abundance) <- gsub("HOM", "HomoDel", colnames(abundance))
colnames(abundance) <- gsub("PWS1", "PWSTI", colnames(abundance))
colnames(abundance) <- gsub("PWS2", "PWSTII", colnames(abundance))
colnames(abundance) <- gsub("rep", "", colnames(abundance))
colnames(abundance) <- gsub("^M_", "MG_", colnames(abundance))
colnames(abundance) <- gsub("^G_", "GM_", colnames(abundance))

proteins_all <- rownames(abundance)


#make columns with protein abundance numeric (default is character)
abundance[,3:ncol(abundance)] <- sapply(abundance[,3:ncol(abundance)], as.numeric)

#downstream NaNs are not allowed and should be replaced by NA
abundance <- abundance %>% as.data.frame(.) %>% mutate_all(~ifelse(is.nan(.), NA, .))

rownames(abundance) <- proteins_all

#remove rows containing only NAs
abundance <- abundance[!!rowSums(!is.na(abundance[,3:ncol(abundance)])),]

#no Inf values, replaced by NA
abundance <- do.call(data.frame,lapply(abundance, function(x) replace(x, is.infinite(x),NA)))
```

Pre-Processing
```{r}
# Generate a SummarizedExperiment object by parsing condition information from the column names
experimental_design <- data.frame(row.names = colnames(abundance[,3:ncol(abundance)]),
                                  genotype = rep(NA, ncol(abundance[,3:ncol(abundance)])),
                                  background = rep(NA, ncol(abundance[,3:ncol(abundance)])),
                                  replicate = rep(NA, ncol(abundance[,3:ncol(abundance)])))



for (i in 1:ncol(abundance[,3:ncol(abundance)])) {
  experimental_design[i, "genotype"] <- str_split(colnames(abundance[,3:ncol(abundance)])[i], pattern = "_")[[1]][2]
  experimental_design[i, "background"] <- str_split(colnames(abundance[,3:ncol(abundance)])[i], pattern = "_")[[1]][1]
  experimental_design[i, "replicate"] <- str_split(colnames(abundance[,3:ncol(abundance)])[i], pattern = "_")[[1]][3]
}

experimental_design$ID <- 1:nrow(experimental_design)
experimental_design$name <- rownames(experimental_design)

colnames(experimental_design)[c(1,5)] <- c("condition", "label")

experimental_design$ID <- factor(experimental_design$ID)
experimental_design$condition <- factor(experimental_design$condition)

saveRDS(experimental_design, file = "UbiEAs_sample_info.rds")

abundance_unique_8330 <- DEP2::make_unique(proteins = abundance[,c(1,2,which(experimental_design$background == "GM")+2)],
                                     names = "PG.Genes",
                                     ids = "PG.ProteinGroups",
                                     delim = ",")

se_8330 <- DEP2::make_se(proteins_unique = abundance_unique_8330,
                   columns = 3:20,
                   expdesign = experimental_design[which(experimental_design$background == "GM"),])


abundance_unique_MGH <- DEP2::make_unique(proteins = abundance[,c(1,2,which(experimental_design$background == "MG")+2)],
                                    names = "PG.Genes",
                                    ids = "PG.ProteinGroups",
                                    delim = ",")

se_MGH <- DEP2::make_se(proteins_unique = abundance_unique_MGH,
                  columns = 3:23,
                  expdesign = experimental_design[which(experimental_design$background == "MG"),])
```

Quality control - Proteins per Sample
```{r}
#8330
p1 <- plot_frequency(se_8330)

#we first filter out proteins that contain too many missing values
#Filter for proteins that are identified in all replicates of at least one condition
data_filt_8330_Ubi <- filter_se(se_8330, thr = 0)

p2 <- plot_numbers(data_filt_8330_Ubi)

plot_coverage(data_filt_8330_Ubi)

#MGH
p3 <- plot_frequency(se_MGH)

#we first filter out proteins that contain too many missing values
#Filter for proteins that are identified in all replicates of at least one condition
data_filt_MGH_Ubi <- filter_se(se_MGH, thr = 0)

p4 <- plot_numbers(data_filt_MGH_Ubi)

plot_coverage(data_filt_MGH_Ubi)

(p2 + p4) / (p1 + p3)
```


Data normalization
```{r}
# Normalize the data
data_norm_8330_Ubi <- normalize_vsn(data_filt_8330_Ubi)

# Visualize normalization by boxplots for all samples before and after normalization
p1 <- plot_normalization(data_filt_8330_Ubi, data_norm_8330_Ubi)

# Normalize the data
data_norm_MGH_Ubi <- normalize_vsn(data_filt_MGH_Ubi)

# Visualize normalization by boxplots for all samples before and after normalization
p2 <- plot_normalization(data_filt_MGH_Ubi, data_norm_MGH_Ubi)

p1 + p2
```
Heatmap with missing values
```{r}
# Plot a heatmap of proteins with missing values
plot_missval(data_filt_8330_Ubi)
plot_missval(data_filt_MGH_Ubi)

# Plot intensity distributions and cumulative fraction of proteins with and without missing values
plot_detect(data_filt_8330_Ubi)
plot_detect(data_filt_MGH_Ubi)

```
Missing values - Part 2
```{r}
#Indeed the proteins with missing values have on average low intensities. This data (MNAR and close to the detection limit) should be imputed by a left-censored imputation method, such as the quantile regression-based left-censored function (“QRILC”) or random draws from a left-shifted distribution (“MinProb” and “man”)

# Impute missing data using random draws from a Gaussian distribution centered around a minimal value (for MNAR)
Ubi_data_imp_8330 <- DEP2::impute(data_norm_8330_Ubi, fun = "MinProb", q = 0.01)

# Plot intensity distributions before and after imputation
plot_imputation(data_norm_8330_Ubi, Ubi_data_imp_8330)


###MGH

# Impute missing data using random draws from a Gaussian distribution centered around a minimal value (for MNAR)
Ubi_data_imp_MGH <- DEP2::impute(data_norm_MGH_Ubi, fun = "MinProb", q = 0.01)

# Plot intensity distributions before and after imputation
plot_imputation(data_norm_MGH_Ubi, Ubi_data_imp_MGH)

```

Differential enrichment analysis
```{r}
# Differential enrichment analysis  based on linear models and empherical Bayes statistics

# Test every sample versus control
data_diff_8330 <- DEP2::test_diff(Ubi_data_imp_8330, type = "all", fdr.type = "BH")

deu_8330 <- add_rejections(data_diff_8330, alpha = 0.1, lfc = 0)

data_diff_MGH <- DEP2::test_diff(Ubi_data_imp_MGH, type = "all", fdr.type = "BH")

deu_MGH <- add_rejections(data_diff_MGH, alpha = 0.1, lfc = 0)

saveRDS(Ubi_data_imp_8330, file = "Ubi_Ubi_data_imp_8330_BH_new.rds")
saveRDS(deu_8330, file = "Ubi_deu_8330_BH_new.rds")
saveRDS(Ubi_data_imp_MGH, file = "Ubi_Ubi_data_imp_MGH_BH_new.rds")
saveRDS(deu_MGH, file = "Ubi_deu_MGH_BH_new.rds")
```

Principal component analysis
```{r}
plot_pca(deu_8330, x = 1, y = 2, n=5000, point_size = 4, , indicate = "condition") +
    theme_pubr() + 
  theme(legend.position = "right") +
  ggtitle("")
plot_pca(deu_MGH, x = 1, y = 2, n=5000, point_size = 4, indicate = "condition") +
  theme_pubr() + 
  theme(legend.position = "right") +
  ggtitle("")

# Plot the Pearson correlation matrix
plot_cor(deu_8330, significant = TRUE, lower = 0, upper = 1, pal = "Reds")
plot_cor(deu_MGH, significant = TRUE, lower = 0, upper = 1, pal = "Reds")
```

Heatmap of PCA

```{r}
plot_heatmap(deu_MGH, type = "centered", col_limit = 3, kmeans = T, k =6)

plot_heatmap(deu_8330, type = "centered", col_limit = 3, kmeans = T, k =6)
```

Volcano plots
```{r}
(plot_volcano(deu_8330, contrast = "DelC_vs_WT", label_size = 3, add_names = TRUE, adjusted = T) +
  plot_volcano(deu_8330, contrast = "DupC_vs_WT", label_size = 3, add_names = TRUE, adjusted = T) +
  plot_volcano(deu_8330, contrast = "HomoDel_vs_WT", label_size = 3, add_names = TRUE, adjusted = T)) /
  (plot_volcano(deu_8330, contrast = "PWSTI_vs_WT", label_size = 3, add_names = TRUE, adjusted = T) +
     plot_volcano(deu_8330, contrast = "PWSTII_vs_WT", label_size = 3, add_names = TRUE, adjusted = T))


(plot_volcano(deu_MGH, contrast = "DelC_vs_WT", label_size = 3, add_names = TRUE, adjusted = T) +
  plot_volcano(deu_MGH, contrast = "DupC_vs_WT", label_size = 3, add_names = TRUE, adjusted = T) +
  plot_volcano(deu_MGH, contrast = "HetDel_vs_WT", label_size = 3, add_names = TRUE, adjusted = T)) /
  (plot_volcano(deu_MGH, contrast = "HomoDel_vs_WT", label_size = 3, add_names = TRUE, adjusted = T) +
     plot_volcano(deu_MGH, contrast = "PWSTI_vs_WT", label_size = 3, add_names = TRUE, adjusted = T) +
     plot_volcano(deu_MGH, contrast = "PWSTII_vs_WT", label_size = 3, add_names = TRUE, adjusted = T))
```

Plotting function 
```{r}
plotAbundance <- function(data, GOI) {
  ggboxplot(
    data.frame(
      expression = assay(data)[GOI,],
      genotype = data$condition
    ),
  x="genotype",
  y="expression",
  fill="genotype",
  add="jitter",
  add.params = list(size=3),
  title = GOI
  ) +
    ylab("normalized & imputed log2 intensity") +
    xlab("") +
    theme(legend.position = "none") +
    theme(text = element_text(size=14))
}
```

Testing plotting function
```{r}


plotAbundance(data = Ubi_data_imp_8330, GOI = "CALM1;CALM2;CALM3")
plotAbundance(data = Ubi_data_imp_MGH, GOI = "CALM1;CALM2;CALM3")

```


Comparison mutants vs WT
```{r}
deu_8330 <- rowData(deu_8330)

comparisons_deu <- c("DelC_vs_WT", "DupC_vs_WT", "HomoDel_vs_WT", "PWSTI_vs_WT", "PWSTII_vs_WT")
list_8330_new_Ubi_up <- list()
for (comp in comparisons_deu) {
  
  # Extract significant results
  signif_genes <- deu_8330$name[which(deu_8330[,paste0(comp, "_p.adj")] < 0.1 &
                                               deu_8330[,paste0(comp, "_diff")] > 0)]
  
  list_8330_new_Ubi_up[[comp]] <- signif_genes
}

list_8330_new_Ubi_down <- list()
for (comp in comparisons_deu) {
  
  # Extract significant results
  signif_genes <- deu_8330$name[which(deu_8330[,paste0(comp, "_p.adj")] < 0.1 &
                                               deu_8330[,paste0(comp, "_diff")] < 0)]
  
  list_8330_new_Ubi_down[[comp]] <- signif_genes
}

names(list_8330_new_Ubi_up) <- c("Protein_DelC_up", "Protein_DupC_up", "Protein_HomoDel_up", "Protein_PWSTI_up", "Protein_PWSTII_up")
names(list_8330_new_Ubi_down) <- c("Protein_DelC_down", "Protein_DupC_down", "Protein_HomoDel_down", "Protein_PWSTI_down", "Protein_PWSTII_down")




deu_MGH <- rowData(deu_MGH)
comparisons_deu <- c("DelC_vs_WT", "DupC_vs_WT", "HetDel_vs_WT", "HomoDel_vs_WT", "PWSTI_vs_WT", "PWSTII_vs_WT")
list_MGH_new_Ubi_up <- list()
for (comp in comparisons_deu) {
  
  # Extract significant results
  signif_genes <- deu_MGH$name[which(deu_MGH[,paste0(comp, "_p.adj")] < 0.1 &
                                               deu_MGH[,paste0(comp, "_diff")] > 0)]
  
  list_MGH_new_Ubi_up[[comp]] <- signif_genes
}

list_MGH_new_Ubi_down <- list()
for (comp in comparisons_deu) {
  
  # Extract significant results
  signif_genes <- deu_MGH$name[which(deu_MGH[,paste0(comp, "_p.adj")] < 0.1 &
                                               deu_MGH[,paste0(comp, "_diff")] < 0)]
  
  list_MGH_new_Ubi_down[[comp]] <- signif_genes
}

names(list_MGH_new_Ubi_up) <- c("Protein_DelC_up", "Protein_DupC_up", "Protein_HetDel_up", "Protein_HomoDel_up", "Protein_PWSTI_up", "Protein_PWSTII_up")
names(list_MGH_new_Ubi_down) <- c("Protein_DelC_down", "Protein_DupC_down", "Protein_HetDel_down", "Protein_HomoDel_down", "Protein_PWSTI_down", "Protein_PWSTII_down")
```


Overlapping the DEUs from both backgrounds
```{r}
# --- Upregulated Proteins ---

# DelC_up
plot(euler(list("MGH" = list_MGH_new_Ubi_up$Protein_DelC_up,
                "8330" = list_8330_new_Ubi_up$Protein_DelC_up)),
     quantities = TRUE,
     main = "Upregulated Proteins - DelC")  # Add a title

# DupC_up
plot(euler(list("MGH" = list_MGH_new_Ubi_up$Protein_DupC_up,
                "8330" = list_8330_new_Ubi_up$Protein_DupC_up)),
     quantities = TRUE,
     main = "Upregulated Proteins - DupC")

# HetDel_up
plot(euler(list("MGH" = list_MGH_new_Ubi_up$Protein_HetDel_up,
                "8330" = list_8330_new_Ubi_up$Protein_HetDel_up)),
     quantities = TRUE,
     main = "Upregulated Proteins - HetDel")

# HomoDel_up
plot(euler(list("MGH" = list_MGH_new_Ubi_up$Protein_HomoDel_up,
                "8330" = list_8330_new_Ubi_up$Protein_HomoDel_up)),
     quantities = TRUE,
     main = "Upregulated Proteins - HomoDel")

# PWSTI_up
plot(euler(list("MGH" = list_MGH_new_Ubi_up$Protein_PWSTI_up,
                "8330" = list_8330_new_Ubi_up$Protein_PWSTI_up)),
     quantities = TRUE,
     main = "Upregulated Proteins - PWSTI")

# PWSTII_up
plot(euler(list("MGH" = list_MGH_new_Ubi_up$Protein_PWSTII_up,
                "8330" = list_8330_new_Ubi_up$Protein_PWSTII_up)),
     quantities = TRUE,
     main = "Upregulated Proteins - PWSTII")


list_Ubi_up_overlap <-
  list(
    "overlap_DEU_DelC" = intersect(list_MGH_new_Ubi_up$Protein_DelC_up,
                                    list_8330_new_Ubi_up$Protein_DelC_up),
    "overlap_DEU_DupC" = intersect(list_MGH_new_Ubi_up$Protein_DupC_up,
                                    list_8330_new_Ubi_up$Protein_DupC_up),
    "overlap_DEU_HetDel" = intersect(list_MGH_new_Ubi_up$Protein_HetDel_up,
                                     list_8330_new_Ubi_up$Protein_HetDel_up),
    "overlap_DEU_HomoDel" = intersect(list_MGH_new_Ubi_up$Protein_HomoDel_up,
                                      list_8330_new_Ubi_up$Protein_HomoDel_up),
    "overlap_DEU_PWSTI" = intersect(list_MGH_new_Ubi_up$Protein_PWSTI_up,
                                    list_8330_new_Ubi_up$Protein_PWSTI_up),
    "overlap_DEU_PWSTII" = intersect(list_MGH_new_Ubi_up$Protein_PWSTII_up,
                                     list_8330_new_Ubi_up$Protein_PWSTII_up)
  )

# --- Downregulated Proteins ---

# DelC_down
plot(euler(list("MGH" = list_MGH_new_Ubi_down$Protein_DelC_down,
                "8330" = list_8330_new_Ubi_down$Protein_DelC_down)),
     quantities = TRUE,
     main = "Downregulated Proteins - DelC") # Add a title

# DupC_down
plot(euler(list("MGH" = list_MGH_new_Ubi_down$Protein_DupC_down,
                "8330" = list_8330_new_Ubi_down$Protein_DupC_down)),
     quantities = TRUE,
     main = "Downregulated Proteins - DupC")

# HetDel_down
plot(euler(list("MGH" = list_MGH_new_Ubi_down$Protein_HetDel_down,
                "8330" = list_8330_new_Ubi_down$Protein_HetDel_down)),
     quantities = TRUE,
     main = "Downregulated Proteins - HetDel")

# HomoDel_down
plot(euler(list("MGH" = list_MGH_new_Ubi_down$Protein_HomoDel_down,
                "8330" = list_8330_new_Ubi_down$Protein_HomoDel_down)),
     quantities = TRUE,
     main = "Downregulated Proteins - HomoDel")

# PWSTI_down
plot(euler(list("MGH" = list_MGH_new_Ubi_down$Protein_PWSTI_down,
                "8330" = list_8330_new_Ubi_down$Protein_PWSTI_down)),
     quantities = TRUE,
     main = "Downregulated Proteins - PWSTI")

# PWSTII_down
plot(euler(list("MGH" = list_MGH_new_Ubi_down$Protein_PWSTII_down,
                "8330" = list_8330_new_Ubi_down$Protein_PWSTII_down)),
     quantities = TRUE,
     main = "Downregulated Proteins - PWSTII")


list_Ubi_down_overlap <-
  list(
    "overlap_DEU_DelC" = intersect(list_MGH_new_Ubi_down$Protein_DelC_down,
                                    list_8330_new_Ubi_down$Protein_DelC_down),
    "overlap_DEU_DupC" = intersect(list_MGH_new_Ubi_down$Protein_DupC_down,
                                    list_8330_new_Ubi_down$Protein_DupC_down),
    "overlap_DEU_HetDel" = intersect(list_MGH_new_Ubi_down$Protein_HetDel_down,
                                     list_8330_new_Ubi_down$Protein_HetDel_down),
    "overlap_DEU_HomoDel" = intersect(list_MGH_new_Ubi_down$Protein_HomoDel_down,
                                      list_8330_new_Ubi_down$Protein_HomoDel_down),
    "overlap_DEU_PWSTI" = intersect(list_MGH_new_Ubi_down$Protein_PWSTI_down,
                                    list_8330_new_Ubi_down$Protein_PWSTI_down),
    "overlap_DEU_PWSTII" = intersect(list_MGH_new_Ubi_down$Protein_PWSTII_down,
                                     list_8330_new_Ubi_down$Protein_PWSTII_down)
  )

lapply(list_Ubi_down_overlap, length)

```



Prepare lists of DEU and significantly enriched GO-terms for the respective comparisons of overlap analysis - MGH
```{r}
deu_MGH_results_new <- readRDS(file = "Ubi_Ubi_dep_MGH_BH_new.rds")
deu_8330_results_new <- readRDS(file = "Ubi_Ubi_dep_8330_BH_new.rds")



deu_MGH_new_results <- rowData(deu_MGH_results_new)

comparisons_DEU <- c("DelC_vs_WT", "DupC_vs_WT", "HetDel_vs_WT", "HomoDel_vs_WT", "PWSTI_vs_WT", "PWSTII_vs_WT")
list_MGH_new_Ubi_up <- list()
for (comp in comparisons_DEU) {
  
  # Extract significant results
  signif_genes <- deu_MGH_new_results$name[which(deu_MGH_new_results[,paste0(comp, "_p.adj")] < 0.1 &
                                               deu_MGH_new_results[,paste0(comp, "_diff")] > 0)]
  
  list_MGH_new_Ubi_up[[comp]] <- signif_genes
}



list_MGH_new_Ubi_down <- list()
for (comp in comparisons_DEU) {
  
  # Extract significant results
  signif_genes <- deu_MGH_new_results$name[which(deu_MGH_new_results[,paste0(comp, "_p.adj")] < 0.1 &
                                               deu_MGH_new_results[,paste0(comp, "_diff")] < 0)]
  
  list_MGH_new_Ubi_down[[comp]] <- signif_genes
}

names(list_MGH_new_Ubi_up) <- c("Protein_DelC_up", "Protein_DupC_up", "Protein_HetDel_up", "Protein_HomoDel_up", "Protein_PWSTI_up", "Protein_PWSTII_up")
names(list_MGH_new_Ubi_down) <- c("Protein_DelC_down", "Protein_DupC_down", "Protein_HetDel_down", "Protein_HomoDel_down", "Protein_PWSTI_down", "Protein_PWSTII_down")
```


Prepare lists of DEU and significantly enriched GO-terms for the respective comparisons of overlap analysis - 8330
```{r}
deu_8330_new_results <- rowData(deu_8330_results_new)

comparisons_DEU <- c("DelC_vs_WT", "DupC_vs_WT", "HomoDel_vs_WT", "PWSTI_vs_WT", "PWSTII_vs_WT")

list_8330_new_Ubi_up <- list()
for (comp in comparisons_DEU) {
  
  # Extract significant results
  signif_genes <- deu_8330_new_results$name[which(deu_8330_new_results[,paste0(comp, "_p.adj")] < 0.1 &
                                               deu_8330_new_results[,paste0(comp, "_diff")] > 0)]
  
  list_8330_new_Ubi_up[[comp]] <- signif_genes
}



list_8330_new_Ubi_down <- list()
for (comp in comparisons_DEU) {
  
  # Extract significant results
  signif_genes <- deu_8330_new_results$name[which(deu_8330_new_results[,paste0(comp, "_p.adj")] < 0.1 &
                                               deu_8330_new_results[,paste0(comp, "_diff")] < 0)]
  
  list_8330_new_Ubi_down[[comp]] <- signif_genes
}

names(list_8330_new_Ubi_up) <- c("Protein_DelC_up", "Protein_DupC_up", "Protein_HomoDel_up", "Protein_PWSTI_up", "Protein_PWSTII_up")
names(list_8330_new_Ubi_down) <- c("Protein_DelC_down", "Protein_DupC_down", "Protein_HomoDel_down", "Protein_PWSTI_down", "Protein_PWSTII_down")
```




Preparing GO lists - MGH Ubiquitination
```{r}
comparisons_DEU <- c("DelC_vs_WT", "DupC_vs_WT", "HetDel_vs_WT", "HomoDel_vs_WT", "PWSTI_vs_WT", "PWSTII_vs_WT")

GO_list_MGH_Ubi_up <- list()
for (comp in comparisons_DEU) {

  # Create background dataset for hypergeometric testing using all genes tested for significance in the results
  all_genes <- rowData(deu_MGH_results_new)$name

  # Extract significant results
  signif_genes <- rowData(deu_MGH_results_new)$name[which(rowData(deu_MGH_results_new)[, paste0(comp, "_p.adj")] < 0.1 &
                                                            rowData(deu_MGH_results_new)[, paste0(comp, "_diff")] > 0)]

  # Symbol to EntrezID
  all_genes <- bitr(all_genes, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = "org.Hs.eg.db")
  signif_genes <- bitr(signif_genes, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = "org.Hs.eg.db")

  ego <- enrichGO(gene = signif_genes$ENTREZID,
                  keyType = "ENTREZID",
                  OrgDb = org.Hs.eg.db,
                  ont = "all",
                  pAdjustMethod = "BH",
                  qvalueCutoff = 0.05,
                  readable = TRUE)

  GO_list_MGH_Ubi_up[[comp]] <- ego
}

###

comparisons_DEU <- c("DelC_vs_WT", "DupC_vs_WT", "HetDel_vs_WT", "HomoDel_vs_WT", "PWSTI_vs_WT", "PWSTII_vs_WT")

GO_list_MGH_Ubi_down <- list()
for (comp in comparisons_DEU) {

  # Create background dataset for hypergeometric testing using all genes tested for significance in the results
  all_genes <- rowData(deu_MGH_results_new)$name

  # Extract significant results
  signif_genes <- rowData(deu_MGH_results_new)$name[which(rowData(deu_MGH_results_new)[, paste0(comp, "_p.adj")] < 0.1 &
                                                            rowData(deu_MGH_results_new)[, paste0(comp, "_diff")] < 0)]

  # Symbol to EntrezID
  all_genes <- bitr(all_genes, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = "org.Hs.eg.db")
  signif_genes <- bitr(signif_genes, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = "org.Hs.eg.db")

  ego <- enrichGO(gene = signif_genes$ENTREZID,
                  keyType = "ENTREZID",
                  OrgDb = org.Hs.eg.db,
                  ont = "all",
                  pAdjustMethod = "BH",
                  qvalueCutoff = 0.05,
                  readable = TRUE)

  GO_list_MGH_Ubi_down[[comp]] <- ego
}


names(GO_list_MGH_Ubi_up) <- c("Protein_DelC_up", "Protein_DupC_up", "Protein_HetDel_up", "Protein_HomoDel_up", "Protein_PWSTI_up", "Protein_PWSTII_up")
names(GO_list_MGH_Ubi_down) <- c("Protein_DelC_down", "Protein_DupC_down", "Protein_HetDel_down", "Protein_HomoDel_down", "Protein_PWSTI_down", "Protein_PWSTII_down")

```


Preparing GO lists - 8330 Ubiquitination
```{r}
comparisons_DEU <- c("DelC_vs_WT", "DupC_vs_WT", "HomoDel_vs_WT", "PWSTI_vs_WT", "PWSTII_vs_WT")

GO_list_8330_Ubi_up <- list()
for (comp in comparisons_DEU) {

  # Create background dataset for hypergeometric testing using all genes tested for significance in the results
  all_genes <- rowData(deu_8330_results_new)$name

  # Extract significant results
  signif_genes <- rowData(deu_8330_results_new)$name[which(rowData(deu_8330_results_new)[, paste0(comp, "_p.adj")] < 0.1 &
                                                            rowData(deu_8330_results_new)[, paste0(comp, "_diff")] > 0)]

  # Symbol to EntrezID
  all_genes <- bitr(all_genes, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = "org.Hs.eg.db")
  signif_genes <- bitr(signif_genes, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = "org.Hs.eg.db")

  ego <- enrichGO(gene = signif_genes$ENTREZID,
                  keyType = "ENTREZID",
                  OrgDb = org.Hs.eg.db,
                  ont = "all",
                  pAdjustMethod = "BH",
                  qvalueCutoff = 0.05,
                  readable = TRUE)

  GO_list_8330_Ubi_up[[comp]] <- ego
}

###

comparisons_DEU <- c("DelC_vs_WT", "DupC_vs_WT", "HomoDel_vs_WT", "PWSTI_vs_WT", "PWSTII_vs_WT")

GO_list_8330_Ubi_down <- list()
for (comp in comparisons_DEU) {

  # Create background dataset for hypergeometric testing using all genes tested for significance in the results
  all_genes <- rowData(deu_8330_results_new)$name

  # Extract significant results
  signif_genes <- rowData(deu_8330_results_new)$name[which(rowData(deu_8330_results_new)[, paste0(comp, "_p.adj")] < 0.1 &
                                                            rowData(deu_8330_results_new)[, paste0(comp, "_diff")] < 0)]

  # Symbol to EntrezID
  all_genes <- bitr(all_genes, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = "org.Hs.eg.db")
  signif_genes <- bitr(signif_genes, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = "org.Hs.eg.db")

  ego <- enrichGO(gene = signif_genes$ENTREZID,
                  keyType = "ENTREZID",
                  OrgDb = org.Hs.eg.db,
                  ont = "all",
                  pAdjustMethod = "BH",
                  qvalueCutoff = 0.05,
                  readable = TRUE)

  GO_list_8330_Ubi_down[[comp]] <- ego
}


names(GO_list_8330_Ubi_up) <- c("Protein_DelC_up", "Protein_DupC_up", "Protein_HomoDel_up", "Protein_PWSTI_up", "Protein_PWSTII_up")
names(GO_list_8330_Ubi_down) <- c("Protein_DelC_down", "Protein_DupC_down", "Protein_HomoDel_down", "Protein_PWSTI_down", "Protein_PWSTII_down")

```

Preparing GO term lists

```{r}

GO_list_MGH_Ubi_up_terms <- list()
for (comp in names(GO_list_MGH_Ubi_up)) {
  res <- GO_list_MGH_Ubi_up[[comp]]@result
  GO_list_MGH_Ubi_up_terms[[comp]] <- res[which(res$p.adjust < 0.05), "Description"]
}
GO_list_MGH_Ubi_down_terms <- list()
for (comp in names(GO_list_MGH_Ubi_down)) {
  res <- GO_list_MGH_Ubi_down[[comp]]@result
  GO_list_MGH_Ubi_down_terms[[comp]] <- res[which(res$p.adjust < 0.05), "Description"]
}

#####


GO_list_8330_Ubi_up_terms <- list()
for (comp in names(GO_list_8330_Ubi_up)) {
  res <- GO_list_8330_Ubi_up[[comp]]@result
  GO_list_8330_Ubi_up_terms[[comp]] <- res[which(res$p.adjust < 0.05), "Description"]
}
GO_list_8330_Ubi_down_terms <- list()
for (comp in names(GO_list_8330_Ubi_down)) {
  res <- GO_list_8330_Ubi_down[[comp]]@result
  GO_list_8330_Ubi_down_terms[[comp]] <- res[which(res$p.adjust < 0.05), "Description"]
}

```




Overlapping GO terms from both backgrounds - Downregulated
```{r}

GO_downregulated_Ubi_overlap <- list(
  GO_MGH_DEU_DelC = GO_list_MGH_Ubi_down_terms$Protein_DelC_down,
  GO_8330_DEU_DelC = GO_list_8330_Ubi_down_terms$Protein_DelC_down,
  GO_MGH_DEU_DupC = GO_list_MGH_Ubi_down_terms$Protein_DupC_down,
  GO_8330_DEU_DupC = GO_list_8330_Ubi_down_terms$Protein_DupC_down,
  GO_MGH_DEU_HomoDel = GO_list_MGH_Ubi_down_terms$Protein_HomoDel_down,
  GO_8330_DEU_HomoDel = GO_list_8330_Ubi_down_terms$Protein_HomoDel_down,
  GO_MGH_DEU_PWSTI = GO_list_MGH_Ubi_down_terms$Protein_PWSTI_down,
  GO_8330_DEU_PWSTI = GO_list_8330_Ubi_down_terms$Protein_PWSTI_down,
  GO_MGH_DEU_PWSTII = GO_list_MGH_Ubi_down_terms$Protein_PWSTII_down,
  GO_8330_DEU_PWSTII = GO_list_8330_Ubi_down_terms$Protein_PWSTII_down
)

lapply(GO_downregulated_Ubi_overlap, length)

# Euler Plots mit Labels und Titeln
plot(euler(list("MGH" = GO_downregulated_Ubi_overlap$GO_MGH_DEU_DelC,
                "8330" = GO_downregulated_Ubi_overlap$GO_8330_DEU_DelC)), 
     quantities = TRUE, labels = TRUE, main = "GO Downregulated Ubiquitination - DelC")

plot(euler(list("MGH" = GO_downregulated_Ubi_overlap$GO_MGH_DEU_DupC,
                "8330" = GO_downregulated_Ubi_overlap$GO_8330_DEU_DupC)), 
     quantities = TRUE, labels = TRUE, main = "GO Downregulated Ubiquitination - DupC")

plot(euler(list("MGH" = GO_downregulated_Ubi_overlap$GO_MGH_DEU_HomoDel,
                "8330" = GO_downregulated_Ubi_overlap$GO_8330_DEU_HomoDel)), 
     quantities = TRUE, labels = TRUE, main = "GO Downregulated Ubiquitination - HomoDel")

plot(euler(list("MGH" = GO_downregulated_Ubi_overlap$GO_MGH_DEU_PWSTI,
                "8330" = GO_downregulated_Ubi_overlap$GO_8330_DEU_PWSTI)), 
     quantities = TRUE, labels = TRUE, main = "GO Downregulated Ubiquitination - PWSTI")

plot(euler(list("MGH" = GO_downregulated_Ubi_overlap$GO_MGH_DEU_PWSTII,
                "8330" = GO_downregulated_Ubi_overlap$GO_8330_DEU_PWSTII)), 
     quantities = TRUE, labels = TRUE, main = "GO Downregulated Ubiquitination - PWSTII")

GO_list_Ubi_down_overlap_terms <- list(
  "GO_overlap_DEU_DelC" = intersect(GO_downregulated_Ubi_overlap$GO_MGH_DEU_DelC,
                                    GO_downregulated_Ubi_overlap$GO_8330_DEU_DelC),
  "GO_overlap_DEU_DupC" = intersect(GO_downregulated_Ubi_overlap$GO_MGH_DEU_DupC,
                                    GO_downregulated_Ubi_overlap$GO_8330_DEU_DupC),
  "GO_overlap_DEU_HomoDel" = intersect(GO_downregulated_Ubi_overlap$GO_MGH_DEU_HomoDel,
                                       GO_downregulated_Ubi_overlap$GO_8330_DEU_HomoDel),
  "GO_overlap_DEU_PWSTI" = intersect(GO_downregulated_Ubi_overlap$GO_MGH_DEU_PWSTI,
                                     GO_downregulated_Ubi_overlap$GO_8330_DEU_PWSTI),
  "GO_overlap_DEU_PWSTII" = intersect(GO_downregulated_Ubi_overlap$GO_MGH_DEU_PWSTII,
                                      GO_downregulated_Ubi_overlap$GO_8330_DEU_PWSTII)
)

```


Overlapping GO terms from both backgrounds - Upregulated
```{r}

GO_upregulated_Ubi_overlap <- list(
  GO_MGH_DEU_DelC = GO_list_MGH_Ubi_up_terms$Protein_DelC_up,
  GO_8330_DEU_DelC = GO_list_8330_Ubi_up_terms$Protein_DelC_up,
  GO_MGH_DEU_DupC = GO_list_MGH_Ubi_up_terms$Protein_DupC_up,
  GO_8330_DEU_DupC = GO_list_8330_Ubi_up_terms$Protein_DupC_up,
  GO_MGH_DEU_HomoDel = GO_list_MGH_Ubi_up_terms$Protein_HomoDel_up,
  GO_8330_DEU_HomoDel = GO_list_8330_Ubi_up_terms$Protein_HomoDel_up,
  GO_MGH_DEU_PWSTI = GO_list_MGH_Ubi_up_terms$Protein_PWSTI_up,
  GO_8330_DEU_PWSTI = GO_list_8330_Ubi_up_terms$Protein_PWSTI_up,
  GO_MGH_DEU_PWSTII = GO_list_MGH_Ubi_up_terms$Protein_PWSTII_up,
  GO_8330_DEU_PWSTII = GO_list_8330_Ubi_up_terms$Protein_PWSTII_up
)

lapply(GO_upregulated_Ubi_overlap, length)

# Euler Plots mit Labels und Titeln
plot(euler(list("MGH" = GO_upregulated_Ubi_overlap$GO_MGH_DEU_DelC,
                "8330" = GO_upregulated_Ubi_overlap$GO_8330_DEU_DelC)), 
     quantities = TRUE, labels = TRUE, main = "GO Upregulated Ubiquitination - DelC")

plot(euler(list("MGH" = GO_upregulated_Ubi_overlap$GO_MGH_DEU_DupC,
                "8330" = GO_upregulated_Ubi_overlap$GO_8330_DEU_DupC)), 
     quantities = TRUE, labels = TRUE, main = "GO Upregulated Ubiquitination - DupC")

plot(euler(list("MGH" = GO_upregulated_Ubi_overlap$GO_MGH_DEU_HomoDel,
                "8330" = GO_upregulated_Ubi_overlap$GO_8330_DEU_HomoDel)), 
     quantities = TRUE, labels = TRUE, main = "GO Upregulated Ubiquitination - HomoDel")

plot(euler(list("MGH" = GO_upregulated_Ubi_overlap$GO_MGH_DEU_PWSTI,
                "8330" = GO_upregulated_Ubi_overlap$GO_8330_DEU_PWSTI)), 
     quantities = TRUE, labels = TRUE, main = "GO Upregulated Ubiquitination - PWSTI")

plot(euler(list("MGH" = GO_upregulated_Ubi_overlap$GO_MGH_DEU_PWSTII,
                "8330" = GO_upregulated_Ubi_overlap$GO_8330_DEU_PWSTII)), 
     quantities = TRUE, labels = TRUE, main = "GO Upregulated Ubiquitination - PWSTII")

GO_list_Ubi_up_overlap_terms <- list(
  "GO_overlap_DEU_DelC" = intersect(GO_upregulated_Ubi_overlap$GO_MGH_DEU_DelC,
                                    GO_upregulated_Ubi_overlap$GO_8330_DEU_DelC),
  "GO_overlap_DEU_DupC" = intersect(GO_upregulated_Ubi_overlap$GO_MGH_DEU_DupC,
                                    GO_upregulated_Ubi_overlap$GO_8330_DEU_DupC),
  "GO_overlap_DEU_HomoDel" = intersect(GO_upregulated_Ubi_overlap$GO_MGH_DEU_HomoDel,
                                       GO_upregulated_Ubi_overlap$GO_8330_DEU_HomoDel),
  "GO_overlap_DEU_PWSTI" = intersect(GO_upregulated_Ubi_overlap$GO_MGH_DEU_PWSTI,
                                     GO_upregulated_Ubi_overlap$GO_8330_DEU_PWSTI),
  "GO_overlap_DEU_PWSTII" = intersect(GO_upregulated_Ubi_overlap$GO_MGH_DEU_PWSTII,
                                      GO_upregulated_Ubi_overlap$GO_8330_DEU_PWSTII)
)

```






10.9 Upset Plots - GO terms - Ubiquitin only - Preparation
```{r}
intersectionsOI_onlyUbi <-
  list(
    #single
    "DEU_DelC", 
    "DEU_DupC", 
    "DEU_HomoDel", 
    "DEU_PWSTI", 
    "DEU_PWSTII",

    #SYS-specific Ubi
    c("DEU_DelC", "DEU_DupC"),

    #PWS-specific Ubi
    c("DEU_PWSTI", "DEU_PWSTII"),

    #deletion-specific Ubi
    c("DEU_HomoDel",
         "DEU_PWSTI", "DEU_PWSTII"),

    #all Ubi without DelC
    c("DEU_DupC",
         "DEU_HomoDel",
         "DEU_PWSTI","DEU_PWSTII"),

    #all Ubi
    c("DEU_DelC","DEU_DupC",
         "DEU_HomoDel",
         "DEU_PWSTI","DEU_PWSTII")

    )



```


10.10 Upset Plots - GO terms - Ubiquitin only - Downregulated
```{r}
downregulated_onlyUbi <- list(
  DEU_DelC = list_Ubi_down_overlap$overlap_DEU_DelC,
  DEU_DupC = list_Ubi_down_overlap$overlap_DEU_DupC,
  DEU_HomoDel = list_Ubi_down_overlap$overlap_DEU_HomoDel,
  DEU_PWSTI = list_Ubi_down_overlap$overlap_DEU_PWSTI,
  DEU_PWSTII = list_Ubi_down_overlap$overlap_DEU_PWSTII
)

downregulated_complex_onlyUbi <- fromList(downregulated_onlyUbi) == 1
downregulated_complex_onlyUbi <- as.data.frame(downregulated_complex_onlyUbi)

my_corrected_theme <- list(
  'intersections_matrix' = theme(
    legend.margin = margin(t = 5, r = 5, b = 5, l = 5, unit = 'pt')
  ),
  'overall_sizes' = theme(
  ),
  'selected_theme' = theme() 
)

#UpSet plot

ComplexUpset::upset(downregulated_complex_onlyUbi,
      colnames(downregulated_complex_onlyUbi),
      sort_sets = FALSE,
      intersections = intersectionsOI_onlyUbi,
      sort_intersections=FALSE,
      theme = my_corrected_theme,

      ) +
  rremove("xlab")




#Query intersections of interest

downregulated_onlyUbi_df <- fromList(downregulated_onlyUbi)
downregulated_onlyUbi_df$GO <- unique(unlist(downregulated_onlyUbi))

downregulated_onlyUbi_df_with_intersection <- downregulated_onlyUbi_df %>%
  tidyr::unite(col = "intersection", -c("GO"), sep = "")

downregulated_onlyUbi_df_with_intersection <-
  downregulated_onlyUbi_df_with_intersection %>%
  group_by(intersection) %>%
  summarise(list = list(GO)) %>%
  mutate(list = setNames(list, intersection)) %>%
  pull(list)

#DelC ubi-specific
downregulated_onlyUbi_df_with_intersection[["10000"]]

#DupC ubi-specific
downregulated_onlyUbi_df_with_intersection[["01000"]]

#HomoDel ubi-specific
downregulated_onlyUbi_df_with_intersection[["00100"]]

#PWSTI ubi-specific
downregulated_onlyUbi_df_with_intersection[["00010"]]

#PWSTII ubi-specific
downregulated_onlyUbi_df_with_intersection[["00001"]]

#SYS ubi only
downregulated_onlyUbi_df_with_intersection[["11000"]]

#PWS ubi only
downregulated_onlyUbi_df_with_intersection[["00011"]]

#PWS and MAGEL2-del ubi only
downregulated_onlyUbi_df_with_intersection[["00111"]]

#all DelC-exclusive ubi only
downregulated_onlyUbi_df_with_intersection[["01111"]]

#all ubi
downregulated_onlyUbi_df_with_intersection[["11111"]]
```

10.10 Upset Plots - GO terms - Ubiquitin only - Upregulated
```{r}
upregulated_onlyUbi <- list(
  DEU_DelC = list_Ubi_up_overlap$overlap_DEU_DelC,
  DEU_DupC = list_Ubi_up_overlap$overlap_DEU_DupC,
  DEU_HomoDel = list_Ubi_up_overlap$overlap_DEU_HomoDel,
  DEU_PWSTI = list_Ubi_up_overlap$overlap_DEU_PWSTI,
  DEU_PWSTII = list_Ubi_up_overlap$overlap_DEU_PWSTII
)

upregulated_complex_onlyUbi <- fromList(upregulated_onlyUbi) == 1
upregulated_complex_onlyUbi <- as.data.frame(upregulated_complex_onlyUbi)

#UpSet plot

ComplexUpset::upset(upregulated_complex_onlyUbi,
      colnames(upregulated_complex_onlyUbi),
      sort_sets = FALSE,
      intersections = intersectionsOI_onlyUbi,
      sort_intersections=FALSE,
      theme = my_corrected_theme,

      ) +
  rremove("xlab")




#Query intersections of interest

upregulated_onlyUbi_df <- fromList(upregulated_onlyUbi)
upregulated_onlyUbi_df$GO <- unique(unlist(upregulated_onlyUbi))

upregulated_onlyUbi_df_with_intersection <- upregulated_onlyUbi_df %>%
  tidyr::unite(col = "intersection", -c("GO"), sep = "")

upregulated_onlyUbi_df_with_intersection <-
  upregulated_onlyUbi_df_with_intersection %>%
  group_by(intersection) %>%
  summarise(list = list(GO)) %>%
  mutate(list = setNames(list, intersection)) %>%
  pull(list)

#DelC ubi-specific
upregulated_onlyUbi_df_with_intersection[["10000"]]

#DupC ubi-specific
upregulated_onlyUbi_df_with_intersection[["01000"]]

#HomoDel ubi-specific
upregulated_onlyUbi_df_with_intersection[["00100"]]

#PWSTI ubi-specific
upregulated_onlyUbi_df_with_intersection[["00010"]]

#PWSTII ubi-specific
upregulated_onlyUbi_df_with_intersection[["00001"]]

#SYS ubi only
upregulated_onlyUbi_df_with_intersection[["11000"]]

#PWS ubi only
upregulated_onlyUbi_df_with_intersection[["00011"]]

#PWS and MAGEL2-del ubi only
upregulated_onlyUbi_df_with_intersection[["00111"]]

#all DelC-exclusive ubi only
upregulated_onlyUbi_df_with_intersection[["01111"]]

#all ubi
upregulated_onlyUbi_df_with_intersection[["11111"]]
```
