---
title: "Transcriptomics_vs_Proteomics"
output: html_document
date: "2025-08-26"
---

```{r}
#RNA-seq
dds_RNA <- readRDS(file = "./dds_RNA_30102024.rds")
DEG_list_MGH <- readRDS(file = "./DEG_list_MGH_30102024.rds")
DEG_list_8330 <- readRDS(file = "./DEG_list_8330_30102024.rds")


#proteomics
data_imp_MGH_new <- readRDS(file = "./data_imp_MGH_BH_new.rds")
dep_MGH_results_new <- readRDS(file = "./dep_MGH_BH_new.rds")

data_imp_8330_new <- readRDS(file = "./data_imp_8330_BH_new.rds")
dep_8330_results_new <- readRDS(file = "./dep_8330_BH_new.rds")

```


Ridgeline Plot of proteomics / transcriptomics ratio 
```{r}
# Step 1: Load necessary libraries
library(SummarizedExperiment)
library(DESeq2)
library(tidyverse)
library(ggridges)

# --- User-defined variables ---
gg_color_hue <- function(n) {
  hues <- seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}
condition_order <- c("DelC", "DupC", "HetDel", "HomoDel", "PWSTI", "PWSTII", "WT")
genotype_colors <- gg_color_hue(length(condition_order))
names(genotype_colors) <- condition_order

# --- Step 2: Normalize RNA-seq data ---
# Note: VST data is already log2-like and library size normalized
vst_data <- vst(dds_RNA, blind = TRUE)
vst_counts <- assay(vst_data)

# --- Step 3: Create a Definitive Sample Mapping Table ---
# We will create one map for each background.

# --- MGH Mapping Table ---
meta_prot_mgh <- as.data.frame(colData(data_imp_MGH_new)) %>%
  rownames_to_column("Proteomics_ID")

map_mgh <- meta_prot_mgh %>%
  mutate(
    # Handle the 'HetDel' vs 'HETDEL' inconsistency
    condition_rna = ifelse(condition == "HetDel", "HETDEL", as.character(condition)),
    # Reconstruct the RNA_ID
    RNA_ID = paste0("MGH", condition_rna, replicate)
  ) %>%
  # Select only the mapping columns
  dplyr::select(Proteomics_ID, RNA_ID, condition, replicate)

# --- 8330 Mapping Table ---
meta_prot_8330 <- as.data.frame(colData(data_imp_8330_new)) %>%
  rownames_to_column("Proteomics_ID")

map_8330 <- meta_prot_8330 %>%
  mutate(
    condition_rna = ifelse(condition == "HetDel", "HETDEL", as.character(condition)),
    RNA_ID = paste0("8330", condition_rna, replicate)
  ) %>%
  dplyr::select(Proteomics_ID, RNA_ID, condition, replicate)


# --- Step 4: Define a new, simpler function that uses the map ---

prepare_ratio_data_mapped <- function(proteomics_se, vst_counts, sample_map, condition_order_vector) {
  # A: Tidy Proteomics Data
  protein_tidy <- assay(proteomics_se) %>%
    as.data.frame() %>%
    rownames_to_column("Gene") %>%
    pivot_longer(-Gene, names_to = "Proteomics_ID", values_to = "Protein_Abundance")

  # B: Tidy RNA-seq Data
  # Subset VST counts to only the samples in our map
  rna_subset <- vst_counts[, colnames(vst_counts) %in% sample_map$RNA_ID, drop = FALSE]
  rna_tidy <- rna_subset %>%
    as.data.frame() %>%
    rownames_to_column("Gene") %>%
    pivot_longer(-Gene, names_to = "RNA_ID", values_to = "RNA_Abundance")

  # C: Merge using the map and Calculate Ratio
  merged_data <- protein_tidy %>%
    left_join(sample_map, by = "Proteomics_ID") %>%
    left_join(rna_tidy, by = c("Gene", "RNA_ID")) %>%
    dplyr::filter(!is.na(RNA_Abundance)) # Keep only matched data

  ratio_data <- merged_data %>%
    mutate(
      Log2Ratio = Protein_Abundance - RNA_Abundance, 
      condition = factor(condition, levels = rev(condition_order_vector))
    )
  return(ratio_data)
}


# --- Step 5: Generate data for each background using the new function and maps ---
mgh_ratio_data <- prepare_ratio_data_mapped(data_imp_MGH_new, vst_counts, map_mgh, condition_order)
h8330_ratio_data <- prepare_ratio_data_mapped(data_imp_8330_new, vst_counts, map_8330, condition_order)


# --- Step 6: Create and Customize the Ridgeline Plots ---

# Define a consistent theme for a cleaner, publication-ready look
publication_theme <- theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
    axis.title.x = element_text(face = "bold", margin = margin(t = 10)),
    axis.text.y = element_text(face = "bold"),
    axis.title.y = element_blank(), # The y-axis labels are self-explanatory
    legend.position = "none" # Hide legend as it's redundant with the y-axis
  )

# Plot for MGH background
mgh_ridgeline_plot <- ggplot(mgh_ratio_data, aes(x = Log2Ratio, y = condition, fill = condition, color = condition)) +
  geom_density_ridges(alpha = 0.7, scale = 1.2, size = 0.8) + # Outline color now mapped in aes()
  geom_vline(xintercept = median(mgh_ratio_data$Log2Ratio, na.rm = TRUE),
             linetype = "dashed", color = "black", size = 0.8) +
  scale_fill_manual(values = genotype_colors) +
  scale_color_manual(values = genotype_colors) + # Added scale for the outline color
  labs(
    title = "Distribution of Protein/RNA Ratios (MGH Background)",
    x = bquote(Log[2]~"(Protein / RNA Abundance)")
  ) +
  # Use coord_cartesian to zoom in without dropping data points
  coord_cartesian(xlim = quantile(mgh_ratio_data$Log2Ratio, c(0.005, 0.995), na.rm = TRUE)) +
  publication_theme

# Plot for 8330 background
h8330_ridgeline_plot <- ggplot(h8330_ratio_data, aes(x = Log2Ratio, y = condition, fill = condition, color = condition)) +
  geom_density_ridges(alpha = 0.7, scale = 1.2, size = 0.8) + # Outline color now mapped in aes()
  geom_vline(xintercept = median(h8330_ratio_data$Log2Ratio, na.rm = TRUE),
             linetype = "dashed", color = "black", size = 0.8) +
  scale_fill_manual(values = genotype_colors) +
  scale_color_manual(values = genotype_colors) + # Added scale for the outline color
  labs(
    title = "Distribution of Protein/RNA Ratios (8330 Background)",
    x = bquote(Log[2]~"(Protein / RNA Abundance)")
  ) +
  coord_cartesian(xlim = quantile(h8330_ratio_data$Log2Ratio, c(0.005, 0.995), na.rm = TRUE)) +
  publication_theme


# --- Step 7: Display the plots ---
print(mgh_ridgeline_plot)
print(h8330_ridgeline_plot)
```



```{r}
mgh_mean_ratios <- mgh_ratio_data %>%
  dplyr::group_by(condition) %>%
  dplyr::summarize(
    mean_log2_ratio = mean(Log2Ratio, na.rm = TRUE),
    sd_log2_ratio = sd(Log2Ratio, na.rm = TRUE)
  )

print(mgh_mean_ratios)

h8330_mean_ratios <- h8330_ratio_data %>%
  dplyr::group_by(condition) %>%
  dplyr::summarize(
    mean_log2_ratio = mean(Log2Ratio, na.rm = TRUE),
    sd_log2_ratio = sd(Log2Ratio, na.rm = TRUE)
  )

print(h8330_mean_ratios)
```

Data Preparation
```{r}
# --- Step 1: Recreate the Master Integrated Data Table ---
# This part is the same as before, ensuring the base data exists.
library(tidyverse)
library(SummarizedExperiment)

comparison_map <- list(
  "MGH_DelC_vs_WT"    = list(prot_lfc = "DelC_vs_WT_diff", prot_padj = "DelC_vs_WT_p.adj"),
  "MGH_DupC_vs_WT"    = list(prot_lfc = "DupC_vs_WT_diff", prot_padj = "DupC_vs_WT_p.adj"),
  "MGH_HetDel_vs_WT"  = list(prot_lfc = "HetDel_vs_WT_diff", prot_padj = "HetDel_vs_WT_p.adj"),
  "MGH_HomoDel_vs_WT" = list(prot_lfc = "HomoDel_vs_WT_diff", prot_padj = "HomoDel_vs_WT_p.adj"),
  "MGH_PWSTI_vs_WT"   = list(prot_lfc = "PWSTI_vs_WT_diff", prot_padj = "PWSTI_vs_WT_p.adj"),
  "MGH_PWSTII_vs_WT"  = list(prot_lfc = "PWSTII_vs_WT_diff", prot_padj = "PWSTII_vs_WT_p.adj"),
  "8330_DelC_vs_WT"    = list(prot_lfc = "DelC_vs_WT_diff", prot_padj = "DelC_vs_WT_p.adj"),
  "8330_DupC_vs_WT"    = list(prot_lfc = "DupC_vs_WT_diff", prot_padj = "DupC_vs_WT_p.adj"),
  "8330_HetDel_vs_WT"  = list(prot_lfc = "HetDel_vs_WT_diff", prot_padj = "HetDel_vs_WT_p.adj"),
  "8330_HomoDel_vs_WT" = list(prot_lfc = "HomoDel_vs_WT_diff", prot_padj = "HomoDel_vs_WT_p.adj"),
  "8330_PWSTI_vs_WT"   = list(prot_lfc = "PWSTI_vs_WT_diff", prot_padj = "PWSTI_vs_WT_p.adj"),
  "8330_PWSTII_vs_WT"  = list(prot_lfc = "PWSTII_vs_WT_diff", prot_padj = "PWSTII_vs_WT_p.adj")
)
all_deg_lists <- list("MGH" = DEG_list_MGH, "8330" = DEG_list_8330)
all_prot_results <- list("MGH" = dep_MGH_results_new, "8330" = dep_8330_results_new)
integrated_list <- list()

for (background in names(all_deg_lists)) {
  deg_list <- all_deg_lists[[background]]
  prot_results_df <- as.data.frame(rowData(all_prot_results[[background]]))
  for (comp_short_name in names(deg_list)) {
    rna_df <- as.data.frame(deg_list[[comp_short_name]]) %>%
      rownames_to_column("Gene") %>%
      dplyr::select(Gene, mRNA_Log2FC = log2FoldChange, mRNA_padj = padj)
    comp_full_name <- paste0(background, "_", gsub("genotype_|_vs_WT", "", comp_short_name), "_vs_WT")
    prot_lfc_col <- comparison_map[[comp_full_name]]$prot_lfc
    prot_padj_col <- comparison_map[[comp_full_name]]$prot_padj
    prot_df <- prot_results_df %>%
      dplyr::select(Gene = PG.Genes, Protein_Log2FC = all_of(prot_lfc_col), Protein_padj = all_of(prot_padj_col))
    integrated_df <- inner_join(rna_df, prot_df, by = "Gene") %>%
      mutate(comparison = comp_full_name) %>%
      na.omit()
    integrated_list[[comp_full_name]] <- integrated_df
  }
}
master_integrated_table <- bind_rows(integrated_list)


# --- Step 2: Recreate the Quadrant Data Table ---
# This is the crucial step that was missing.
quadrant_data <- master_integrated_table %>%
  mutate(
    quadrant = case_when(
      mRNA_Log2FC > 0 & Protein_Log2FC > 0   ~ "Up-Up (Correlated)",
      mRNA_Log2FC < 0 & Protein_Log2FC < 0   ~ "Down-Down (Correlated)",
      mRNA_Log2FC > 0 & Protein_Log2FC < 0   ~ "Up-Down (Post-transcriptional Down-regulation)",
      mRNA_Log2FC < 0 & Protein_Log2FC > 0   ~ "Down-Up (Post-transcriptional Up-regulation)",
      TRUE                                   ~ "No Change"
    )
  )

# --- Step 3: Verify the result ---
message("quadrant_data has been successfully recreated.")
print(head(quadrant_data))
```





```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(ggrepel)
library(stringr)

# ==============================================================================
# STEP 1: DEFINE THE "STRICT" UNIVERSE (Same as before)
# ==============================================================================

conditions <- c("DelC", "DupC", "HetDel", "HomoDel", "PWSTI", "PWSTII")
backgrounds <- c("MGH", "8330")

all_comparisons <- unlist(lapply(backgrounds, function(bg) {
  paste0(bg, "_", conditions, "_vs_WT")
}))

# Initialize strict list
first_comp <- all_comparisons[1]
strict_common_genes <- master_integrated_table %>%
  dplyr::filter(comparison == first_comp & !is.na(Protein_Log2FC) & !is.na(mRNA_Log2FC)) %>%
  pull(Gene)

# Intersect across all
for(comp in all_comparisons[-1]) {
  current_genes <- master_integrated_table %>%
    dplyr::filter(comparison == comp & !is.na(Protein_Log2FC) & !is.na(mRNA_Log2FC)) %>%
    pull(Gene)
  strict_common_genes <- intersect(strict_common_genes, current_genes)
}

cat("\n--- Strict Universe Check ---\n")
cat("Proteins detected in ALL 12 comparisons:", length(strict_common_genes), "\n\n")


# ==============================================================================
# STEP 2: SCATTER PLOT FUNCTION (Protein-Driven Discordance)
# ==============================================================================

create_rna_prot_scatter <- function(cond_name) {
  
  # Construct comparison strings
  mgh_comp_string  <- paste0("MGH_", cond_name, "_vs_WT")
  comp8330_string  <- paste0("8330_", cond_name, "_vs_WT")
  
  cat(paste("Processing:", cond_name, "\n"))

  # --- A. Prepare MGH Data ---
  df_mgh <- master_integrated_table %>%
    dplyr::filter(comparison == mgh_comp_string) %>%
    dplyr::filter(Gene %in% strict_common_genes) %>%
    dplyr::select(Gene, mRNA_Log2FC, Protein_Log2FC, mRNA_padj, Protein_padj) %>%
    rename(
      MGH_RNA_FC = mRNA_Log2FC, 
      MGH_Prot_FC = Protein_Log2FC,
      MGH_RNA_padj = mRNA_padj,
      MGH_Prot_padj = Protein_padj
    )
  
  # --- B. Prepare 8330 Data ---
  df_8330 <- master_integrated_table %>%
    dplyr::filter(comparison == comp8330_string) %>%
    dplyr::filter(Gene %in% strict_common_genes) %>%
    dplyr::select(Gene, mRNA_Log2FC, Protein_Log2FC, mRNA_padj, Protein_padj) %>%
    rename(
      bg8330_RNA_FC = mRNA_Log2FC, 
      bg8330_Prot_FC = Protein_Log2FC,
      bg8330_RNA_padj = mRNA_padj,
      bg8330_Prot_padj = Protein_padj
    )
  
  # Check data existence
  if(nrow(df_mgh) == 0 || nrow(df_8330) == 0) {
    return(paste("ERROR: Missing data for", cond_name))
  }

  # --- C. Merge ---
  merged_df <- inner_join(df_mgh, df_8330, by = "Gene")
  cat(paste("  > Points to plot:", nrow(merged_df), "\n"))
  
  # --- D. Apply New "Protein-Only" Logic ---
  
  # Helper function for your specific logic
  # Returns TRUE if Protein is driving the change (Prot Sig, RNA Not Sig in same direction)
  is_protein_driven <- function(prot_padj, prot_fc, rna_padj, rna_fc) {
    # Case 1: Protein Sig UP, RNA NOT Sig UP
    prot_up <- (prot_padj < 0.05 & prot_fc > 0)
    rna_up  <- (rna_padj < 0.05 & rna_fc > 0)
    case1   <- prot_up & !rna_up
    
    # Case 2: Protein Sig DOWN, RNA NOT Sig DOWN
    prot_down <- (prot_padj < 0.05 & prot_fc < 0)
    rna_down  <- (rna_padj < 0.05 & rna_fc < 0)
    case2     <- prot_down & !rna_down
    
    return(case1 | case2)
  }

  merged_df <- merged_df %>%
    mutate(
      # Apply logic to MGH
      is_discordant_MGH = is_protein_driven(MGH_Prot_padj, MGH_Prot_FC, MGH_RNA_padj, MGH_RNA_FC),
      
      # Apply logic to 8330
      is_discordant_8330 = is_protein_driven(bg8330_Prot_padj, bg8330_Prot_FC, bg8330_RNA_padj, bg8330_RNA_FC),
      
      # Check if Protein Direction is consistent between backgrounds
      # (We only care if the protein change is consistent)
      prot_dir_match = sign(MGH_Prot_FC) == sign(bg8330_Prot_FC),
      
      # Final Category: Must be discordant in BOTH backgrounds and protein direction must match
      plot_category = case_when(
        is_discordant_MGH & is_discordant_8330 & prot_dir_match ~ "Protein-Driven (Significant)",
        TRUE ~ "Other"
      )
    )
  
  # --- E. Labeling ---
  proteins_to_label <- c("NRAS","SLC2A1","SMAD4","NONO", "AKT2", "DYNC2H1", "PTPN11", "BRAF", "PAFAH1B1", "GRIA3", "NTRK2", "NDN")
  merged_df$label <- ifelse(merged_df$Gene %in% proteins_to_label, merged_df$Gene, "")
  
  # --- F. Plotting ---
  p <- ggplot(merged_df, aes(x = MGH_RNA_FC, y = MGH_Prot_FC, color = plot_category)) +
    
    # 1. Non-significant (Light Grey)
    geom_point(data = subset(merged_df, plot_category == "Other"), 
               color = "#ebebeb", size = 2, alpha = 0.8) +
    
    # 2. Significant (Dark Grey/Silver)
    geom_point(data = subset(merged_df, plot_category == "Protein-Driven (Significant)"), 
               color = "#c0c0c0", size = 2, alpha = 1) +
    
    # 3. Labeled points (Red)
    geom_point(data = subset(merged_df, label != ""), 
               color = "#b34747", size = 2) +
    
    # 4. Text Labels
    geom_text_repel(aes(label = label), data = subset(merged_df, label != ""), 
                    color = "#b34747", size = 3, 
                    box.padding = 0.5, point.padding = 0.5, max.overlaps = 20) +
    
    # 5. Styling
    labs(
      title = paste0("MGH ", cond_name, ": Protein-Driven Discordance"),
      subtitle = "Highlighting significant protein changes NOT driven by mRNA (Consistent in 8330)",
      x = "mRNA Log2 Fold Change",
      y = "Protein Log2 Fold Change",
      color = ""
    ) +
    theme_bw() +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      legend.position = "bottom"
    ) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
    geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
    coord_cartesian(xlim = c(-7, 7), ylim = c(-5, 5))
  
  return(p)
}

# --- Execute for all conditions ---
for (cond in conditions) {
  print(create_rna_prot_scatter(cond))
}
```



scatter plot v2
```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(ggrepel)
library(stringr)
library(tidyverse)

# ==============================================================================
# STEP 1: STRICT FILTERING (The "Must be in Bucket 1/2 everywhere" Rule)
# ==============================================================================

# --- A. Define Bucket 1 (Protein UP / RNA Not UP) ---
bucket1_all <- master_integrated_table %>%
  dplyr::filter(
    Protein_padj < 0.05 & Protein_Log2FC > 0,      # Prot Sig UP
    !(mRNA_padj < 0.05 & mRNA_Log2FC > 0)          # RNA NOT Sig UP
  )

# Find genes present in ALL 12 Bucket 1 lists
list_b1 <- bucket1_all %>% group_split(comparison) %>% lapply(function(df) df$Gene)
if(length(list_b1) < 12) {
  strict_up_genes <- character(0) 
} else {
  strict_up_genes <- Reduce(intersect, list_b1)
}

# --- B. Define Bucket 2 (Protein DOWN / RNA Not DOWN) ---
bucket2_all <- master_integrated_table %>%
  dplyr::filter(
    Protein_padj < 0.05 & Protein_Log2FC < 0,      # Prot Sig DOWN
    !(mRNA_padj < 0.05 & mRNA_Log2FC < 0)          # RNA NOT Sig DOWN
  )

# Find genes present in ALL 12 Bucket 2 lists
list_b2 <- bucket2_all %>% group_split(comparison) %>% lapply(function(df) df$Gene)
if(length(list_b2) < 12) {
  strict_down_genes <- character(0)
} else {
  strict_down_genes <- Reduce(intersect, list_b2)
}

cat("\n--- Strict Selection Check ---\n")
cat("Genes strictly discordant UP in all 12 conditions:", length(strict_up_genes), "\n")
cat("Genes strictly discordant DOWN in all 12 conditions:", length(strict_down_genes), "\n")


# ==============================================================================
# STEP 2: RANK BY DIVERGENCE (Pick Top 5 from the Strict Lists)
# ==============================================================================

# Helper to calculate mean divergence for a specific gene list
get_top_divergent <- function(gene_list, top_n, direction = "up") {
  if(length(gene_list) == 0) return(character(0))
  
  ranked <- master_integrated_table %>%
    dplyr::filter(Gene %in% gene_list) %>%
    mutate(diff = Protein_Log2FC - mRNA_Log2FC) %>%
    group_by(Gene) %>%
    summarize(mean_diff = mean(diff, na.rm = TRUE))
  
  if(direction == "up") {
    # Largest Positive Difference (Prot >> RNA)
    return(head(arrange(ranked, desc(mean_diff)), top_n)$Gene)
  } else {
    # Largest Negative Difference (Prot << RNA)
    return(head(arrange(ranked, mean_diff), top_n)$Gene)
  }
}

# Get Top 5 Strict UP and Top 5 Strict DOWN
final_top_5_up   <- get_top_divergent(strict_up_genes, 5, "up")
final_top_5_down <- get_top_divergent(strict_down_genes, 5, "down")

targets_global <- c(final_top_5_up, final_top_5_down)

cat("\n--- Final Targets to Highlight ---\n")
print(targets_global)


# ==============================================================================
# STEP 3: PLOTTING (With Halo Effect for Strict Targets)
# ==============================================================================

# We need the "Strict Universe" just for the background data filtering in the plot function
# (This ensures the plot only shows genes detected in all experiments, consistent with previous plots)
conditions <- c("DelC", "DupC", "HetDel", "HomoDel", "PWSTI", "PWSTII")
backgrounds <- c("MGH", "8330")
all_comparisons <- unlist(lapply(backgrounds, function(bg) paste0(bg, "_", conditions, "_vs_WT")))
first_comp <- all_comparisons[1]
base_strict_genes <- master_integrated_table %>% filter(comparison == first_comp) %>% pull(Gene)
for(comp in all_comparisons[-1]) {
  curr <- master_integrated_table %>% filter(comparison == comp) %>% pull(Gene)
  base_strict_genes <- intersect(base_strict_genes, curr)
}

create_rna_prot_scatter <- function(cond_name) {
  
  mgh_comp_string  <- paste0("MGH_", cond_name, "_vs_WT")
  comp8330_string  <- paste0("8330_", cond_name, "_vs_WT")
  
  # Prepare Data (Filtering for the base strict universe)
  df_mgh <- master_integrated_table %>%
    dplyr::filter(comparison == mgh_comp_string, Gene %in% base_strict_genes) %>%
    dplyr::select(Gene, mRNA_Log2FC, Protein_Log2FC, mRNA_padj, Protein_padj) %>%
    rename(MGH_RNA_FC=mRNA_Log2FC, MGH_Prot_FC=Protein_Log2FC, MGH_RNA_padj=mRNA_padj, MGH_Prot_padj=Protein_padj)
  
  df_8330 <- master_integrated_table %>%
    dplyr::filter(comparison == comp8330_string, Gene %in% base_strict_genes) %>%
    dplyr::select(Gene, mRNA_Log2FC, Protein_Log2FC, mRNA_padj, Protein_padj) %>%
    rename(bg8330_RNA_FC=mRNA_Log2FC, bg8330_Prot_FC=Protein_Log2FC, bg8330_RNA_padj=mRNA_padj, bg8330_Prot_padj=Protein_padj)
  
  if(nrow(df_mgh) == 0 || nrow(df_8330) == 0) return(NULL)
  
  merged_df <- inner_join(df_mgh, df_8330, by = "Gene")
  
  # Logic: Red/Blue only if they meet the criteria locally (for the colors)
  is_protein_driven <- function(prot_padj, prot_fc, rna_padj, rna_fc) {
    prot_up <- (prot_padj < 0.05 & prot_fc > 0); rna_up <- (rna_padj < 0.05 & rna_fc > 0)
    prot_down <- (prot_padj < 0.05 & prot_fc < 0); rna_down <- (rna_padj < 0.05 & rna_fc < 0)
    return((prot_up & !rna_up) | (prot_down & !rna_down))
  }

  merged_df <- merged_df %>%
    mutate(
      is_discordant_MGH = is_protein_driven(MGH_Prot_padj, MGH_Prot_FC, MGH_RNA_padj, MGH_RNA_FC),
      is_discordant_8330 = is_protein_driven(bg8330_Prot_padj, bg8330_Prot_FC, bg8330_RNA_padj, bg8330_RNA_FC),
      prot_dir_match = sign(MGH_Prot_FC) == sign(bg8330_Prot_FC),
      plot_category = case_when(
        is_discordant_MGH & is_discordant_8330 & prot_dir_match & MGH_Prot_FC > 0 ~ "Protein-Driven Up",
        is_discordant_MGH & is_discordant_8330 & prot_dir_match & MGH_Prot_FC < 0 ~ "Protein-Driven Down",
        TRUE ~ "Other"
      )
    )
  
  # --- LABELING STRATEGY ---
  # Only label if the gene is in our globally strict top 10 list
  merged_df$label <- ifelse(merged_df$Gene %in% targets_global, merged_df$Gene, "")
  
  # Plot
  p <- ggplot(merged_df, aes(x = MGH_RNA_FC, y = MGH_Prot_FC)) +
    geom_point(data = subset(merged_df, plot_category == "Other"), color = "#ebebeb", size = 2, alpha = 0.8) +
    geom_point(data = subset(merged_df, plot_category == "Protein-Driven Down"), color = "#0072B2", size = 2, alpha = 1) + 
    geom_point(data = subset(merged_df, plot_category == "Protein-Driven Up"), color = "#b34747", size = 2, alpha = 1) + 
    
    # --- HALO HIGHLIGHT FOR TARGETS ---
    # 1. Black Ring
    geom_point(data = subset(merged_df, label != ""),
               color = "black", size = 4, stroke = 1.2, shape = 21, fill = NA) + 
    # 2. Colored Center
    geom_point(data = subset(merged_df, label != ""),
               aes(color = plot_category), size = 3) +
    # 3. Labels
    geom_text_repel(aes(label = label), data = subset(merged_df, label != ""), 
                    color = "black", size = 4, fontface = "bold",
                    min.segment.length = 0, box.padding = 0.8, max.overlaps = Inf) +
    
    scale_color_manual(values = c("Protein-Driven Up" = "#b34747", "Protein-Driven Down" = "#0072B2", "Other" = "#ebebeb")) +
    labs(title = paste0("MGH ", cond_name), subtitle = "Highlighted: Top Divergent Genes (Strictly Consistent in All 12 Conditions)",
         x = "mRNA Log2 Fold Change", y = "Protein Log2 Fold Change") +
    theme_bw() +
    theme(panel.grid = element_blank(), plot.title = element_text(face = "bold"), legend.position = "none") +
    geom_vline(xintercept = 0, linetype = "dashed") + geom_hline(yintercept = 0, linetype = "dashed") +
    coord_cartesian(xlim = c(-3.5, 3.5), ylim = c(-9, 9))
  
  return(p)
}

# --- Execute ---
for (cond in conditions) {
  print(create_rna_prot_scatter(cond))
}
```





```{r}
# Filter for Bucket 1 criteria
bucket1_genes <- quadrant_data %>%
  dplyr::filter(
    # Protein is significantly UP
    Protein_padj < 0.05 & Protein_Log2FC > 0,
    # AND RNA is NOT significantly UP
    !(mRNA_padj < 0.05 & mRNA_Log2FC > 0)
  )

# Split the single dataframe into a list of 12 dataframes (one for each comparison)
bucket1_list <- bucket1_genes %>%
  dplyr::group_split(comparison, .keep = TRUE) %>%
  # Set the names of the list elements to be the comparison names
  setNames(sapply(., function(df) unique(df$comparison)))

# --- How to view the results ---

# See the names of all 12 comparisons in your list
print(names(bucket1_list))


# Extract the list of gene names from each of the 12 dataframes in bucket1_list
bucket1_gene_vectors <- lapply(bucket1_list, function(df) df$Gene)

# Iteratively find the common genes across all 12 lists
overlapping_bucket1_genes <- Reduce(intersect, bucket1_gene_vectors)

# --- View the results ---
print("Overlapping genes for Bucket 1:")
print(overlapping_bucket1_genes)
```



```{r}
# Filter for Bucket 2 criteria
bucket2_genes <- quadrant_data %>%
  dplyr::filter(
    # Protein is significantly DOWN
    Protein_padj < 0.05 & Protein_Log2FC < 0,
    # AND RNA is NOT significantly DOWN
    !(mRNA_padj < 0.05 & mRNA_Log2FC < 0)
  )

# Split into a list of 12 dataframes
bucket2_list <- bucket2_genes %>%
  dplyr::group_split(comparison, .keep = TRUE) %>%
  setNames(sapply(., function(df) unique(df$comparison)))

# --- How to view the results ---

# See the names of all 12 comparisons
print(names(bucket2_list))


# Extract the list of gene names from each of the 12 dataframes in bucket2_list
bucket2_gene_vectors <- lapply(bucket2_list, function(df) df$Gene)

# Find the common genes across all 12 lists
overlapping_bucket2_genes <- Reduce(intersect, bucket2_gene_vectors)

# --- View the results ---
print("Overlapping genes for Bucket 2:")
print(overlapping_bucket2_genes)
```


Scatter Plot v2
```{r}

```




Preparation for MotifmatchR - Bucket 1
```{r}
library(biomaRt)
library(Biostrings)
library(tidyverse)

# --- Step 1: Prepare Gene Lists (Modified) ---

# MODIFICATION 1: Use the overlapping genes for the foreground set
# This assumes you have already run the code to create 'overlapping_bucket1_genes'
foreground_genes <- overlapping_bucket1_genes

# The background set remains the same: all genes identified in the experiment
background_genes <- unique(master_integrated_table$Gene)

print(paste("Foreground set now contains", length(foreground_genes), "overlapping genes."))
print(paste("Background set contains", length(background_genes), "genes."))


# --- Step 2: Retrieve 3' UTR Sequences using biomaRt ---

# Connect to the Ensembl database for human genes
ensembl_mart <- useMart(biomart = "ensembl", dataset = "hsapiens_gene_ensembl")

# Fetch sequences for the foreground gene set
foreground_seqs <- getSequence(
  id = foreground_genes,
  type = "hgnc_symbol",
  seqType = "3utr",
  mart = ensembl_mart
)

# Fetch sequences for the background gene set
background_seqs <- getSequence(
  id = background_genes,
  type = "hgnc_symbol",
  seqType = "3utr",
  mart = ensembl_mart
)


# --- Step 3: Write Sequences to FASTA Files (Modified with Error Fix) ---

# This function is now more robust and will filter out invalid sequences
write_fasta_from_biomart <- function(seq_df, filename) {
  seq_df_clean <- seq_df %>%
    # Remove entries without a sequence
    filter(nchar(`3utr`) > 0) %>%
    # MODIFICATION 2 (ERROR FIX): Filter for valid DNA sequences only (A,C,G,T,N)
    # The regex "^[ACGTN]+$" ensures the sequence only contains DNA characters
    filter(grepl("^[ACGTN]+$", `3utr`, ignore.case = TRUE)) %>%
    # Keep only the first unique entry for each gene symbol
    distinct(hgnc_symbol, .keep_all = TRUE)

  if (nrow(seq_df_clean) > 0) {
    # Create a DNAStringSet object
    dna_string_set <- DNAStringSet(seq_df_clean$`3utr`)
    names(dna_string_set) <- seq_df_clean$hgnc_symbol

    # Write to file
    writeXStringSet(dna_string_set, filename)
    print(paste("Wrote", length(dna_string_set), "sequences to", filename))
  } else {
    print(paste("Warning: No valid sequences found to write to", filename))
  }
}

# Write the FASTA files to your working directory
write_fasta_from_biomart(foreground_seqs, "foreground_genes_overlapping_3utr.fasta")
write_fasta_from_biomart(background_seqs, "background_genes_3utr.fasta")
```

```{r}
library(biomaRt)
library(Biostrings)
library(tidyverse)

# --- Step 1: Prepare Gene Lists (Modified) ---

# MODIFICATION 1: Use the overlapping genes for the foreground set
# This assumes you have already run the code to create 'overlapping_bucket1_genes'
foreground_genes <- overlapping_bucket2_genes

# The background set remains the same: all genes identified in the experiment
background_genes <- unique(master_integrated_table$Gene)

print(paste("Foreground set now contains", length(foreground_genes), "overlapping genes."))
print(paste("Background set contains", length(background_genes), "genes."))


# --- Step 2: Retrieve 3' UTR Sequences using biomaRt ---

# Connect to the Ensembl database for human genes
ensembl_mart <- useMart(biomart = "ensembl", dataset = "hsapiens_gene_ensembl")

# Fetch sequences for the foreground gene set
foreground_seqs <- getSequence(
  id = foreground_genes,
  type = "hgnc_symbol",
  seqType = "3utr",
  mart = ensembl_mart
)

# Fetch sequences for the background gene set
background_seqs <- getSequence(
  id = background_genes,
  type = "hgnc_symbol",
  seqType = "3utr",
  mart = ensembl_mart
)


# --- Step 3: Write Sequences to FASTA Files (Modified with Error Fix) ---

# This function is now more robust and will filter out invalid sequences
write_fasta_from_biomart <- function(seq_df, filename) {
  seq_df_clean <- seq_df %>%
    # Remove entries without a sequence
    filter(nchar(`3utr`) > 0) %>%
    # MODIFICATION 2 (ERROR FIX): Filter for valid DNA sequences only (A,C,G,T,N)
    # The regex "^[ACGTN]+$" ensures the sequence only contains DNA characters
    filter(grepl("^[ACGTN]+$", `3utr`, ignore.case = TRUE)) %>%
    # Keep only the first unique entry for each gene symbol
    distinct(hgnc_symbol, .keep_all = TRUE)

  if (nrow(seq_df_clean) > 0) {
    # Create a DNAStringSet object
    dna_string_set <- DNAStringSet(seq_df_clean$`3utr`)
    names(dna_string_set) <- seq_df_clean$hgnc_symbol

    # Write to file
    writeXStringSet(dna_string_set, filename)
    print(paste("Wrote", length(dna_string_set), "sequences to", filename))
  } else {
    print(paste("Warning: No valid sequences found to write to", filename))
  }
}

# Write the FASTA files to your working directory
write_fasta_from_biomart(foreground_seqs, "2foreground_genes_overlapping_3utr.fasta")
write_fasta_from_biomart(background_seqs, "2background_genes_3utr.fasta")
```



MotifMatchR Foreground Bucket 1
```{r}
# --- Step 1: Recreate the Quadrant Data ---
# This step is needed to get the classification and p-values.
quadrant_data <- master_integrated_table %>%
  mutate(
    quadrant = case_when(
      mRNA_Log2FC > 0 & Protein_Log2FC > 0   ~ "Up-Up (Correlated)",
      mRNA_Log2FC < 0 & Protein_Log2FC < 0   ~ "Down-Down (Correlated)",
      mRNA_Log2FC > 0 & Protein_Log2FC < 0   ~ "Up-Down (Post-transcriptional Down-regulation)",
      mRNA_Log2FC < 0 & Protein_Log2FC > 0   ~ "Down-Up (Post-transcriptional Up-regulation)",
      TRUE                                   ~ "No Change"
    )
  )

# --- Step 2: Recreate Bucket 1 ---
# Filter for genes significantly UP in protein but NOT significantly UP in mRNA.
bucket1_genes <- quadrant_data %>%
  dplyr::filter(
    Protein_padj < 0.05 & Protein_Log2FC > 0,
    !(mRNA_padj < 0.05 & mRNA_Log2FC > 0)
  )

# Split into a list of dataframes, one for each condition.
bucket1_list <- bucket1_genes %>%
  dplyr::group_split(comparison, .keep = TRUE) %>%
  setNames(sapply(., function(df) unique(df$comparison)))

# --- Step 3: Find Overlapping Genes ---
# Extract the gene names from each of the 12 dataframes.
bucket1_gene_vectors <- lapply(bucket1_list, function(df) df$Gene)

# Find the intersection of all 12 gene lists.
overlapping_bucket1_genes <- Reduce(intersect, bucket1_gene_vectors)

# --- Step 4: Recreate the final foreground_genes list ---
foreground_genes_bucket1 <- overlapping_bucket1_genes

# --- Step 5: Verify the result ---
print("foreground_genes list has been recreated:")
print(paste("Number of overlapping genes:", length(foreground_genes_bucket1)))
print(head(foreground_genes_bucket1))
```


MotifMatchR Foreground Bucket 2
```{r}
# --- Step 1: Recreate Bucket 2 ---
# Filter for genes significantly DOWN in protein but NOT significantly DOWN in mRNA.
bucket2_genes <- quadrant_data %>%
  dplyr::filter(
    Protein_padj < 0.05 & Protein_Log2FC < 0,
    !(mRNA_padj < 0.05 & mRNA_Log2FC < 0)
  )

# Split into a list of dataframes, one for each condition.
bucket2_list <- bucket2_genes %>%
  dplyr::group_split(comparison, .keep = TRUE) %>%
  setNames(sapply(., function(df) unique(df$comparison)))

# --- Step 2: Find Overlapping Genes ---
# Extract the gene names from each of the 12 dataframes.
bucket2_gene_vectors <- lapply(bucket2_list, function(df) df$Gene)

# Find the intersection of all 12 gene lists.
overlapping_bucket2_genes <- Reduce(intersect, bucket2_gene_vectors)

# --- Step 3: Create the final foreground_genes list for Bucket 2 ---
foreground_genes_bucket2 <- overlapping_bucket2_genes

# --- Step 4: Verify the result ---
print("foreground_genes list for Bucket 2 has been recreated:")
print(paste("Number of overlapping genes:", length(foreground_genes_bucket2)))
print(head(foreground_genes_bucket2))
```


MotifMatchR Filtering for RNA-binding proteins
```{r}
# --- Step 1: Load Libraries ---
library(MotifDb)
library(dplyr)
library(tibble)
library(data.table)

# --- Step 2: Read RBP Data and Filter by Score ---
file_path <- "Table_HS_RBP.txt" # Assumes the RBP2GO file is in your directory
rbp_data <- data.table::fread(file_path)

# --- SET YOUR CONFIDENCE THRESHOLD HERE ---
score_cutoff <- 10

# Filter for high-confidence RBPs using the RBP2GO_Score
high_confidence_rbps <- rbp_data %>%
  dplyr::filter(RBP2GO_Score >= score_cutoff) %>%
  dplyr::pull(Gene_Name) %>%
  unique()

message(paste("Filtered to", length(high_confidence_rbps), "high-confidence RBPs with a score >=", score_cutoff))


# --- Step 3: Filter MotifDb with the High-Confidence List ---

# Get the full metadata from your MotifDb installation
all_metadata <- as.data.frame(values(MotifDb)) %>%
  tibble::rownames_to_column(var = "motif")

# Filter the metadata to find rows where the geneSymbol is in our new high-confidence list
rbp_metadata_subset <- all_metadata %>%
  dplyr::filter(geneSymbol %in% high_confidence_rbps & organism == "Hsapiens")

# Get the unique names of the motifs that passed the filter
rbp_motif_names <- rbp_metadata_subset %>%
  dplyr::pull(motif) %>%
  unique()


# --- Step 4: Create the New High-Confidence RBP-only MotifDb Object ---

# Subset the main MotifDb object using our filtered list of motif names
rbp_only_motifs <- MotifDb[rbp_motif_names]


# --- Step 5: Verify the Result ---
message(paste("Successfully created a new database with", length(rbp_only_motifs), "motifs for the high-confidence RBPs."))
print(rbp_only_motifs)
length(rbp_only_motifs)
```

MotifMatchR enrichment for Bucket 1
```{r}
# --- Step 1: Load Libraries ---
library(motifmatchr)
library(MotifDb)
library(GenomicRanges)
library(BSgenome.Hsapiens.UCSC.hg38)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(org.Hs.eg.db)
library(TFBSTools)
library(SummarizedExperiment)
library(data.table)
library(tidyverse)

# --- Step 3: Prepare Sequences and Gene Lists ---
message("Fetching 3' UTRs and mapping to gene symbols...")
utrs <- threeUTRsByTranscript(TxDb.Hsapiens.UCSC.hg38.knownGene, use.names = TRUE)
tx_names <- names(utrs)

tx2geneid <- AnnotationDbi::select(
  TxDb.Hsapiens.UCSC.hg38.knownGene, keys = tx_names,
  keytype = "TXNAME", columns = "GENEID"
)
geneid2symbol <- AnnotationDbi::select(
  org.Hs.eg.db, keys = unique(tx2geneid$GENEID),
  keytype = "ENTREZID", columns = "SYMBOL"
)
tx2symbol <- dplyr::left_join(tx2geneid, geneid2symbol, by = c("GENEID" = "ENTREZID")) %>%
  dplyr::filter(!is.na(SYMBOL) & !is.na(TXNAME))

foreground_tx_names <- tx2symbol %>%
  dplyr::filter(SYMBOL %in% foreground_genes_bucket1) %>%
  dplyr::pull(TXNAME) %>% unique()

background_tx_names <- tx2symbol %>%
  dplyr::filter(SYMBOL %in% background_genes) %>%
  dplyr::pull(TXNAME) %>% unique()

if (length(foreground_tx_names) == 0) {
  stop("Error: None of the foreground gene symbols could be mapped to a transcript.")
}
message("Extracting genomic sequences...")
utrs_foreground <- utrs[names(utrs) %in% foreground_tx_names]
utrs_background <- utrs[names(utrs) %in% background_tx_names]
seqs_foreground <- getSeq(BSgenome.Hsapiens.UCSC.hg38, utrs_foreground)
seqs_background <- getSeq(BSgenome.Hsapiens.UCSC.hg38, utrs_background)


# --- Step 4: Create and Prepare the HIGH-CONFIDENCE RBP-only Motif Database ---
message("Creating a focused database of high-confidence RBP-only motifs...")

# Define the path to your RBP2GO data file
file_path <- "Table_HS_RBP.txt" # Make sure this file is in your directory
rbp_data <- data.table::fread(file_path, skip = 5)

# --- SET YOUR CONFIDENCE THRESHOLD HERE ---
score_cutoff <- 10

# **NEW:** Filter for high-confidence RBPs using the RBP2GO_Score
high_confidence_rbp_genes <- rbp_data %>%
  dplyr::filter(RBP2GO_Score >= score_cutoff) %>%
  dplyr::pull(Gene_Name) %>%
  unique()

message(paste("Filtered to", length(high_confidence_rbp_genes), "high-confidence RBPs with a score >=", score_cutoff))

# Get the full metadata from your MotifDb installation
all_metadata <- as.data.frame(values(MotifDb)) %>%
  tibble::rownames_to_column(var = "motif")

# **UPDATED:** Filter the metadata using the new high_confidence_rbp_genes list
rbp_metadata_subset <- all_metadata %>%
  dplyr::filter(geneSymbol %in% high_confidence_rbp_genes & organism == "Hsapiens")

# Get the unique names of the motifs that passed the filter
rbp_motif_names <- rbp_metadata_subset %>%
  dplyr::pull(motif) %>%
  unique()

# Subset the main MotifDb object using our filtered list of motif names
rbp_only_motifs <- MotifDb[rbp_motif_names]

message("Preparing high-confidence RBP motifs for matching...")
motif_list_raw <- as.list(rbp_only_motifs)
pfm_object_list <- mapply(
    function(matrix, name) { TFBSTools::PFMatrix(ID = name, name = name, profileMatrix = matrix) },
    motif_list_raw, names(motif_list_raw), SIMPLIFY = FALSE
)
rbp_motifs_final <- do.call(PFMatrixList, pfm_object_list)
message(paste("Successfully prepared", length(rbp_motifs_final), "high-confidence RBP motifs for analysis."))


# --- Step 5: Find Motif Matches ---
message("Scanning sequences for motif matches...")
seqs_foreground_flat <- unlist(seqs_foreground)
seqs_background_flat <- unlist(seqs_background)

motif_results_foreground <- motifmatchr::matchMotifs(rbp_motifs_final, seqs_foreground_flat, out = "scores")
motif_results_background <- motifmatchr::matchMotifs(rbp_motifs_final, seqs_background_flat, out = "scores")

motif_hits_foreground <- motifCounts(motif_results_foreground)
motif_hits_background <- motifCounts(motif_results_background)


# --- Step 6: Perform Enrichment Analysis for Bucket 1 ---
message("Performing Fisher's exact test for enrichment...")
final_results_bucket1 <- tibble() 

if (nrow(motif_hits_foreground) > 0) {
    foreground_hits <- colSums(motif_hits_foreground > 0)
    background_hits <- colSums(motif_hits_background > 0)
    
    num_foreground_seqs <- nrow(motif_hits_foreground)
    num_background_seqs <- nrow(motif_hits_background)

    fisher_results <- lapply(1:length(rbp_motifs_final), function(i) {
        tbl <- matrix(c(
            foreground_hits[i],
            background_hits[i],
            num_foreground_seqs - foreground_hits[i],
            num_background_seqs - background_hits[i]
        ), nrow = 2)
        
        fisher.test(tbl, alternative = "greater")
    })

    final_results_bucket1 <- tibble(
        motif = names(rbp_motifs_final),
        p_value = sapply(fisher_results, "[[", "p.value"),
        odds_ratio = sapply(fisher_results, function(x) x$estimate)
    ) %>%
    dplyr::mutate(p_adj = p.adjust(p_value, method = "BH")) %>%
    dplyr::arrange(p_adj)
} else {
    warning("No foreground sequences remained after mapping and filtering.")
}

print("Top 20 enriched RBP motifs for Bucket 1:")
print(head(final_results_bucket1, 20))


# --- Step 7: Visualize the Results for Bucket 1 ---

if (nrow(final_results_bucket1) > 0) {
  message("Generating final results dot plot for Bucket 1...")
  
  top_motifs_plot_bucket1 <- final_results_bucket1 %>%
    # Data preparation steps
    dplyr::filter(is.finite(odds_ratio) & p_adj < 0.05) %>%
    dplyr::top_n(15, -p_adj) %>%
    dplyr::mutate(
      log10_padj = -log10(p_adj),
      # --- CHANGE HERE ---
      # The y-axis (motif) is now sorted by odds_ratio
      motif = fct_reorder(motif, odds_ratio) 
    ) %>%
    
    # ggplot call remains the same
    ggplot(aes(x = odds_ratio, y = motif, size = odds_ratio, color = log10_padj)) +
    geom_point(alpha = 0.8) +
    
    # Scales remain the same
    scale_color_gradient(low = "lightblue", high = "firebrick") +
    scale_size_continuous(range = c(4, 10), breaks = c(3, 3.5, 4, 4.5)) +
    
    # Labs remain the same
    labs(
      title = "Top 15 Enriched RBP Motifs in Foreground Genes (Bucket 1)",
      x = "Odds Ratio",
      y = "RBP Motif",
      size = "Odds Ratio",
      color = bquote(-Log[10]~"(p-adj)")
    ) +
    
    # Theme remains the same
    theme_minimal() +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
    )

  print(top_motifs_plot_bucket1)
  
} else {
  message("No significant motifs found to plot for Bucket 1.")
}

# --- Step 8: Link Significant Motifs to RBP Gene Symbols for Bucket 1 ---
message("--- Linking enriched motifs to their corresponding RBPs for Bucket 1 ---")
significant_motifs_bucket1 <- final_results_bucket1 %>%
  dplyr::filter(p_adj < 0.05) %>%
  dplyr::pull(motif)

if (length(significant_motifs_bucket1) > 0) {
    rbp_identities_bucket1 <- as.data.frame(values(MotifDb)) %>%
      tibble::rownames_to_column(var = "motif") %>%
      dplyr::filter(motif %in% significant_motifs_bucket1) %>%
      dplyr::select(motif, geneSymbol, proteinId, dataSource) %>%
      dplyr::arrange(match(motif, significant_motifs_bucket1))

    print(head(rbp_identities_bucket1, 20))
} else {
    message("No significantly enriched motifs to report for Bucket 1.")
}

# --- Step 9: Create and Print Final Summary Table for Bucket 1 ---
# Make sure the knitr package is installed: install.packages("knitr")
library(knitr)

message("--- Final Summary of Enriched RBP Motifs (p_adj < 0.05) ---")

if (exists("rbp_identities_bucket1") && nrow(rbp_identities_bucket1) > 0) {
  
  # Join the statistical results with the RBP identities
  summary_table_bucket1 <- dplyr::left_join(
    final_results_bucket1, 
    rbp_identities_bucket1, 
    by = "motif"
  ) %>%
  # Filter for significant results
  dplyr::filter(p_adj < 0.05) %>%
  # Select and order the most important columns
  dplyr::select(motif, geneSymbol, odds_ratio, p_value, p_adj) %>%
  # Arrange by significance
  dplyr::arrange(p_adj)

  # Print the final results in a clean table format
  if(nrow(summary_table_bucket1) > 0){
    print(knitr::kable(summary_table_bucket1, digits = 3, caption = "Enriched RBP Motifs (Bucket 1)"))
  } else {
    message("No motifs passed the significance threshold of p_adj < 0.05.")
  }
  
} else {
  message("No significantly enriched motifs were found to create a summary table.")
}

# --- Step 10: Overlap Enriched Motifs with a Specific Protein List ---

message("--- Overlapping Enriched Motifs with Your Protein List ---")

# First, make sure the summary table from Step 9 exists and has results
if (exists("summary_table_bucket1") && nrow(summary_table_bucket1) > 0) {
  
  # Filter the summary table to find rows where the geneSymbol is in your protein_list
  overlapping_results <- summary_table_bucket1 %>%
    dplyr::filter(geneSymbol %in% protein_list)
  
  # Print the results of the overlap
  if(nrow(overlapping_results) > 0){
    print(knitr::kable(
      overlapping_results, 
      digits = 3, 
      caption = "Enriched Motifs Overlapping with Your Protein List"
    ))
  } else {
    message("No overlap found between the significantly enriched motifs and your protein list.")
  }
  
} else {
  message("No significant motifs were found in Step 9 to overlap with.")
}


# --- Step 11: Create a Summarized Dot Plot (Unique RBPs) ---
message("--- Generating a summarized plot with unique RBPs for Bucket 2 ---")

# First, make sure the final results and identities tables exist and are not empty
if (exists("final_results_bucket1") && exists("rbp_identities_bucket2") && nrow(final_results_bucket1) > 0) {

  # 1. Join the statistical results with the RBP gene symbol identities
  enriched_rbps_all <- dplyr::left_join(
    final_results_bucket1,
    rbp_identities_bucket1,
    by = "motif"
  ) %>%
  # Filter for significant results and remove any RBPs that couldn't be mapped to a gene symbol
  dplyr::filter(p_adj < 0.05 & !is.na(geneSymbol) & is.finite(odds_ratio))

  # 2. For each RBP (geneSymbol), find its single best result (lowest p-value)
  enriched_rbps_unique <- enriched_rbps_all %>%
    dplyr::group_by(geneSymbol) %>%
    # slice_min() keeps the row with the minimum p_adj for each geneSymbol group
    dplyr::slice_min(order_by = p_adj, n = 1, with_ties = FALSE) %>%
    ungroup() # Ungroup for further operations

  # 3. Create the dot plot using this summarized data
  if(nrow(enriched_rbps_unique) > 0) {
    summarized_dot_plot <- enriched_rbps_unique %>%
      dplyr::top_n(15, -p_adj) %>%
      dplyr::mutate(
        log10_padj = -log10(p_adj),
        # Reorder the y-axis by the odds ratio for better visualization
        geneSymbol = fct_reorder(geneSymbol, odds_ratio)
      ) %>%
      ggplot(aes(x = odds_ratio, y = geneSymbol, size = odds_ratio, color = log10_padj)) +
      geom_point(alpha = 0.8) +
      scale_color_gradient(low = "lightblue", high = "firebrick") +
      scale_size_continuous(range = c(4, 10)) + # Adjust range as needed
      labs(
        title = "Top 15 Enriched RBPs (Unique) in Foreground Genes (Bucket 1)",
        x = "Odds Ratio",
        y = "RBP (Gene Symbol)",
        size = "Odds Ratio",
        color = bquote(-Log[10]~"(p-adj)")
      ) +
      theme_minimal() +
      theme(panel.grid = element_blank())

    print(summarized_dot_plot)

  } else {
    message("No significant motifs remained after summarizing by gene symbol for Bucket 2.")
  }

} else {
  message("No significant motifs found in the previous step to create a summarized plot for Bucket 2.")
}
```


MotifMatchR enrichment for Bucket 2
```{r}
# --- Step 1: Load Libraries ---
library(motifmatchr)
library(MotifDb)
library(GenomicRanges)
library(BSgenome.Hsapiens.UCSC.hg38)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(org.Hs.eg.db)
library(TFBSTools)
library(SummarizedExperiment)
library(data.table)
library(knitr)
library(tidyverse)


# --- Step 3: Prepare Sequences and Gene Lists ---
message("Fetching 3' UTRs and mapping to gene symbols...")
utrs <- threeUTRsByTranscript(TxDb.Hsapiens.UCSC.hg38.knownGene, use.names = TRUE)
tx_names <- names(utrs)

tx2geneid <- AnnotationDbi::select(
  TxDb.Hsapiens.UCSC.hg38.knownGene, keys = tx_names,
  keytype = "TXNAME", columns = "GENEID"
)
geneid2symbol <- AnnotationDbi::select(
  org.Hs.eg.db, keys = unique(tx2geneid$GENEID),
  keytype = "ENTREZID", columns = "SYMBOL"
)
tx2symbol <- dplyr::left_join(tx2geneid, geneid2symbol, by = c("GENEID" = "ENTREZID")) %>%
  dplyr::filter(!is.na(SYMBOL) & !is.na(TXNAME))

foreground_tx_names <- tx2symbol %>%
  dplyr::filter(SYMBOL %in% foreground_genes_bucket2) %>%
  dplyr::pull(TXNAME) %>% unique()

background_tx_names <- tx2symbol %>%
  dplyr::filter(SYMBOL %in% background_genes) %>%
  dplyr::pull(TXNAME) %>% unique()

if (length(foreground_tx_names) == 0) {
  stop("Error: None of the foreground gene symbols could be mapped to a transcript.")
}
message("Extracting genomic sequences...")
utrs_foreground <- utrs[names(utrs) %in% foreground_tx_names]
utrs_background <- utrs[names(utrs) %in% background_tx_names]
seqs_foreground <- getSeq(BSgenome.Hsapiens.UCSC.hg38, utrs_foreground)
seqs_background <- getSeq(BSgenome.Hsapiens.UCSC.hg38, utrs_background)



# --- Step 5: Find Motif Matches ---
message("Scanning sequences for motif matches...")
seqs_foreground_flat <- unlist(seqs_foreground)
seqs_background_flat <- unlist(seqs_background)

motif_results_foreground <- motifmatchr::matchMotifs(rbp_motifs_final, seqs_foreground_flat, out = "scores")
motif_results_background <- motifmatchr::matchMotifs(rbp_motifs_final, seqs_background_flat, out = "scores")

motif_hits_foreground <- motifCounts(motif_results_foreground)
motif_hits_background <- motifCounts(motif_results_background)


# --- Step 6: Perform Enrichment Analysis for Bucket 2 ---
message("Performing Fisher's exact test for enrichment...")
final_results_bucket2 <- tibble() 

if (nrow(motif_hits_foreground) > 0) {
    foreground_hits <- colSums(motif_hits_foreground > 0)
    background_hits <- colSums(motif_hits_background > 0)
    
    num_foreground_seqs <- nrow(motif_hits_foreground)
    num_background_seqs <- nrow(motif_hits_background)

    fisher_results <- lapply(1:length(rbp_motifs_final), function(i) {
        tbl <- matrix(c(
            foreground_hits[i],
            background_hits[i],
            num_foreground_seqs - foreground_hits[i],
            num_background_seqs - background_hits[i]
        ), nrow = 2)
        
        fisher.test(tbl, alternative = "greater")
    })

    final_results_bucket2 <- tibble(
        motif = names(rbp_motifs_final),
        p_value = sapply(fisher_results, "[[", "p.value"),
        odds_ratio = sapply(fisher_results, function(x) x$estimate)
    ) %>%
    dplyr::mutate(p_adj = p.adjust(p_value, method = "BH")) %>%
    dplyr::arrange(p_adj)
} else {
    warning("No foreground sequences remained after mapping and filtering.")
}

print("Top 20 enriched RBP motifs for Bucket 2:")
print(head(final_results_bucket2, 20))


# --- Step 7: Visualize the Results for Bucket 2 ---
# Load necessary libraries if you haven't already
# library(ggplot2)
# library(dplyr)
# library(forcats)

if (nrow(final_results_bucket2) > 0) {
  message("Generating updated results dot plot for Bucket 2...")
  
  top_motifs_plot_bucket2 <- final_results_bucket2 %>%
    # Data preparation steps
    dplyr::filter(is.finite(odds_ratio) & p_adj < 0.05) %>%
    dplyr::top_n(15, -p_adj) %>%
    dplyr::mutate(
      log10_padj = -log10(p_adj),
      # Sort the y-axis (motif) by the odds_ratio
      motif = fct_reorder(motif, odds_ratio) 
    ) %>%
    
    # ggplot call with updated aesthetics
    ggplot(aes(x = odds_ratio, y = motif, size = odds_ratio, color = log10_padj)) +
    geom_point(alpha = 0.8) +
    
    # Set the color gradient for significance
    scale_color_gradient(low = "lightblue", high = "firebrick") +
    
    # Set the size range and legend breaks
    # NOTE: You may need to adjust the numbers in `breaks` to fit Bucket 2's data
    scale_size_continuous(range = c(4, 10), breaks = c(3, 3.5, 4, 4.5)) +
    
    # Update labels for the new axes and legends
    labs(
      title = "Top 15 Enriched RBP Motifs in Foreground Genes (Bucket 2)",
      x = "Odds Ratio",
      y = "RBP Motif",
      size = "Odds Ratio",
      color = bquote(-Log[10]~"(p-adj)")
    ) +
    
    # Apply a minimal theme and remove gridlines
    theme_minimal() +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
    )

  print(top_motifs_plot_bucket2)
  
} else {
  message("No significant motifs found to plot for Bucket 2.")
}

# --- Step 8: Link Significant Motifs to RBP Gene Symbols for Bucket 2 ---
message("--- Linking enriched motifs to their corresponding RBPs for Bucket 2 ---")
significant_motifs_bucket2 <- final_results_bucket2 %>%
  dplyr::filter(p_adj < 0.05) %>%
  dplyr::pull(motif)

if (length(significant_motifs_bucket2) > 0) {
    rbp_identities_bucket2 <- as.data.frame(values(MotifDb)) %>%
      tibble::rownames_to_column(var = "motif") %>%
      dplyr::filter(motif %in% significant_motifs_bucket2) %>%
      dplyr::select(motif, geneSymbol, proteinId, dataSource) %>%
      dplyr::arrange(match(motif, significant_motifs_bucket2))

    print(head(rbp_identities_bucket2, 20))
} else {
    message("No significantly enriched motifs to report for Bucket 2.")
}


# --- Step 9: Create and Print Final Summary Table for Bucket 2 ---
message("--- Final Summary of Enriched RBP Motifs (p_adj < 0.05) for Bucket 2 ---")
if (exists("rbp_identities_bucket2") && nrow(rbp_identities_bucket2) > 0) {
  
  summary_table_bucket2 <- dplyr::left_join(
    final_results_bucket2, 
    rbp_identities_bucket2, 
    by = "motif"
  ) %>%
  dplyr::filter(p_adj < 0.05) %>%
  dplyr::select(motif, geneSymbol, odds_ratio, p_value, p_adj) %>%
  dplyr::arrange(p_adj)

  if(nrow(summary_table_bucket2) > 0){
    print(knitr::kable(summary_table_bucket2, digits = 3, caption = "Enriched RBP Motifs (Bucket 2)"))
  } else {
    message("No motifs passed the significance threshold of p_adj < 0.05.")
  }
  
} else {
  message("No significantly enriched motifs were found to create a summary table.")
}


# --- Step 10: Overlap Enriched Motifs with a Specific Protein List for Bucket 2 ---
message("--- Overlapping Enriched Motifs with Your Protein List for Bucket 2 ---")
if (exists("summary_table_bucket2") && nrow(summary_table_bucket2) > 0) {
  
  overlapping_results_bucket2 <- summary_table_bucket2 %>%
    dplyr::filter(geneSymbol %in% protein_list)
  
  if(nrow(overlapping_results_bucket2) > 0){
    print(knitr::kable(
      overlapping_results_bucket2, 
      digits = 3, 
      caption = "Enriched Motifs (Bucket 2) Overlapping with Your Protein List"
    ))
  } else {
    message("No overlap found between the significantly enriched motifs and your protein list.")
  }
  
} else {
  message("No significant motifs were found in Step 9 to overlap with.")
}


# --- Step 11: Create a Summarized Dot Plot (Unique RBPs) ---
message("--- Generating a summarized plot with unique RBPs for Bucket 2 ---")

# First, make sure the final results and identities tables exist and are not empty
if (exists("final_results_bucket2") && exists("rbp_identities_bucket2") && nrow(final_results_bucket2) > 0) {

  # 1. Join the statistical results with the RBP gene symbol identities
  enriched_rbps_all <- dplyr::left_join(
    final_results_bucket2,
    rbp_identities_bucket2,
    by = "motif"
  ) %>%
  # Filter for significant results and remove any RBPs that couldn't be mapped to a gene symbol
  dplyr::filter(p_adj < 0.05 & !is.na(geneSymbol) & is.finite(odds_ratio))

  # 2. For each RBP (geneSymbol), find its single best result (lowest p-value)
  enriched_rbps_unique <- enriched_rbps_all %>%
    dplyr::group_by(geneSymbol) %>%
    # slice_min() keeps the row with the minimum p_adj for each geneSymbol group
    dplyr::slice_min(order_by = p_adj, n = 1, with_ties = FALSE) %>%
    ungroup() # Ungroup for further operations

  # 3. Create the dot plot using this summarized data
  if(nrow(enriched_rbps_unique) > 0) {
    summarized_dot_plot <- enriched_rbps_unique %>%
      dplyr::top_n(15, -p_adj) %>%
      dplyr::mutate(
        log10_padj = -log10(p_adj),
        # Reorder the y-axis by the odds ratio for better visualization
        geneSymbol = fct_reorder(geneSymbol, odds_ratio)
      ) %>%
      ggplot(aes(x = odds_ratio, y = geneSymbol, size = odds_ratio, color = log10_padj)) +
      geom_point(alpha = 0.8) +
      scale_color_gradient(low = "lightblue", high = "firebrick") +
      scale_size_continuous(range = c(4, 10)) + # Adjust range as needed
      labs(
        title = "Top 15 Enriched RBPs (Unique) in Foreground Genes (Bucket 2)",
        x = "Odds Ratio",
        y = "RBP (Gene Symbol)",
        size = "Odds Ratio",
        color = bquote(-Log[10]~"(p-adj)")
      ) +
      theme_minimal() +
      theme(panel.grid = element_blank())

    print(summarized_dot_plot)

  } else {
    message("No significant motifs remained after summarizing by gene symbol for Bucket 2.")
  }

} else {
  message("No significant motifs found in the previous step to create a summarized plot for Bucket 2.")
}

#step 12



# --- Step 2: Load Libraries and Prepare Data ---
# Make sure the VennDiagram package is installed: install.packages("VennDiagram")
library(VennDiagram)
library(dplyr)

message("--- Comparing Enriched RBPs (Bucket 2) with Your Protein List ---")

# First, ensure the summary table from the previous step exists and is not empty
if (exists("summary_table_bucket2") && nrow(summary_table_bucket2) > 0) {

  # 1. Extract the unique gene symbols of significantly enriched RBPs from Bucket 2
  enriched_rbps_bucket2 <- summary_table_bucket2 %>%
    dplyr::filter(p_adj < 0.05) %>%
    dplyr::pull(geneSymbol) %>%
    unique()

  # 2. Prepare the data for the Venn diagram
  venn_list_overlap <- list(
    `Enriched RBPs (Bucket 2)` = enriched_rbps_bucket2,
    `Your Protein List` = protein_list
  )

  # 3. Create and draw the Venn diagram
  grid.newpage() # Clears any previous plots

  venn.plot <- venn.diagram(
    x = venn_list_overlap,
    filename = NULL, # This makes the plot appear in RStudio's plot pane
    output = TRUE,

    # --- Style Options ---
    imagetype = "png",
    col = c("#00BFC4", "#7CAE00"), # Colors for the circles
    fill = c(alpha("#00BFC4", 0.3), alpha("#7CAE00", 0.3)),
    cex = 1.5, # Size of the numbers in the circles
    fontfamily = "sans",
    cat.cex = 1.2, # Size of the category labels
    cat.fontfamily = "sans",
    cat.col = c("#00BFC4", "#7CAE00")
  )

  grid.draw(venn.plot)

  # 4. List the common and unique proteins for clarity
  common_proteins <- intersect(enriched_rbps_bucket2, protein_list)
  unique_to_enriched <- setdiff(enriched_rbps_bucket2, protein_list)
  unique_to_your_list <- setdiff(protein_list, enriched_rbps_bucket2)

  cat("\n\n--- Summary of Overlap ---\n")
  cat("\nProteins common to BOTH lists:", paste(common_proteins, collapse=", "), "\n")
  cat("\nEnriched RBPs NOT in your list:", paste(head(unique_to_enriched), collapse=", "), "...\n")
  cat("\nProteins from your list that were NOT enriched:", paste(unique_to_your_list, collapse=", "), "\n")

} else {
  message("No significant motifs were found in the previous step to create a Venn diagram.")
}



message("--- Comparing Enriched RBPs (Bucket 2) with Your Protein List ---")

# First, ensure the necessary summary tables exist
if (exists("summary_table_bucket2") && exists("rbp_only_motifs") && nrow(summary_table_bucket2) > 0) {

  # 1. Extract the list of significantly enriched RBPs from Bucket 2
  enriched_rbps_bucket2 <- summary_table_bucket2 %>%
    dplyr::filter(p_adj < 0.05) %>%
    dplyr::pull(geneSymbol) %>%
    unique()

  # 2. **CORRECTED UNIVERSE:** Define the total pool of proteins as ONLY those RBPs
  # for which we had a motif in our filtered database.
  total_rbp_universe <- unique(values(rbp_only_motifs)$geneSymbol)
  message(paste("Statistical universe defined as the", length(total_rbp_universe), "RBPs with motifs in our database."))

  # 3. Prepare the data for the Venn diagram
  venn_list_overlap <- list(
    `Enriched RBPs (Bucket 2)` = enriched_rbps_bucket2,
    `Your Protein List` = protein_list
  )
  grid.newpage()
  venn.plot <- venn.diagram(x = venn_list_overlap, filename = NULL, output = TRUE,
                            col = c("#00BFC4", "#7CAE00"), fill = c(alpha("#00BFC4", 0.3), alpha("#7CAE00", 0.3)))
  grid.draw(venn.plot)

  # 4. Find the overlapping proteins
  common_proteins <- intersect(enriched_rbps_bucket2, protein_list)

  # 5. Perform the Hypergeometric Test with the CORRECT universe
  message("\n--- Calculating Statistical Significance of Overlap ---")

  q <- length(common_proteins)
  m <- length(enriched_rbps_bucket2)
  k <- length(intersect(protein_list, total_rbp_universe)) # Your list, but only members present in the universe
  n <- length(total_rbp_universe) - m

  # Calculate the p-value
  p_value <- phyper(q - 1, m, n, k, lower.tail = FALSE)

  cat("Overlap size:", q, "\n")
  cat("Hypergeometric test p-value:", format.pval(p_value, digits = 3), "\n")
  if (p_value < 0.05) {
    cat("Result: The overlap is STATISTICALLY SIGNIFICANT.\n")
  } else {
    cat("Result: The overlap is NOT statistically significant.\n")
  }

} else {
  message("No significant motifs were found to perform the overlap test.")
}
```
Overlap with MAGEL2 interaction partner list filtered for RBPs with known motifs
```{r}
# --- Step 12: Compare Enriched RBPs (Bucket 2) with Your FILTERED Protein List ---

# --- MODIFIED LIBRARIES ---
# Make sure the eulerr package is installed: install.packages("eulerr")
library(eulerr)     # For plotting area-proportional Euler diagrams
library(grid)       # For drawing the plot grob
library(dplyr)

message("--- Comparing Enriched RBPs (Bucket 2) with Your FILTERED & ALIGNED Protein List ---")

# First, ensure all necessary data objects exist
if (exists("summary_table_bucket2") && 
    exists("rbp_only_motifs") && 
    exists("protein_list") && 
    exists("rbp_data") &&
    nrow(summary_table_bucket2) > 0) {

  # 1. Define the Statistical Universe:
  # ONLY those RBPs for which we had a motif in our filtered database
  total_rbp_universe <- unique(values(rbp_only_motifs)$geneSymbol)
  message(paste("Statistical universe defined as the", length(total_rbp_universe), "RBPs with motifs in our database."))

  # 2. Define List A: Enriched RBPs
  # These are already a subset of our universe by definition
  enriched_rbps_bucket2 <- summary_table_bucket2 %>%
    dplyr::filter(p_adj < 0.05) %>%
    dplyr::pull(geneSymbol) %>%
    unique()

  # 3. Define List B: Your Filtered & Aligned Protein List
  
  # First, get all high-confidence RBPs from the score table
  high_conf_rbps_from_table <- rbp_data %>%
    dplyr::filter(RBP2GO_Score >= 10) %>%
    dplyr::pull(Gene_Name) %>%
    unique()
  
  # Filter your list by score
  protein_list_high_conf <- intersect(protein_list, high_conf_rbps_from_table)
  
  # **NEW/UPDATED STEP:** Filter this high-conf list *again* by the universe
  # This ensures List B only contains genes that were *possible* to find
  protein_list_final <- intersect(protein_list_high_conf, total_rbp_universe)
  
  message(paste("Original 'protein_list' had", length(protein_list), "genes."))
  message(paste("After filtering for RBP2GO_Score >= 10, it has", length(protein_list_high_conf), "genes."))
  message(paste("After *also* filtering for presence in motif universe, your final list (List B) has", length(protein_list_final), "genes."))


  # 4. Prepare the data for the Euler diagram
  venn_list_overlap <- list(
    `Enriched RBPs (List A)` = enriched_rbps_bucket2,
    `Your Aligned List (List B)` = protein_list_final 
  )
  
  # --- MODIFIED PLOTTING STEP ---
  # 5. Create and draw the Euler diagram
  # This replaces the old VennDiagram code
  message("\n--- Generating Euler Diagram ---")
  grid.newpage() # Clear the plot device
  
  # Create the euler plot object using the style from your example
  # quantities = TRUE adds the numbers to the segments
  euler_plot_grob <- plot(euler(venn_list_overlap), quantities = TRUE)
  
  # Draw the plot to the graphics device
  grid.draw(euler_plot_grob)
  # --- END OF PLOTTING MODIFICATION ---


  # 6. Find the overlapping proteins (This was step 5)
  common_proteins <- intersect(enriched_rbps_bucket2, protein_list_final)

  # 7. Perform the Hypergeometric Test (This was step 6)
  message("\n--- Calculating Statistical Significance of Overlap ---")

  # q = number of overlapping genes (our successes)
  q <- length(common_proteins)
  
  # m = total number of "successes" in the universe (List A)
  m <- length(enriched_rbps_bucket2)
  
  # k = the size of our "draw" (List B)
  k <- length(protein_list_final) 
  
  # n = total number of "failures" in the universe (all RBPs in universe - List A)
  n <- length(total_rbp_universe) - m

  # Calculate the p-value:
  p_value <- phyper(q - 1, m, n, k, lower.tail = FALSE)

  cat("--- Test Parameters --- \n")
  cat("Overlap size (q):", q, "\n")
  cat("Size of Enriched List 'm' (List A):", m, "\n")
  cat("Size of Your Aligned List 'k' (List B):", k, "\n")
  cat("Size of Universe 'm+n':", m + n, "\n")
  cat("---------------------- \n")
  cat("Hypergeometric test p-value:", format.pval(p_value, digits = 3), "\n")
  
  if (p_value < 0.05) {
    cat("Result: The overlap is STATISTICALLY SIGNIFICANT.\n")
  } else {
    cat("Result: The overlap is NOT statistically significant.\n")
  }

} else {
  message("One or more required objects are missing ('summary_table_bucket2', 'rbp_only_motifs', 'protein_list', 'rbp_data') or no significant motifs were found.")
}
```




Overlap of bucket 1 and bucket 2
```{r}
# --- Step 11: Overlap Bucket 1 and Bucket 2 Results ---

# --- MODIFIED LIBRARIES ---
# Make sure the eulerr package is installed: install.packages("eulerr")
library(eulerr)     # For plotting area-proportional Euler diagrams
library(grid)       # For drawing the plot grob
library(dplyr)

message("--- Comparing Enriched RBPs from Bucket 1 and Bucket 2 ---")

# First, ensure the necessary summary tables exist and are not empty
if (exists("summary_table_bucket1") && exists("summary_table_bucket2") &&
    exists("rbp_only_motifs") && # Added check for universe
    nrow(summary_table_bucket1) > 0) {

  # 1. Extract the unique gene symbols of significantly enriched RBPs from each bucket
  bucket1_rbps <- summary_table_bucket1 %>%
    dplyr::filter(p_adj < 0.05) %>%
    dplyr::pull(geneSymbol) %>%
    unique()

  bucket2_rbps <- summary_table_bucket2 %>%
    dplyr::filter(p_adj < 0.05) %>%
    dplyr::pull(geneSymbol) %>%
    unique()

  # 2. Prepare the data for the Euler diagram
  venn_list <- list(
    `Bucket 1 (Prot UP)` = bucket1_rbps,
    `Bucket 2 (Prot DOWN)` = bucket2_rbps
  )

  # --- MODIFIED PLOTTING STEP ---
  # 3. Create and draw the Euler diagram
  message("\n--- Generating Euler Diagram ---")
  grid.newpage() # Clear the plot device
  
  # Create the euler plot object
  euler_plot_grob <- plot(euler(venn_list), quantities = TRUE)
  
  # Draw the plot to the graphics device
  grid.draw(euler_plot_grob)
  # --- END OF PLOTTING MODIFICATION ---


  # 4. List the common and unique proteins
  common_rbps <- intersect(bucket1_rbps, bucket2_rbps)
  cat("\n--- Summary of Overlap ---\n")
  cat("\nRBPs common to BOTH buckets:", paste(common_rbps, collapse=", "), "\n")


  # --- Step 5: Perform the Hypergeometric Test for Statistical Significance ---
  message("\n--- Calculating Statistical Significance of Overlap ---")

  # Define the "universe" as all RBPs with motifs that were tested in our analysis
  total_rbp_universe <- unique(values(rbp_only_motifs)$geneSymbol)

  # Define the four numbers for the test:
  q <- length(common_rbps) # Overlap
  m <- length(bucket1_rbps) # Size of List 1
  k <- length(bucket2_rbps) # Size of List 2
  n <- length(total_rbp_universe) - m # All non-List 1 items in universe

  # Calculate the p-value for the overlap
  p_value <- phyper(q - 1, m, n, k, lower.tail = FALSE)

  cat("\nOverlap size (q):", q, "\n")
  cat("Size of List 1 'm' (Bucket 1):", m, "\n")
  cat("Size of List 2 'k' (Bucket 2):", k, "\n")
  cat("Size of Universe 'm+n':", m + n, "\n")
  cat("---------------------- \n")
  cat("Hypergeometric test p-value:", format.pval(p_value, digits = 3), "\n")

  if (p_value < 0.05) {
    cat("Result: The overlap between Bucket 1 and Bucket 2 enriched RBPs is STATISTICALLY SIGNIFICANT.\n")
  } else {
    cat("Result: The overlap is NOT statistically significant.\n")
  }

} else {
  message("One or more required objects are missing ('summary_table_bucket1', 'summary_table_bucket2', 'rbp_only_motifs') or the summary tables are empty. Cannot perform overlap analysis.")
}
```

```{r}
# --- STEP 1: SET YOUR GENES OF INTEREST HERE ---
# This list now contains all the RBPs you provided.
genes_to_plot <- c(
  # Common to both buckets
  "TSN", "PTCD1", "FHL2", "MRPL2", "CNOT6", "RBM42", "ZNF71", "RBM22", "ZSCAN31", "RAN", "ATF2", "WDR83", "ESRP1", "PURG", "HBP1", "DHX36", "DTL", "SNRNP70", "CHD2", "CDX1", "EWSR1", "GATA1", "SRF", "HES7", "SOX4", "WT1", "CEBPB", "ATF1", "MAX", "SMAD2", "TFE3", "JUN", "CREB1", "XBP1", "BPTF", "FOXK1", "FUBP1", "HLTF", "KLF12", "SMAD1", "FOXA2", "PAX6", "SMAD3", "STAT3", "ZNF41", "POU2F1", "FOXP1", "PBX2", "ZNF317", "ZNF257", "ZNF148", "CTCFL", "ZNF263", "BNC2", "PATZ1", "ZNF281", "NRF1", "CDC5L", "GZF1", "PRDM14", "STAT6", "WRNIP1", "ZBTB33", "GABPA", "SOX2", "FOXA3", "PSMC2", "LRRFIP1", "RELA", "YY1", "TFAP4", "RAB2A", "TIMM8A", "CEBPZ", "DLX4", "RARA", "NME1", "SPI1", "FOXC1", "KLF3",
  # Unique to Bucket 1
  "LARP1", "MDM2", "SRP9", "ADARB1", "TRIP10", "SUCLG1", "XRCC1", "E4F1", "AGGF1", "CYB5R1", "GCM2", "MRPL1", "NKX2-1", "MRPS25", "HSPA5", "POLE3", "FOXP4", "EXOSC3", "RFX3", "ZNF143", "ZNF140", "ZNF76", "ZNF324", "ZNF354A", "ZNF582", "ZNF701",
  # Unique to Bucket 2
  "GATA3", "ACO1", "ASPSCR1", "ESRRA", "HOXB6", "SND1", "R3HDM2", "TSNAX", "RFC3", "NELFB", "PTPMT1", "SOX9", "RPL35", "DLX2", "NR0B1", "RBBP5", "ESRP2", "MEX3C", "ZNF124", "SF1", "ZNF207", "SRRM3", "NMI", "CSNK2B", "GAR1", "DDX43", "LARP4", "RPS4X", "TRMT1", "TFDP1", "MYEF2", "NFIC", "MSI2", "ZEB1", "LAS1L", "TFAM", "TCEAL2", "NKX3-1", "PSMA6", "SNAPC4", "RBM3", "NUP107", "RBFOX2", "UBXN1", "PDCD11", "KLF13", "SRBD1", "HDAC8", "RNASEH2C", "BOLL", "SMAD4", "HNRNPH3", "PICK1", "UTP18", "RAB18", "MAPK1", "SPATS2", "GIT2", "HIRIP3", "USP39", "BRCA1", "PPARG", "ZNF830", "DDX4", "TOB2", "PITX3", "KLF7", "SMARCC1", "RNF138", "NFKB1", "EGR1", "NFKB2", "MAZ", "RFX5", "C19orf25", "CCDC25", "CSTF2", "DIS3", "ENO1", "GTPBP6", "RPL6", "STAT1", "FOXK2", "NR1H4", "POU5F1", "NKX2-5", "CBFB", "YBX1", "PAXIP1", "PKM", "UGP2", "CELF4", "TFEC", "BAX", "ZNF706", "ABCF2", "TPI1", "GTPBP1", "NOC2L", "OTUD4", "ZRSR2", "PPP1R10", "MGA", "UBE2V1", "MSX2", "HOXD4", "CPEB1", "GRHPR", "KDM5A", "A1CF", "MSI1", "MBD2", "GTF2B", "STUB1", "DAZAP1", "MECP2", "RIOK2", "PCK2", "RFC2", "LSM6", "CLK1", "CELF6", "HNRNPC"
)



# --- STEP 2: LOAD LIBRARIES AND YOUR DATA ---
library(ggplot2)
library(ggpubr)
library(dplyr)
library(SummarizedExperiment)
library(DESeq2)
library(patchwork) # Library for combining plots

message("Loading data...")
tryCatch({
    # Proteome
    new_data_imp_8330 <- readRDS("new_data_imp_8330_BH_new.rds"); dep_8330 <- readRDS("dep_8330_BH_new.rds")
    new_data_imp_MGH <- readRDS("new_data_imp_MGH_BH_new.rds"); dep_MGH <- readRDS("dep_MGH_BH_new.rds")
    # Ubiquitome
    Ubi_data_imp_8330 <- readRDS("Ubi_Ubi_data_imp_8330_BH_new.rds"); Ubi_dep_8330 <- readRDS("Ubi_Ubi_dep_8330_BH_new.rds")
    Ubi_data_imp_MGH <- readRDS("Ubi_Ubi_data_imp_MGH_BH_new.rds"); Ubi_dep_MGH <- readRDS("Ubi_Ubi_dep_MGH_BH_new.rds")
    # RNA-Seq
    dds_RNA <- readRDS(file = "dds_RNA_30102024.rds"); DEG_list_MGH <- readRDS(file = "DEG_list_MGH_30102024.rds"); DEG_list_8330 <- readRDS(file = "DEG_list_8330_30102024.rds")
    
    # Prepare RNA Data
    dds_RNA_8330 <- dds_RNA[, dds_RNA$background == "8330"]; dds_RNA_MGH <- dds_RNA[, dds_RNA$background == "MGH"]
    colData(dds_RNA_8330)$condition <- colData(dds_RNA_8330)$genotype; colData(dds_RNA_MGH)$condition <- colData(dds_RNA_MGH)$genotype
    
    message("All data loaded and prepared successfully.")
}, error = function(e) {
    stop("Data loading failed: ", e$message)
})

# --- STEP 3: DEFINE MASTER ORDER AND PLOTTING FUNCTION ---

master_condition_order <- levels(colData(dds_RNA_8330)$condition)
message("Master condition order set to: ", paste(master_condition_order, collapse=", "))

gg_color_hue <- function(n) { hues = seq(15, 375, length = n + 1); hcl(h = hues, l = 65, c = 100)[1:n] }
get_significance_label <- function(p_adj) {
  if (is.na(p_adj)) { return("") } else if (p_adj <= 0.0001) { return("****") }
  else if (p_adj <= 0.001) { return("***") } else if (p_adj <= 0.01) { return("**") }
  else if (p_adj <= 0.05) { return("*") } else { return("ns") }
}

### Final plotAbundance function with updated theme ###
plotAbundance <- function(data, GOI, dep_results, ref_group = "WT",
                          plot_color = "grey80", data_source_label = "",
                          master_order,
                          star_vjust = -0.5, star_size = 6) {

  if (!GOI %in% rownames(data)) {
     return(ggplot() + theme_void() + annotate("text", x=0.5, y=0.5, label="Gene not found\nin this dataset") + ggtitle(GOI) + labs(subtitle=data_source_label))
  }

  if (inherits(data, "DESeqDataSet")) {
      if (is.null(sizeFactors(data))) data <- estimateSizeFactors(data); expr_values <- counts(data, normalized = TRUE)[GOI, ]; y_axis_label <- "Normalized Counts"
  } else if (is(data, "SummarizedExperiment")) {
      expr_values <- assay(data)[GOI, ]; y_axis_label <- "Normalized log2 Intensity"
  }

  plot_df <- data.frame(expression = expr_values, condition = colData(data)$condition)
  plot_df$condition <- factor(plot_df$condition, levels = master_order)
  other_conditions <- setdiff(master_order, ref_group)

  annotation_info <- list(); y_offset <- diff(range(plot_df$expression, na.rm = TRUE)) * 0.05
  if (!is.finite(y_offset) || y_offset == 0) y_offset <- max(abs(plot_df$expression), na.rm=TRUE) * 0.05
  
  if (is(dep_results, "SummarizedExperiment")) { results_metadata <- as.data.frame(elementMetadata(dep_results)); goi_idx <- which(results_metadata$name == GOI)[1]
    if (length(goi_idx) > 0 && !is.na(goi_idx)) { for (cond in other_conditions) { p_col_name <- paste0(cond, "_vs_", ref_group, "_p.adj")
        if (p_col_name %in% names(results_metadata)) { p_adj <- results_metadata[goi_idx, p_col_name]; signif_label <- get_significance_label(p_adj)
          if (signif_label != "ns" && signif_label != "") { max_y <- max(plot_df$expression[plot_df$condition == cond], na.rm = TRUE); annotation_info[[cond]] <- data.frame(condition = cond, y_pos = max_y + y_offset, label = signif_label) }
        }
    }}
  } else if (is(dep_results, "list")) {
    for (cond in other_conditions) { res_name <- paste0("genotype_", cond, "_vs_", ref_group)
      if (res_name %in% names(dep_results)) { res_obj <- dep_results[[res_name]]
        if (is(res_obj, "DESeqResults") && GOI %in% rownames(res_obj)) { p_adj <- res_obj[GOI, "padj"]; signif_label <- get_significance_label(p_adj)
          if (signif_label != "ns" && signif_label != "") { max_y <- max(plot_df$expression[plot_df$condition == cond], na.rm = TRUE); annotation_info[[cond]] <- data.frame(condition = cond, y_pos = max_y + y_offset, label = signif_label) }
        }
      }
    }
  }
  annotation_df <- bind_rows(annotation_info)

  p <- ggplot(plot_df, aes(x = condition, y = expression)) +
    geom_boxplot(fill = plot_color, alpha = 0.6, color = NA, outlier.shape = NA) +
    geom_boxplot(fill = NA, color = plot_color, outlier.shape = NA) +
    geom_jitter(color = plot_color, size = 3.5, width = 0.1, alpha = 0.7) +
    scale_x_discrete(drop = FALSE) +
    labs(title = GOI, subtitle = data_source_label, y = y_axis_label, x = NULL) +
    theme_bw(base_size = 16) +
    theme(
      text = element_text(color = "black"), axis.text = element_text(color = "black"),
      axis.ticks = element_line(color = "black"), axis.title.y = element_text(size = 12),
      panel.border = element_blank(), # <<< CHANGE 1: Remove the full border
      axis.line.x = element_line(color = "black"), # <<< CHANGE 2: Add back only the x-axis line
      axis.line.y = element_line(color = "black"), # <<< CHANGE 3: Add back only the y-axis line
      legend.position = "none",
      plot.title = element_text(hjust = 0.5, face = "bold"),
      plot.subtitle = element_text(hjust = 0.5), axis.text.x = element_text(angle = 45, hjust = 1),
      panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
      plot.margin = margin(t = 20, r = 5, b = 5, l = 5, "pt")
    )

  if (nrow(annotation_df) > 0) {
    p <- p + geom_text(data = annotation_df, aes(x = condition, y = y_pos, label = label), inherit.aes = FALSE, size = star_size, vjust = star_vjust)
  }
  
  min_y <- min(plot_df$expression, na.rm = TRUE)
  max_y_data <- max(plot_df$expression, na.rm = TRUE)
  max_y_annot <- if (nrow(annotation_df) > 0) max(annotation_df$y_pos, na.rm = TRUE) else -Inf
  highest_point <- max(max_y_data, max_y_annot, na.rm = TRUE)
  final_upper_limit <- highest_point + (diff(range(plot_df$expression, na.rm=TRUE)) * 0.15)
  p <- p + coord_cartesian(ylim = c(min_y, final_upper_limit), clip = "off")

  return(p)
}

# --- STEP 4: GENERATE AND PRINT PLOTS FOR EACH GENE ---
plot_colors <- gg_color_hue(2)
my_color_8330 <- plot_colors[1]
my_color_MGH <- plot_colors[2]

for (current_goi in genes_to_plot) {
  message(paste("\n\n<<<<<<<<<< GENERATING PLOTS FOR GENE:", current_goi, ">>>>>>>>>>\n"))

  tryCatch({
  
    p_prot_8330 <- plotAbundance(new_data_imp_8330, current_goi, dep_8330, master_order = master_condition_order, plot_color = my_color_8330, data_source_label = "Proteome - 8330")
    p_prot_MGH <- plotAbundance(new_data_imp_MGH, current_goi, dep_MGH, master_order = master_condition_order, plot_color = my_color_MGH, data_source_label = "Proteome - MGH")
  
    message(paste("\n--- Individual Plots for", current_goi, "---"))
   print(p_prot_8330); print(p_prot_MGH)
    
    message(paste("\n--- Combined RNA-Seq and Proteome Grid for", current_goi, "---"))
    

  }, error = function(e) {
    message(paste("!!! An error occurred while plotting", current_goi, ":", e$message, "!!!"))
  })
  
  message(paste("\n<<<<<<<<<< Finished plotting for", current_goi, ">>>>>>>>>>"))
}
```

```{r}
# --- STEP 1: SET YOUR GENES OF INTEREST HERE ---
# This list now contains all the RBPs you provided.
genes_to_plot <- c("RBM3", "HNRNPH3", "ENO1", "YBX1", "HNRNPC"
  
)



# --- STEP 2: LOAD LIBRARIES AND YOUR DATA ---
library(ggplot2)
library(ggpubr)
library(dplyr)
library(SummarizedExperiment)
library(DESeq2)
library(patchwork) # Library for combining plots

message("Loading data...")
tryCatch({
    # Proteome
    new_data_imp_8330 <- readRDS("new_data_imp_8330_BH_new.rds"); dep_8330 <- readRDS("dep_8330_BH_new.rds")
    new_data_imp_MGH <- readRDS("new_data_imp_MGH_BH_new.rds"); dep_MGH <- readRDS("dep_MGH_BH_new.rds")
    # Ubiquitome
    Ubi_data_imp_8330 <- readRDS("Ubi_Ubi_data_imp_8330_BH_new.rds"); Ubi_dep_8330 <- readRDS("Ubi_Ubi_dep_8330_BH_new.rds")
    Ubi_data_imp_MGH <- readRDS("Ubi_Ubi_data_imp_MGH_BH_new.rds"); Ubi_dep_MGH <- readRDS("Ubi_Ubi_dep_MGH_BH_new.rds")
    # RNA-Seq
    dds_RNA <- readRDS(file = "dds_RNA_30102024.rds"); DEG_list_MGH <- readRDS(file = "DEG_list_MGH_30102024.rds"); DEG_list_8330 <- readRDS(file = "DEG_list_8330_30102024.rds")
    
    # Prepare RNA Data
    dds_RNA_8330 <- dds_RNA[, dds_RNA$background == "8330"]; dds_RNA_MGH <- dds_RNA[, dds_RNA$background == "MGH"]
    colData(dds_RNA_8330)$condition <- colData(dds_RNA_8330)$genotype; colData(dds_RNA_MGH)$condition <- colData(dds_RNA_MGH)$genotype
    
    message("All data loaded and prepared successfully.")
}, error = function(e) {
    stop("Data loading failed: ", e$message)
})

# --- STEP 3: DEFINE MASTER ORDER AND PLOTTING FUNCTION ---

master_condition_order <- levels(colData(dds_RNA_8330)$condition)
message("Master condition order set to: ", paste(master_condition_order, collapse=", "))

gg_color_hue <- function(n) { hues = seq(15, 375, length = n + 1); hcl(h = hues, l = 65, c = 100)[1:n] }
get_significance_label <- function(p_adj) {
  if (is.na(p_adj)) { return("") } else if (p_adj <= 0.0001) { return("****") }
  else if (p_adj <= 0.001) { return("***") } else if (p_adj <= 0.01) { return("**") }
  else if (p_adj <= 0.05) { return("*") } else { return("ns") }
}

### Final plotAbundance function with updated theme ###
plotAbundance <- function(data, GOI, dep_results, ref_group = "WT",
                          plot_color = "grey80", data_source_label = "",
                          master_order,
                          star_vjust = -0.5, star_size = 6) {

  if (!GOI %in% rownames(data)) {
     return(ggplot() + theme_void() + annotate("text", x=0.5, y=0.5, label="Gene not found\nin this dataset") + ggtitle(GOI) + labs(subtitle=data_source_label))
  }

  if (inherits(data, "DESeqDataSet")) {
      if (is.null(sizeFactors(data))) data <- estimateSizeFactors(data); expr_values <- counts(data, normalized = TRUE)[GOI, ]; y_axis_label <- "Normalized Counts"
  } else if (is(data, "SummarizedExperiment")) {
      expr_values <- assay(data)[GOI, ]; y_axis_label <- "Normalized log2 Intensity"
  }

  plot_df <- data.frame(expression = expr_values, condition = colData(data)$condition)
  plot_df$condition <- factor(plot_df$condition, levels = master_order)
  other_conditions <- setdiff(master_order, ref_group)

  annotation_info <- list(); y_offset <- diff(range(plot_df$expression, na.rm = TRUE)) * 0.05
  if (!is.finite(y_offset) || y_offset == 0) y_offset <- max(abs(plot_df$expression), na.rm=TRUE) * 0.05
  
  if (is(dep_results, "SummarizedExperiment")) { results_metadata <- as.data.frame(elementMetadata(dep_results)); goi_idx <- which(results_metadata$name == GOI)[1]
    if (length(goi_idx) > 0 && !is.na(goi_idx)) { for (cond in other_conditions) { p_col_name <- paste0(cond, "_vs_", ref_group, "_p.adj")
        if (p_col_name %in% names(results_metadata)) { p_adj <- results_metadata[goi_idx, p_col_name]; signif_label <- get_significance_label(p_adj)
          if (signif_label != "ns" && signif_label != "") { max_y <- max(plot_df$expression[plot_df$condition == cond], na.rm = TRUE); annotation_info[[cond]] <- data.frame(condition = cond, y_pos = max_y + y_offset, label = signif_label) }
        }
    }}
  } else if (is(dep_results, "list")) {
    for (cond in other_conditions) { res_name <- paste0("genotype_", cond, "_vs_", ref_group)
      if (res_name %in% names(dep_results)) { res_obj <- dep_results[[res_name]]
        if (is(res_obj, "DESeqResults") && GOI %in% rownames(res_obj)) { p_adj <- res_obj[GOI, "padj"]; signif_label <- get_significance_label(p_adj)
          if (signif_label != "ns" && signif_label != "") { max_y <- max(plot_df$expression[plot_df$condition == cond], na.rm = TRUE); annotation_info[[cond]] <- data.frame(condition = cond, y_pos = max_y + y_offset, label = signif_label) }
        }
      }
    }
  }
  annotation_df <- bind_rows(annotation_info)

  p <- ggplot(plot_df, aes(x = condition, y = expression)) +
    geom_boxplot(fill = plot_color, alpha = 0.6, color = NA, outlier.shape = NA) +
    geom_boxplot(fill = NA, color = plot_color, outlier.shape = NA) +
    geom_jitter(color = plot_color, size = 3.5, width = 0.1, alpha = 0.7) +
    scale_x_discrete(drop = FALSE) +
    labs(title = GOI, subtitle = data_source_label, y = y_axis_label, x = NULL) +
    theme_bw(base_size = 16) +
    theme(
      text = element_text(color = "black"), axis.text = element_text(color = "black"),
      axis.ticks = element_line(color = "black"), axis.title.y = element_text(size = 12),
      panel.border = element_blank(), # <<< CHANGE 1: Remove the full border
      axis.line.x = element_line(color = "black"), # <<< CHANGE 2: Add back only the x-axis line
      axis.line.y = element_line(color = "black"), # <<< CHANGE 3: Add back only the y-axis line
      legend.position = "none",
      plot.title = element_text(hjust = 0.5, face = "bold"),
      plot.subtitle = element_text(hjust = 0.5), axis.text.x = element_text(angle = 45, hjust = 1),
      panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
      plot.margin = margin(t = 20, r = 5, b = 5, l = 5, "pt")
    )

  if (nrow(annotation_df) > 0) {
    p <- p + geom_text(data = annotation_df, aes(x = condition, y = y_pos, label = label), inherit.aes = FALSE, size = star_size, vjust = star_vjust)
  }
  
  min_y <- min(plot_df$expression, na.rm = TRUE)
  max_y_data <- max(plot_df$expression, na.rm = TRUE)
  max_y_annot <- if (nrow(annotation_df) > 0) max(annotation_df$y_pos, na.rm = TRUE) else -Inf
  highest_point <- max(max_y_data, max_y_annot, na.rm = TRUE)
  final_upper_limit <- highest_point + (diff(range(plot_df$expression, na.rm=TRUE)) * 0.15)
  p <- p + coord_cartesian(ylim = c(min_y, final_upper_limit), clip = "off")

  return(p)
}

# --- STEP 4: GENERATE AND PRINT PLOTS FOR EACH GENE ---
plot_colors <- gg_color_hue(2)
my_color_8330 <- plot_colors[1]
my_color_MGH <- plot_colors[2]

for (current_goi in genes_to_plot) {
  message(paste("\n\n<<<<<<<<<< GENERATING PLOTS FOR GENE:", current_goi, ">>>>>>>>>>\n"))

  tryCatch({
  
    p_prot_8330 <- plotAbundance(new_data_imp_8330, current_goi, dep_8330, master_order = master_condition_order, plot_color = my_color_8330, data_source_label = "Proteome - 8330")
    p_prot_MGH <- plotAbundance(new_data_imp_MGH, current_goi, dep_MGH, master_order = master_condition_order, plot_color = my_color_MGH, data_source_label = "Proteome - MGH")
  
    message(paste("\n--- Individual Plots for", current_goi, "---"))
   print(p_prot_8330); print(p_prot_MGH)
    
    message(paste("\n--- Combined RNA-Seq and Proteome Grid for", current_goi, "---"))
    

  }, error = function(e) {
    message(paste("!!! An error occurred while plotting", current_goi, ":", e$message, "!!!"))
  })
  
  message(paste("\n<<<<<<<<<< Finished plotting for", current_goi, ">>>>>>>>>>"))
}
```


MotifMatchR miRNAs
```{r}
# --- Step 1: Load Libraries ---
library(DBI)
library(targetscan.Hs.eg.db)
library(Biostrings)
library(GenomicRanges)
library(BSgenome.Hsapiens.UCSC.hg38)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(org.Hs.eg.db)
library(knitr)
library(tidyverse)


# This script assumes 'foreground_genes_bucket1' and 'background_genes' are pre-defined.


# --- Step 3: Prepare Sequences and Gene Lists ---
message("Fetching 3' UTRs and mapping to gene symbols for Bucket 1...")
utrs <- threeUTRsByTranscript(TxDb.Hsapiens.UCSC.hg38.knownGene, use.names = TRUE)
tx_names <- names(utrs)
tx2geneid <- AnnotationDbi::select(
  TxDb.Hsapiens.UCSC.hg38.knownGene, keys = tx_names,
  keytype = "TXNAME", columns = "GENEID"
)
geneid2symbol <- AnnotationDbi::select(
  org.Hs.eg.db, keys = unique(tx2geneid$GENEID),
  keytype = "ENTREZID", columns = "SYMBOL"
)
tx2symbol <- dplyr::left_join(tx2geneid, geneid2symbol, by = c("GENEID" = "ENTREZID")) %>%
  dplyr::filter(!is.na(SYMBOL) & !is.na(TXNAME))
foreground_tx_names <- tx2symbol %>%
  dplyr::filter(SYMBOL %in% foreground_genes_bucket1) %>%
  dplyr::pull(TXNAME) %>% unique()
background_tx_names <- tx2symbol %>%
  dplyr::filter(SYMBOL %in% background_genes) %>%
  dplyr::pull(TXNAME) %>% unique()
if (length(foreground_tx_names) == 0) {
  stop("Error: None of the foreground gene symbols could be mapped to a transcript.")
}
message("Extracting genomic sequences...")
utrs_foreground <- utrs[names(utrs) %in% foreground_tx_names]
utrs_background <- utrs[names(utrs) %in% background_tx_names]
seqs_foreground_dna <- getSeq(BSgenome.Hsapiens.UCSC.hg38, utrs_foreground)
seqs_background_dna <- getSeq(BSgenome.Hsapiens.UCSC.hg38, utrs_background)


# --- Step 4: Prepare miRNA Motifs from TargetScan ---
message("Preparing miRNA seed region motifs from TargetScan...")
db_path <- system.file("extdata", "targetscan.Hs.eg.sqlite", package = "targetscan.Hs.eg.db")
con <- DBI::dbConnect(RSQLite::SQLite(), db_path)
family_table <- DBI::dbReadTable(con, "mirna_family")
mirbase_table <- DBI::dbReadTable(con, "mirbase")
DBI::dbDisconnect(con)
master_table <- dplyr::inner_join(family_table, mirbase_table, by = c("X_id" = "family"))
mirs_filtered <- master_table %>%
  dplyr::select(miR_Family = name, seed_m8) %>%
  dplyr::distinct()

mir_seeds_dna <- gsub("U", "T", mirs_filtered$seed_m8)
names(mir_seeds_dna) <- mirs_filtered$miR_Family
mir_seeds_dna <- mir_seeds_dna[!is.na(mir_seeds_dna)]
message(paste("Successfully prepared", length(mir_seeds_dna), "unique miRNA motifs for analysis."))


# --- Step 5: Find Motif Matches using Biostrings ---
message("Scanning DNA sequences for miRNA motif matches...")
seqs_foreground_flat <- unlist(seqs_foreground_dna)
seqs_background_flat <- unlist(seqs_background_dna)

foreground_counts_list <- lapply(mir_seeds_dna, function(seed) {
    Biostrings::vcountPattern(pattern = seed, 
                              subject = seqs_foreground_flat, 
                              max.mismatch = 0)
})
motif_hits_foreground <- do.call(cbind, foreground_counts_list)

background_counts_list <- lapply(mir_seeds_dna, function(seed) {
    Biostrings::vcountPattern(pattern = seed, 
                              subject = seqs_background_flat, 
                              max.mismatch = 0)
})
motif_hits_background <- do.call(cbind, background_counts_list)


# --- Step 6: Perform Enrichment Analysis for miRNAs (Bucket 1) ---
message("Performing Fisher's exact test for enrichment...")
final_results_miRNA_bucket1 <- tibble() 
if (ncol(motif_hits_foreground) > 0) {
    foreground_hits <- colSums(motif_hits_foreground > 0)
    background_hits <- colSums(motif_hits_background > 0)
    
    num_foreground_seqs <- nrow(motif_hits_foreground)
    num_background_seqs <- nrow(motif_hits_background)

    fisher_results <- lapply(1:length(mir_seeds_dna), function(i) {
        tbl <- matrix(c(
            foreground_hits[i], background_hits[i],
            num_foreground_seqs - foreground_hits[i],
            num_background_seqs - background_hits[i]
        ), nrow = 2)
        fisher.test(tbl, alternative = "greater")
    })

    final_results_miRNA_bucket1 <- tibble(
        miR_Family = names(mir_seeds_dna),
        p_value = sapply(fisher_results, "[[", "p.value"),
        odds_ratio = sapply(fisher_results, function(x) x$estimate)
    ) %>%
    dplyr::mutate(p_adj = p.adjust(p_value, method = "BH")) %>%
    dplyr::arrange(p_adj)
} else {
    warning("Motif matching returned an empty result.")
}
print("Top 20 enriched miRNA Families for Bucket 1:")
print(head(final_results_miRNA_bucket1, 20))


# --- Step 7: Visualize the Results for miRNAs (Bucket 1) ---


if (nrow(final_results_miRNA_bucket1) > 0) {
  message("Generating final results dot plot for miRNAs (Bucket 1)...")
  
  top_motifs_plot_miRNA_bucket1 <- final_results_miRNA_bucket1 %>%
    # Data preparation steps
    dplyr::filter(is.finite(odds_ratio) & p_adj < 0.05) %>%
    dplyr::top_n(15, -p_adj) %>%
    dplyr::mutate(
      log10_padj = -log10(p_adj),
      miR_Family = fct_reorder(miR_Family, odds_ratio) 
    ) %>%
    
    # ggplot call with aesthetics
    ggplot(aes(x = odds_ratio, y = miR_Family, size = odds_ratio, color = log10_padj)) +
    geom_point(alpha = 0.8) +
    
    # Define the color scale
    scale_color_gradient(low = "lightblue", high = "firebrick") +
    
    # --- CHANGE HERE ---
    # Define the size scale with your four specified breaks
    scale_size_continuous(range = c(4, 10), breaks = c(1.4, 1.6, 1.8, 2.0)) +
    
    # Use labs() to provide distinct titles for each legend
    labs(
      title = "Top 15 Enriched miRNA Motifs in Foreground Genes (Bucket 1)",
      x = "Odds Ratio",
      y = "miRNA Family",
      size = "Odds Ratio",
      color = bquote(-Log[10]~"(p-adj)")
    ) +
    
    # Apply a minimal theme and remove gridlines
    theme_minimal() +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
    )

  print(top_motifs_plot_miRNA_bucket1)
  
} else {
  message("No significant miRNA motifs found to plot for Bucket 1.")
}


# --- Step 8: Print Final Summary Table for miRNAs (Bucket 1) ---
message("--- Final Summary of Enriched miRNA Motifs (p_adj < 0.05) for Bucket 1 ---")
if (nrow(final_results_miRNA_bucket1) > 0) {
  
  summary_table_miRNA_bucket1 <- final_results_miRNA_bucket1 %>%
    dplyr::filter(p_adj < 0.05)
  
  if(nrow(summary_table_miRNA_bucket1) > 0){
    print(knitr::kable(summary_table_miRNA_bucket1, digits = 3, caption = "Enriched miRNA Motifs (Bucket 1)"))
  } else {
    message("No motifs passed the significance threshold of p_adj < 0.05.")
  }
} else {
  message("No enrichment results were generated to create a summary table.")
}
```

Enriched miRNA motifs in bucket 2
```{r}
# --- Step 1: Load Libraries ---
library(DBI)
library(targetscan.Hs.eg.db)
library(Biostrings)
library(GenomicRanges)
library(BSgenome.Hsapiens.UCSC.hg38)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(org.Hs.eg.db)
library(knitr)
library(tidyverse)


# --- Step 3: Prepare Sequences and Gene Lists ---
message("Fetching 3' UTRs and mapping to gene symbols for Bucket 2...")
utrs <- threeUTRsByTranscript(TxDb.Hsapiens.UCSC.hg38.knownGene, use.names = TRUE)
tx_names <- names(utrs)
tx2geneid <- AnnotationDbi::select(
  TxDb.Hsapiens.UCSC.hg38.knownGene, keys = tx_names,
  keytype = "TXNAME", columns = "GENEID"
)
geneid2symbol <- AnnotationDbi::select(
  org.Hs.eg.db, keys = unique(tx2geneid$GENEID),
  keytype = "ENTREZID", columns = "SYMBOL"
)
tx2symbol <- dplyr::left_join(tx2geneid, geneid2symbol, by = c("GENEID" = "ENTREZID")) %>%
  dplyr::filter(!is.na(SYMBOL) & !is.na(TXNAME))
foreground_tx_names <- tx2symbol %>%
  dplyr::filter(SYMBOL %in% foreground_genes_bucket2) %>%
  dplyr::pull(TXNAME) %>% unique()
background_tx_names <- tx2symbol %>%
  dplyr::filter(SYMBOL %in% background_genes) %>%
  dplyr::pull(TXNAME) %>% unique()
if (length(foreground_tx_names) == 0) {
  stop("Error: None of the foreground gene symbols could be mapped to a transcript.")
}
message("Extracting genomic sequences...")
utrs_foreground <- utrs[names(utrs) %in% foreground_tx_names]
utrs_background <- utrs[names(utrs) %in% background_tx_names]
seqs_foreground_dna <- getSeq(BSgenome.Hsapiens.UCSC.hg38, utrs_foreground)
seqs_background_dna <- getSeq(BSgenome.Hsapiens.UCSC.hg38, utrs_background)


# --- Step 4: Prepare miRNA Motifs from TargetScan ---
message("Preparing miRNA seed region motifs from TargetScan...")
db_path <- system.file("extdata", "targetscan.Hs.eg.sqlite", package = "targetscan.Hs.eg.db")
con <- DBI::dbConnect(RSQLite::SQLite(), db_path)
family_table <- DBI::dbReadTable(con, "mirna_family")
mirbase_table <- DBI::dbReadTable(con, "mirbase")
DBI::dbDisconnect(con)
master_table <- dplyr::inner_join(family_table, mirbase_table, by = c("X_id" = "family"))
mirs_filtered <- master_table %>%
  dplyr::select(miR_Family = name, seed_m8) %>%
  dplyr::distinct()
mir_seeds_dna <- gsub("U", "T", mirs_filtered$seed_m8)
names(mir_seeds_dna) <- mirs_filtered$miR_Family
mir_seeds_dna <- mir_seeds_dna[!is.na(mir_seeds_dna)]
message(paste("Successfully prepared", length(mir_seeds_dna), "unique miRNA motifs for analysis."))


# --- Step 5: Find Motif Matches using Biostrings ---
message("Scanning DNA sequences for miRNA motif matches...")
seqs_foreground_flat <- unlist(seqs_foreground_dna)
seqs_background_flat <- unlist(seqs_background_dna)

foreground_counts_list <- lapply(mir_seeds_dna, function(seed) {
    Biostrings::vcountPattern(pattern = seed, 
                              subject = seqs_foreground_flat, 
                              max.mismatch = 0)
})
motif_hits_foreground <- do.call(cbind, foreground_counts_list)

background_counts_list <- lapply(mir_seeds_dna, function(seed) {
    Biostrings::vcountPattern(pattern = seed, 
                              subject = seqs_background_flat, 
                              max.mismatch = 0)
})
motif_hits_background <- do.call(cbind, background_counts_list)


# --- Step 6: Perform Enrichment Analysis for miRNAs (Bucket 2) ---
message("Performing Fisher's exact test for enrichment...")
final_results_miRNA_bucket2 <- tibble() 
if (ncol(motif_hits_foreground) > 0) {
    foreground_hits <- colSums(motif_hits_foreground > 0)
    background_hits <- colSums(motif_hits_background > 0)
    
    num_foreground_seqs <- nrow(motif_hits_foreground)
    num_background_seqs <- nrow(motif_hits_background)

    fisher_results <- lapply(1:length(mir_seeds_dna), function(i) {
        tbl <- matrix(c(
            foreground_hits[i], background_hits[i],
            num_foreground_seqs - foreground_hits[i],
            num_background_seqs - background_hits[i]
        ), nrow = 2)
        fisher.test(tbl, alternative = "greater")
    })

    final_results_miRNA_bucket2 <- tibble(
        miR_Family = names(mir_seeds_dna),
        p_value = sapply(fisher_results, "[[", "p.value"),
        odds_ratio = sapply(fisher_results, function(x) x$estimate)
    ) %>%
    dplyr::mutate(p_adj = p.adjust(p_value, method = "BH")) %>%
    dplyr::arrange(p_adj)
} else {
    warning("Motif matching returned an empty result.")
}
print("Top 20 enriched miRNA Families for Bucket 2:")
print(head(final_results_miRNA_bucket2, 20))


# --- Step 7: Visualize the Results for miRNAs (Bucket 2) ---

if (nrow(final_results_miRNA_bucket2) > 0) {
  message("Generating results dot plot for miRNAs (Bucket 2)...")
  
  top_motifs_plot_miRNA_bucket2 <- final_results_miRNA_bucket2 %>%
    # Data preparation steps
    dplyr::filter(is.finite(odds_ratio) & p_adj < 0.05) %>%
    
    # --- TYPO FIX HERE ---
    dplyr::top_n(15, -p_adj) %>% # Corrected from -p-adj
    
    dplyr::mutate(
      log10_padj = -log10(p_adj),
      miR_Family = fct_reorder(miR_Family, odds_ratio) 
    ) %>%
    
    # ggplot call with updated aesthetics
    ggplot(aes(x = odds_ratio, y = miR_Family, size = odds_ratio, color = log10_padj)) +
    geom_point(alpha = 0.8) +
    
    # Define the color scale
    scale_color_gradient(low = "lightblue", high = "firebrick") +
    
    # Define the size scale with your specified breaks
    # Note: You may need to adjust these breaks for the Bucket 2 data range
    scale_size_continuous(range = c(4, 10), breaks = c(1.4, 1.6, 1.8, 2.0)) +
    
    # Use labs() to provide distinct titles for each legend
    labs(
      title = "Top 15 Enriched miRNA Motifs in Foreground Genes (Bucket 2)",
      x = "Odds Ratio",
      y = "miRNA Family",
      size = "Odds Ratio",
      color = bquote(-Log[10]~"(p-adj)")
    ) +
    
    # Apply a minimal theme and remove gridlines
    theme_minimal() +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
    )

  print(top_motifs_plot_miRNA_bucket2)
  
} else {
  message("No significant miRNA motifs found to plot for Bucket 2.")
}


# --- Step 8: Print Final Summary Table for miRNAs (Bucket 2) ---
message("--- Final Summary of Enriched miRNA Motifs (p_adj < 0.05) for Bucket 2 ---")
if (nrow(final_results_miRNA_bucket2) > 0) {
  
  summary_table_miRNA_bucket2 <- final_results_miRNA_bucket2 %>%
    dplyr::filter(p_adj < 0.05)
  
  if(nrow(summary_table_miRNA_bucket2) > 0){
    print(knitr::kable(summary_table_miRNA_bucket2, digits = 3, caption = "Enriched miRNA Motifs (Bucket 2)"))
  } else {
    message("No motifs passed the significance threshold of p_adj < 0.05.")
  }
} else {
  message("No enrichment results were generated to create a summary table.")
}
```


```{r}
# --- Step 11: Overlap Bucket 1 and Bucket 2 miRNA Results ---
# Make sure the VennDiagram package is installed: install.packages("VennDiagram")
library(VennDiagram)
library(dplyr)
library(grid) # Required for grid.newpage() and grid.draw()

message("--- Comparing Enriched miRNA Families from Bucket 1 and Bucket 2 ---")

# 1. Extract the unique names of significantly enriched miRNA families from each bucket
bucket1_miRNAs <- final_results_miRNA_bucket1 %>%
  dplyr::filter(p_adj < 0.05) %>%
  dplyr::pull(miR_Family) %>%
  unique()

bucket2_miRNAs <- final_results_miRNA_bucket2 %>%
  dplyr::filter(p_adj < 0.05) %>%
  dplyr::pull(miR_Family) %>%
  unique()

# 2. Prepare the data for the Venn diagram
venn_list_miRNA <- list(
  "Bucket 1" = bucket1_miRNAs,
  "Bucket 2" = bucket2_miRNAs
)

# 3. Create and draw the Venn diagram
# This clears any previous plots
grid.newpage() 

# Generate the plot object
# Load the necessary library
library(VennDiagram)

# Your list of items for the Venn diagram
# venn_list_miRNA <- list(Bucket1 = ..., Bucket2 = ...)

# --- Corrected Plotting Code ---
venn.diagram(
  x = venn_list_miRNA,
  
  # 1. Set a filename
  filename = 'venn_plot.png',
  output = TRUE,
  
  # 2. Set equal width and height for a square canvas
  imagetype="png" ,
  height = 2000, 
  width = 2000, 
  resolution = 300,
  compression = "lzw",
  
  # --- Style Options ---
  col = c("seagreen", "darkorange"),
  fill = c(alpha("seagreen", 0.3), alpha("darkorange", 0.3)),
  cex = 1.5,
  fontfamily = "sans",
  cat.cex = 1.2,
  cat.fontfamily = "sans",
  cat.col = c("seagreen", "darkorange")
)

# Draw the plot to the active graphics device
grid.draw(venn_plot_miRNA)


# 4. List the common and unique miRNA families
common_miRNAs <- intersect(bucket1_miRNAs, bucket2_miRNAs)
unique_to_bucket1_miRNAs <- setdiff(bucket1_miRNAs, bucket2_miRNAs)
unique_to_bucket2_miRNAs <- setdiff(bucket2_miRNAs, bucket1_miRNAs)

cat("\n--- Summary of miRNA Overlap ---\n")
cat("\nmiRNA Families common to BOTH buckets:", paste(common_miRNAs, collapse=", "), "\n")
cat("\nmiRNA Families unique to Bucket 1:", paste(unique_to_bucket1_miRNAs, collapse=", "), "\n")
cat("\nmiRNA Families unique to Bucket 2:", paste(unique_to_bucket2_miRNAs, collapse=", "), "\n")
```

```{r}
# --- Step 11: Overlap Bucket 1 and Bucket 2 miRNA Results ---

# --- MODIFIED LIBRARIES ---
# Make sure the eulerr package is installed: install.packages("eulerr")
library(eulerr)     # For plotting area-proportional Euler diagrams
library(grid)       # For drawing the plot grob
library(dplyr)

message("--- Comparing Enriched miRNA Families from Bucket 1 and Bucket 2 ---")

# --- Pre-flight Check ---
# Ensure the necessary result tables exist
if (exists("final_results_miRNA_bucket1") && exists("final_results_miRNA_bucket2")) {

  # 1. Extract the unique names of significantly enriched miRNA families from each bucket
  bucket1_miRNAs <- final_results_miRNA_bucket1 %>%
    dplyr::filter(p_adj < 0.05) %>%
    dplyr::pull(miR_Family) %>%
    unique()

  bucket2_miRNAs <- final_results_miRNA_bucket2 %>%
    dplyr::filter(p_adj < 0.05) %>%
    dplyr::pull(miR_Family) %>%
    unique()
    
  message(paste("Found", length(bucket1_miRNAs), "significant families in Bucket 1."))
  message(paste("Found", length(bucket2_miRNAs), "significant families in Bucket 2."))

  # 2. Prepare the data for the Euler diagram
  venn_list_miRNA <- list(
    "Bucket 1" = bucket1_miRNAs,
    "Bucket 2" = bucket2_miRNAs
  )

  # --- MODIFIED PLOTTING STEP ---
  # 3. Create and draw the Euler diagram
  message("\n--- Generating Euler Diagram ---")
  grid.newpage() # Clear the plot device
  
  # Create the euler plot object
  # quantities = TRUE adds the numbers to the segments
  euler_plot_grob <- plot(euler(venn_list_miRNA), quantities = TRUE)
  
  # Draw the plot to the graphics device
  grid.draw(euler_plot_grob)
  # --- END OF PLOTTING MODIFICATION ---


  # 4. List the common and unique miRNA families
  common_miRNAs <- intersect(bucket1_miRNAs, bucket2_miRNAs)
  unique_to_bucket1_miRNAs <- setdiff(bucket1_miRNAs, bucket2_miRNAs)
  unique_to_bucket2_miRNAs <- setdiff(bucket2_miRNAs, bucket1_miRNAs)

  cat("\n--- Summary of miRNA Overlap ---\n")
  cat("Total overlapping families:", length(common_miRNAs), "\n")
  cat("\nFamilies common to BOTH buckets:", paste(common_miRNAs, collapse=", "), "\n")
  cat("\nFamilies unique to Bucket 1:", paste(unique_to_bucket1_miRNAs, collapse=", "), "\n")
  cat("\nFamilies unique to Bucket 2:", paste(unique_to_bucket2_miRNAs, collapse=", "), "\n")
  
  
  # --- NEW STATISTICAL ANALYSIS STEP ---
  # 5. Perform the Hypergeometric Test
  message("\n--- Calculating Statistical Significance of Overlap ---")
  
  # --- Define the Statistical Universe ---
  # We define the universe as all unique miRNA families present in EITHER 
  # of the full (unfiltered) results tables. This represents all families
  # that were tested and could have been found significant.
  
  all_b1_families <- unique(final_results_miRNA_bucket1$miR_Family)
  all_b2_families <- unique(final_results_miRNA_bucket2$miR_Family)
  total_mirna_universe <- union(all_b1_families, all_b2_families)
  
  if(length(total_mirna_universe) == 0) {
    message("Error: Universe size is zero. Cannot perform hypergeometric test.")
  } else {
    message(paste("Statistical universe defined as all families tested:", length(total_mirna_universe)))

    # Define the four numbers for the test:
    q <- length(common_miRNAs) # Overlap
    m <- length(bucket1_miRNAs) # Size of List 1 (significant in B1)
    k <- length(bucket2_miRNAs) # Size of List 2 (significant in B2)
    n <- length(total_mirna_universe) - m # All non-List 1 items in universe

    # Calculate the p-value for the overlap
    p_value <- phyper(q - 1, m, n, k, lower.tail = FALSE)

    cat("--- Test Parameters --- \n")
    cat("Overlap size (q):", q, "\n")
    cat("Size of List 1 'm' (Bucket 1):", m, "\n")
    cat("Size of List 2 'k' (Bucket 2):", k, "\n")
    cat("Size of Universe 'm+n':", m + n, "\n")
    cat("---------------------- \n")
    cat("Hypergeometric test p-value:", format.pval(p_value, digits = 3), "\n")

    if (p_value < 0.05) {
      cat("Result: The overlap between Bucket 1 and Bucket 2 enriched miRNA families is STATISTICALLY SIGNIFICANT.\n")
    } else {
      cat("Result: The overlap is NOT statistically significant.\n")
    }
  }
  
} else {
  message("Error: 'final_results_miRNA_bucket1' or 'final_results_miRNA_bucket2' not found. Cannot perform overlap analysis.")
}
```



Export results from ridgeline plots
```{r}
# Load necessary library
library(openxlsx)
library(dplyr)

# ==============================================================================
# EXPORT: AVERAGED TRIPLICATES 
# ==============================================================================

# 1. Prepare and Combine Raw Data
mgh_ratio_data$Background <- "MGH"
h8330_ratio_data$Background <- "8330"

raw_combined <- rbind(mgh_ratio_data, h8330_ratio_data)

# 2. Average the Triplicates
# We group by Gene, Background, and Genotype to calculate the mean across replicates
averaged_table <- raw_combined %>%
  group_by(Gene, Background, condition) %>%
  summarise(
    Mean_Protein_Abundance_Log2 = mean(Protein_Abundance, na.rm = TRUE),
    Mean_RNA_Abundance_VST = mean(RNA_Abundance, na.rm = TRUE),
    Mean_Protein_RNA_Log2_Ratio = mean(Log2Ratio, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(condition, Background, Gene) # Sort for clarity

# 3. Initialize Workbook
wb <- createWorkbook()
header_style <- createStyle(textDecoration = "bold", halign = "center")

# 4. Loop Logic: Create sheets for Genotype + Background combinations
unique_genotypes <- unique(averaged_table$condition)
unique_backgrounds <- c("MGH", "8330")

for (geno in unique_genotypes) {
  for (bg in unique_backgrounds) {
    
    # A. Filter data for this specific Genotype AND Background
    sheet_data <- averaged_table %>% 
      filter(condition == geno, Background == bg) %>%
      dplyr::select(-condition, -Background) 
    
    if(nrow(sheet_data) > 0) {
      
      sheet_name <- paste0(geno, "_", bg)
      
      sheet_name <- substr(sheet_name, 1, 31)
      
      addWorksheet(wb, sheet_name)
      
      writeData(wb, sheet = sheet_name, x = sheet_data)
      
      addStyle(wb, sheet = sheet_name, style = header_style, rows = 1, cols = 1:ncol(sheet_data))
      setColWidths(wb, sheet = sheet_name, cols = 1:ncol(sheet_data), widths = "auto")
    }
  }
}

# 5. Save
saveWorkbook(wb, "Supplementary_Table_Protein_RNA_Ratios_Averaged.xlsx", overwrite = TRUE)

cat("\nSuccess! 'Supplementary_Table_Protein_RNA_Ratios_Averaged.xlsx' created.")
cat("\nTriplicates have been averaged. Sheets are organized by Genotype_Background.")
```


Export lists of bucket 1 and 2
```{r}
# Load necessary libraries
library(dplyr)
library(openxlsx)

# ==============================================================================
# EXPORT: OVERLAPPING PROTEOME-SPECIFIC GENES (SIMPLE LIST)
# ==============================================================================

# --- 1. Define Bucket 1 (UP) Intersection ---
# Filter all genes fitting the criteria
bucket1_all <- master_integrated_table %>%
  dplyr::filter(
    Protein_padj < 0.05 & Protein_Log2FC > 0,
    !(mRNA_padj < 0.05 & mRNA_Log2FC > 0)
  )

# Split by comparison and find the intersection (genes present in ALL lists)
bucket1_list <- bucket1_all %>% group_split(comparison)
bucket1_vectors <- lapply(bucket1_list, function(df) df$Gene)
bucket1_overlap <- Reduce(intersect, bucket1_vectors)

# --- 2. Define Bucket 2 (DOWN) Intersection ---
# Filter all genes fitting the criteria
bucket2_all <- master_integrated_table %>%
  dplyr::filter(
    Protein_padj < 0.05 & Protein_Log2FC < 0,
    !(mRNA_padj < 0.05 & mRNA_Log2FC < 0)
  )

# Split by comparison and find the intersection
bucket2_list <- bucket2_all %>% group_split(comparison)
bucket2_vectors <- lapply(bucket2_list, function(df) df$Gene)
bucket2_overlap <- Reduce(intersect, bucket2_vectors)

# --- 3. Prepare Data Frames for Export ---
# Convert the vectors to dataframes so they can be written to Excel
df_bucket1 <- data.frame(Gene = bucket1_overlap)
df_bucket2 <- data.frame(Gene = bucket2_overlap)

# --- 4. Create and Save Excel File ---
wb <- createWorkbook()
header_style <- createStyle(textDecoration = "bold", halign = "center")

# Sheet 1: Proteome-Spec Upregulation
sheet1 <- "Proteome-Spec Upregulation"
addWorksheet(wb, sheet1)
writeData(wb, sheet1, df_bucket1)
addStyle(wb, sheet1, header_style, rows = 1, cols = 1)

# Sheet 2: Proteome-Spec Downregulation
sheet2 <- "Proteome-Spec Downregulation"
addWorksheet(wb, sheet2)
writeData(wb, sheet2, df_bucket2)
addStyle(wb, sheet2, header_style, rows = 1, cols = 1)

# Save
saveWorkbook(wb, "Supplementary_Table_Core_Proteome_Discordance.xlsx", overwrite = TRUE)

cat("\nSuccess! 'Supplementary_Table_Core_Proteome_Discordance.xlsx' created.")
cat("\nProteome-Specific UP Genes (Intersection):", length(bucket1_overlap))
cat("\nProteome-Specific DOWN Genes (Intersection):", length(bucket2_overlap))
```


```{r}
# --- Step 2: Convert Gene Symbols to Entrez IDs ---
foreground_genes_bucket1_entrez <- bitr(
  foreground_genes_bucket1, 
  fromType = "SYMBOL", 
  toType = "ENTREZID", 
  OrgDb = "org.Hs.eg.db"
)

# Check the conversion
print(head(foreground_genes_bucket1_entrez))

# --- Step 3: Run GO Enrichment Analysis ---
bucket2_enrich <- enrichGO(
  gene          = foreground_genes_bucket1_entrez$ENTREZID,  # <--- FIX: Select the column
  keyType       = "ENTREZID",
  OrgDb         = org.Hs.eg.db,
  ont           = "ALL",        # Runs BP, CC, and MF. Use "BP" for just Biological Process.
  pAdjustMethod = "BH",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  readable      = TRUE          # Maps IDs back to Symbols in the output
)

# --- Step 4: View Results ---
if (!is.null(bucket1_enrich)) {
  print(head(bucket1_enrich))
} else {
  print("No significant GO terms found. Try increasing pvalueCutoff to 0.1 or 0.2 to check for weak signal.")
}
```



```{r}
# --- Step 2: Convert Gene Symbols to Entrez IDs ---
foreground_genes_bucket2_entrez <- bitr(
  foreground_genes_bucket2, 
  fromType = "SYMBOL", 
  toType = "ENTREZID", 
  OrgDb = "org.Hs.eg.db"
)

# Check the conversion
print(head(foreground_genes_bucket2_entrez))

# --- Step 3: Run GO Enrichment Analysis ---
bucket2_enrich <- enrichGO(
  gene          = foreground_genes_bucket2_entrez$ENTREZID,  # <--- FIX: Select the column
  keyType       = "ENTREZID",
  OrgDb         = org.Hs.eg.db,
  ont           = "ALL",        # Runs BP, CC, and MF. Use "BP" for just Biological Process.
  pAdjustMethod = "BH",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05,
  readable      = TRUE          # Maps IDs back to Symbols in the output
)

# --- Step 4: View Results ---
if (!is.null(bucket2_enrich)) {
  print(head(bucket2_enrich))
} else {
  print("No significant GO terms found. Try increasing pvalueCutoff to 0.1 or 0.2 to check for weak signal.")
}
```



Heatmap of RBP-enriched proteins
```{r}
suppressPackageStartupMessages({
  library(SummarizedExperiment)
  library(pheatmap)
  library(RColorBrewer)
})

# ----------------------------
# 0) Inputs
# ----------------------------
se_mgh  <- data_imp_MGH_new
se_8330 <- data_imp_8330_new

proteins_to_plot <- c(
  "GLYR1","CHD2","ACO1","RPL35","HSPA5","HNRNPAO","AHCTF1","RBBP5",
  "R3HDM2","SND1","TSNAX","RAN","RFC3","NELFB","PSMC2","CEBPZ"
)

# ----------------------------
# Helpers
# ----------------------------
unwrap_se <- function(x) {
  if (inherits(x, "SummarizedExperiment")) return(x)
  if (is.list(x)) {
    idx <- which(vapply(x, inherits, logical(1), what="SummarizedExperiment"))[1]
    if (!is.na(idx)) return(x[[idx]])
  }
  stop("Could not unwrap into SummarizedExperiment")
}

get_assay_matrix <- function(se) {
  se <- unwrap_se(se)

  an <- assayNames(se)
  if (length(an) == 0) {
    mat <- assay(se, 1)
  } else {
    preferred <- c("imputed","Imputed","log2","intensity","abundance","exprs","counts")
    use <- preferred[preferred %in% an]
    if (length(use) == 0) use <- an[1] else use <- use[1]
    mat <- assay(se, use)
  }
  mat <- as.matrix(mat)
  mode(mat) <- "numeric"
  mat
}

get_symbol_map <- function(se) {
  se <- unwrap_se(se)
  rd <- as.data.frame(rowData(se))

  candidate_symbol_cols <- c("name","Gene","gene","gene_name","SYMBOL","Symbol","Protein","protein","PG.Genes")
  symbol_col <- candidate_symbol_cols[candidate_symbol_cols %in% colnames(rd)][1]

  sym <- if (!is.na(symbol_col)) as.character(rd[[symbol_col]]) else rownames(se)

  m <- setNames(seq_along(sym), sym)
  m <- m[!duplicated(names(m))]
  m
}

cal_z_row <- function(mat) t(scale(t(mat)))

gg_color_hue <- function(n) {
  hues <- seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}

# ----------------------------
# 1) Extract matrices and subset proteins
# ----------------------------
se_mgh  <- unwrap_se(se_mgh)
se_8330 <- unwrap_se(se_8330)

mat_mgh_raw  <- get_assay_matrix(se_mgh)
mat_8330_raw <- get_assay_matrix(se_8330)

map_mgh  <- get_symbol_map(se_mgh)
map_8330 <- get_symbol_map(se_8330)

valid_proteins <- intersect(proteins_to_plot, intersect(names(map_mgh), names(map_8330)))
if (length(valid_proteins) < 2) stop("Fewer than 2 proteins found in BOTH backgrounds.")

mat_mgh_sub  <- mat_mgh_raw[ unname(map_mgh[valid_proteins]),  , drop=FALSE]
mat_8330_sub <- mat_8330_raw[unname(map_8330[valid_proteins]), , drop=FALSE]

rownames(mat_mgh_sub)  <- valid_proteins
rownames(mat_8330_sub) <- valid_proteins

# ----------------------------
# 2) IMPORTANT FIX: make sample names unique BEFORE combining
# ----------------------------
colnames(mat_mgh_sub)  <- paste0("MGH_",  colnames(mat_mgh_sub))
colnames(mat_8330_sub) <- paste0("8330_", colnames(mat_8330_sub))

# ----------------------------
# 3) Z-score separately by background, recombine
# ----------------------------
z_mgh  <- cal_z_row(mat_mgh_sub)
z_8330 <- cal_z_row(mat_8330_sub)
z_matrix_combined <- cbind(z_mgh, z_8330)

# ----------------------------
# 4) Column annotations (genotype/condition + background) and sorting
# ----------------------------
col_anno_mgh  <- as.data.frame(colData(se_mgh))
col_anno_8330 <- as.data.frame(colData(se_8330))

# rownames must match the NEW prefixed sample names
rownames(col_anno_mgh)  <- colnames(z_mgh)   # MGH_* names
rownames(col_anno_8330) <- colnames(z_8330)  # 8330_* names

col_anno_mgh$background  <- "MGH"
col_anno_8330$background <- "8330"

# your "genotype" is stored as condition
col_anno_mgh$genotype  <- as.character(col_anno_mgh$condition)
col_anno_8330$genotype <- as.character(col_anno_8330$condition)

col_annotation_all <- rbind(col_anno_mgh, col_anno_8330)
col_annotation_all <- col_annotation_all[colnames(z_matrix_combined), , drop=FALSE]

# strict factor levels (same as your RNA heatmap)
col_annotation_all$genotype <- factor(
  col_annotation_all$genotype,
  levels = c("DelC","DupC","HetDel","HomoDel","PWSTI","PWSTII","WT")
)
col_annotation_all$background <- factor(
  col_annotation_all$background,
  levels = c("MGH","8330")
)

# sort columns genotype -> background (same style)
col_annotation_sorted <- col_annotation_all[
  order(col_annotation_all$genotype, col_annotation_all$background), , drop=FALSE
]
ordered_samples <- rownames(col_annotation_sorted)

# final matrix in sorted order
heatmap_matrix_final <- z_matrix_combined[, ordered_samples, drop=FALSE]

# ----------------------------
# 5) Colors + pheatmap (match your RNA style)
# ----------------------------
genotype_colors <- setNames(
  gg_color_hue(length(levels(col_annotation_sorted$genotype))),
  levels(col_annotation_sorted$genotype)
)
background_colors <- setNames(
  gg_color_hue(length(levels(col_annotation_sorted$background))),
  levels(col_annotation_sorted$background)
)
ann_colors_final <- list(
  genotype = genotype_colors,
  background = background_colors
)

pheatmap(
  heatmap_matrix_final,

  annotation_col = col_annotation_sorted[, c("genotype","background"), drop=FALSE],
  annotation_colors = ann_colors_final,

  cluster_rows = TRUE,
  cluster_cols = FALSE,

  show_rownames = TRUE,
  show_colnames = TRUE,
  fontsize_row = 8,
  fontsize_col = 8,

  color = colorRampPalette(rev(brewer.pal(n = 10, name = "RdBu")))(100),
  main = "Proteomics (MGH + 8330): Selected Proteins (Z-score by Background)",
  scale = "none",
  breaks = seq(-2, 2, length.out = 101),
  border_color = NA
)
```

